### 2024-04-12 latex/paper.tex
feat: first commit, draft, some results

\documentclass[modern]{aastex631}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{bm}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{float}
\usepackage{soul}
\usepackage{mathtools}

% commands
\newcommand{\TODO}{\texttt{todo}}
\newcommand{\set}[1]{\{\,#1\,\}}
\newcommand{\footlink}[1]{\footnote{\url{#1}}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\let\subsectionautorefname\sectionautorefname
\let\subsubsectionautorefname\sectionautorefname

% no indent
\setlength\parindent{0pt}

% code style
\lstdefinestyle{mystyle}{
    backgroundcolor=\color{white},
    commentstyle=\color{gray!70},
    keywordstyle=\color{Bittersweet},
    stringstyle=\color{RoyalBlue},
    basicstyle=\fontsize{8.5}{13}\fontfamily{DejaVuSansMono-TLF}\selectfont,
    breakatwhitespace=false,         
    breaklines=true,
    rulecolor=\color{black!15},
    numbers=none,
    numberstyle=\fontsize{7}{11}\fontfamily{DejaVuSansMono-TLF}\selectfont\color{gray!50},
    framerule=0pt,
    breakindent=5pt,
    resetmargins=true,
    numbersep=10pt,
    frame=single,
    aboveskip=1em,
    belowskip=1em,
    xleftmargin=6pt,
    framexleftmargin=4pt
}
\lstset{style=mystyle}

\begin{document}

\title{starry 2}

% author info
\author{Lionel J. Garcia}
\author{Soichiro Hattori}
\author{Daniel Foreman-Mackey}
\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY, USA}

\keywords{exoplanet detection methods, stellar activity, time series analysis, gaussian processes regression, computational methods, GPU computing}

\begin{abstract}
\end{abstract}

\section*{Introduction}
tldr: Modeling occultation light curves serve a wide variety of science cases, especially in exoplanetary science. The latest models account for non-uniform surfaces of stars (starry), but given the large dataset available, could be optimized further. We present several optimizations that improve on the starry model.

Stellar light curves and radial velocity models provided an important tool to study exoplanetary systems. At the origin of these models lies the first principles of orbital mechanics, the Kepler equations, and the algorithms used to solve them. On top of that, models like Agol, of occultation light curves extended our capability to study these systems in greater details, using transits. In practice, computing this model and making inference based on these observables has been enabled by the implementation of these models with optimized frameworks and inference tools (pymc3), providing robust implementation of sampling algorithms and other inferrence tools. A good example of this synergy is the development of exoplanet and starry in c++ backed by Pymc3 which enabled major discoveries based on large datasets. With the recent lunch of JWST, we crumble under data. For example, transmission spectra on hundreds of channels, multi-planet systems and a large number of free parameters with their priors. At the same time, machine learning continue to grow, and novel framework appear and lead to application in physics. One of this framework is jax, developped by google, and gaining a huge popularity on the astronomy community. JAX is a hp machine-learning library that pre-compile code to LAX for paralilisation. It allows complex model to be written direclty in python, where usually a lower-level langage like C is preferable, and parallelised on CPU, GPU or any lax-compatible device. Not only it's good because of the higher level syntax of python that made it popular in science. But also because of the capability to run expansive models on GPU, a trend observed in other field of ML and DL where scaling is required.

In this paper, we describe an implementation of the starry formalism in jax, allowing for the inference of a large number of parameters on complex models, exploiting the latest dev in ML machinery.

\newpage
\section{starry}\label{starry}

\section{optimization}\label{optimization}
\subsection{Limb-darkening multiplicative maps}
\subsection{Spherical harmonics rotations}

% tldr: rotating SH involves costly matrices computations. Hence its smart to figure out which one to pre-compute. That what RL did, amd it led to 6 elementary rotations. However, now that we use jax we don't care that much, and we can reduce the all thing to a single rotation that is optimized at compiled time.\\\\

With \textit{starry}, every light curve computation at a given time $t$ involves a rotation of the spherical harmonics basis, from the rest-frame of the star (\autoref{fig:rotation_basis}, left) to its sky-projected orientation (\autoref{fig:rotation_basis}, right), with a final rotation relative to the position of the occulting body (see Figure 2. from \citealt{starry}).
\begin{figure}[H]
    \begin{center}
        \includegraphics[width=0.6\textwidth]{../workflows/rotations/figures/rotation_basis.pdf}
        \caption{Rotation of the spherical harmonics basis to place the star in the sky-projected frame of the observer.}
        \label{fig:rotation_basis}
    \end{center}
\end{figure}

The rotation of spherical harmonics involves the computation of Wigner-D-matrices, commonly obtained through robust recursion relations in both degree $l$ and order $m$, but leading to costly computations. Hence, knowing how to decompose the full spherical harmonics rotation using pre-computed rotations is essential to achieve optimal performance in the \textit{starry} framework. \autoref{fig:rotation_starry} shows the 6 elementary rotations currently used in the \texttt{starry} implementation.
\begin{figure}[H]
    \begin{center}
        \includegraphics[width=\textwidth]{../workflows/rotations/figures/rotation_starry.pdf}
        \caption{Consecutive rotations of the spherical harmonics basis in the \texttt{starry} implementation. Each schematic shows a sphere rotated from a previous orientation around an axis displayed as a black vector and an angle shown on top.}
        \label{fig:rotation_starry}
    \end{center}
\end{figure}
In this figure, rotations 1 to 3 correspond to a rotation of the star around its rotation axis, and rotations 3 to 5 place the star on a sky-projected frame. Finally, rotation 6 aligns the spherical harmonics map to the occulting body on the $y$ axis in order to apply Green's theorem in the appropriate basis (Figure 2 from \citealt{starry}). The first thing to notice is that rotations 1 to 3 could be simply reduced to a rotation of angle $\theta$ about the y-axis, turning steps 1 to 3 into one. However, rotations around the pole, i.e.\, with a rotation axis along $z$, have simpler expressions that can be implemented separately. In addition, pre-computing these matrices at lower cost is particularly important for steps 2 and 6, involving a potentially large set of angles $\{\theta\}$ and $\{\theta_z\}$, defined over times $\{t\}$, for which to compute the full light curve. This explains the current decomposition of the complete rotation in six separate steps.\\\\
In \texttt{jaxoplanet}, we compute the Wigner-D-matrices by employing the Risbo recursion relations \citep{Risbo1996} implemented in \texttt{JAX} as part of the \texttt{s2fft} Python package \citep{price:s2fft}. In addition, we merge rotations 3, 4 and 5 (see \autoref{fig:rotation_jaxoplanet}) into a single compound rotation of axis
\begin{equation}
    \bm{v} = \frac{1}{\sqrt{1 - \cos^2{\left(\frac{inc}{2} \right)} \cos^2{\left(\frac{obl}{2}\right)}}} \begin{pmatrix}
        \sin{\left(\frac{inc}{2} \right)} \cos{\left(\frac{obl}{2}\right)}\\
        \sin{\left(\frac{inc}{2} \right)} \sin{\left(\frac{obl}{2}\right)}\\
         - \cos{\left(\frac{inc}{2} \right)} \sin{\left(\frac{obl}{2}\right)}\\
    \end{pmatrix},
\end{equation}
and angle
\begin{equation}
    \omega = 2 \cos^{-1}{\left(\cos{\left(\frac{inc}{2} \right)} \cos{\left(\frac{obl}{2}\right)} \right)}.
\end{equation}
This way, the complete rotation reduces to the four separate steps shown in \autoref{fig:rotation_jaxoplanet}.
\begin{figure}[H]
    \begin{center}
        \includegraphics[width=\textwidth]{../workflows/rotations/figures/rotation_jaxoplanet_1.pdf}
        \caption{Consecutive rotations of the spherical harmonics basis in the \texttt{jaxoplanet} implementation. Each schematic shows a sphere rotated from a previous orientation around an axis displayed as a black vector and an angle shown on top.}
        \label{fig:rotation_jaxoplanet}
    \end{center}
\end{figure}

\subsection{Wigner-D matrices}
\subsection{Numerical integration of the solution vectors}

\section{performance}\label{performances}

\section{Case study: ...}

\bibliography{ref}

\end{document}

### 2024-07-01 latex/paper.tex
rewrite abstract

@@
 
-\title{starry 2}
+\title{Hardware-Accelerated Computation of Stellar Light Curves}
 
@@
 
-\keywords{exoplanet detection methods, stellar activity, time series analysis, gaussian processes regression, computational methods, GPU computing}
+\keywords{}
 
 \begin{abstract}
+    The measurement of stellar fluxes over time serves a wide variety of science cases.
+Transit and occultation light curves, for example, are used to characterize the orbital parameters of stellar systems, and infer the atmospheric properties of their transiting bodies.
+Phase curves, on the other hand, can be used to study the thermal emission of exoplanets, and the atmospheric properties of stars.
+In order to infer these properties, a fast, accurate and robust model of stellar light curves is needed, one that accounts for the non-uniform surface intensity of spherical bodies.
+The analytical framework known as \texttt{starry} satisfies these requirements, and was successfully applied over the years following its development.
+However, as datasets and model complexity grow, the tractable inference of these parameters becomes increasingly challenging, and often motivates biased approximations.
+In this paper, we present a new implementation of the \texttt{starry} framework in \texttt{jax}, a high-performance machine-learning library designed for large-scale computations.
+We describe the changes made to the original version of \texttt{starry} and evaluate the performance of its new implementation that we release in the open-source Python package \texttt{jaxoplanet}.
+Through this implementation, we provide the community with a differentiable model of stellar light curves, and enable the use of powerful machine-learning tools available in the \texttt{jax} ecosystem.
+
 \end{abstract}
@@
 \section*{Introduction}
-tldr: Modeling occultation light curves serve a wide variety of science cases, especially in exoplanetary science.
-The latest models account for non-uniform surfaces of stars (starry), but given the large dataset available, could be optimized further.
-We present several optimizations that improve on the starry model.
 
-Stellar light curves and radial velocity models provided an important tool to study exoplanetary systems.
+% the computation of \textit{starry} light curves can be computationally expensive, especially when considering the large number of parameters that need to be inferred from the data.
+In this paper, we present a new implementation of the \textit{starry} model that leverages the \texttt{jax} library to perform hardware-accelerated computations on GPUs.
+We show that this implementation can be used to compute light curves up to 100 times faster than the original implementation, and we demonstrate its application to the inference of the orbital parameters of exoplanets from transit light curves.
+
+% Stellar light curves and radial velocity models provided an important tool to study exoplanetary systems.
 At the origin of these models lies the first principles of orbital mechanics, the Kepler equations, and the algorithms used to solve them.
@@
 
-In this paper, we describe an implementation of the starry formalism in jax, allowing for the inference of a large number of parameters on complex models, exploiting the latest dev in ML machinery.
+% In this paper, we describe an implementation of the starry formalism in jax, allowing for the inference of a large number of parameters on complex models, exploiting the latest dev in ML machinery.
 
@@
 The first thing to notice is that rotations 1 to 3 could be simply reduced to a rotation of angle $\theta$ about the y-axis, turning steps 1 to 3 into one.
-However, rotations around the pole, i.e.\, with a rotation axis along $z$, have simpler expressions that can be implemented separately.
+However, rotations around the pole, i.e.\,with a rotation axis along $z$, have simpler expressions that can be implemented separately.
 In addition, pre-computing these matrices at lower cost is particularly important for steps 2 and 6, involving a potentially large set of angles $\{\theta\}$ and $\{\theta_z\}$, defined over times $\{t\}$, for which to compute the full light curve.
@@
 
-\subsection{Wigner-D matrices}
-\subsection{Numerical integration of the solution vectors}
+\subsection{Computation of the solution vectors}
+\subsection{Autodiff?}
 

### 2024-07-02 latex/paper.tex
proper workflow for error computation

@@
 Phase curves, on the other hand, can be used to study the thermal emission of exoplanets, and the atmospheric properties of stars.
-In order to infer these properties, a fast, accurate and robust model of stellar light curves is needed, one that accounts for the non-uniform surface intensity of spherical bodies.
+In order to infer these properties, a fast, accurate and robust model of stellar light curves is needed, one that accounts for the non-uniform surface intensity of spherical bodies and their mutual occultations.
 The analytical framework known as \texttt{starry} satisfies these requirements, and was successfully applied over the years following its development.
@@
 
-\subsection{Computation of the solution vectors}
+\subsection{Computation of the solution vectors}\label{solution_vectors}
+
 \subsection{Autodiff?}
 
-\section{performance}\label{performances}
+\newpage
+\section{Performance}\label{performances}
+
+As described in \autoref{solution_vectors}, we compute \textit{starry} light curves using the Gauss Legendre quadrature to approximate the $\mathcal{P}$ integral involved in the expression of the solution vector $s_{n}$.
+Hence, the precision of the computed light curve depends on the order of the Legendre polynomial which defines the number of points used to approximate $\mathcal{P}$.
+For this reason, users must carefully set the order parameter to reach the precision required for their application, keeping in mind that a higher order leads to a longer computation time.
+In this section, we evaluate the performance of \texttt{jaxoplanet} against the C++ implementation of \texttt{starry}.\\\\
+\autoref{fig:relative_error_1} shows the relative numerical error in the flux of a linearly limb-darkened star transited by an opaque companion of radius $r$ over a range of impact parameters $b$.
+Evaluated for different Legendre polynomial orders $n$, this figure shows the decrease of the relative error with the order of the Gauss-Legendre approximation.\\\\
+At a fixed occultor radius, \autoref{fig:relative_error_2} shows the time required to compute the flux of the system for a range of $N$ impact parameters $b$, plotted against the maximum relative error reached, the order $n$ of the Gauss-Legendre approximation, and the hardware used for the computation.
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[width=\textwidth]{../workflows/performance/relative_error_1.png}
+        \caption{}
+        \label{fig:relative_error_1}
+    \end{center}
+\end{figure}
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[width=\textwidth]{../workflows/performance/relative_error_2.pdf}
+        \caption{}
+        \label{fig:relative_error_2}
+    \end{center}
+\end{figure}
+These computations were run on an apple M1 CPU and an NVIDIA ... GPU. Thanks to the use of JAX, no harware-specific... \\\\
+
+By computing $\mathcal{P}$ numerically, we show that our vectorized implementation reaches machine-precision on a GPU, up to an order of magnitude faster than the current C++ implementation of \texttt{starry} ran on a CPU.\\\\
+
+Plot of the precision versus speed of the computed light curves.
+
+Plot of the solution vectors computation similar to figure 11 and 12 of luger (maybe in the appendix)
+
+\subsection{Speed}
 

### 2024-07-02 workflows/performance/scripts/flux_jaxoplanet.py

@@ -1,0 +1,125 @@
+import jax
+
+jax.config.update("jax_platform_name", "cpu")
+jax.config.update("jax_enable_x64", True)
+
+import numpy as np
+from jaxoplanet.experimental.starry import Surface
+import jax.numpy as jnp
+import timeit
+
+data = dict(np.load(snakemake.input[0]))
+order = int(snakemake.wildcards.order)
+u = data["u"]
+surface = Surface(period=None, u=(u))
+
+from jaxoplanet.experimental.starry.light_curves import *
+
+
+# TODO: figure out the sparse matrices (and Pijk) to avoid todense()
+def surface_light_curve(
+    map,
+    r: float = None,
+    xo: float = None,
+    yo: float = None,
+    zo: float = None,
+    theta: float = 0.0,
+    order: int = 20,
+):
+    """Light curve of an occulted map.
+
+    Args:
+        map (Map): map object
+        r (float or None): radius of the occulting body, relative to the current map
+           body
+        xo (float or None): x position of the occulting body, relative to the current
+           map body
+        yo (float or None): y position of the occulting body, relative to the current
+           map body
+        zo (float or None): z position of the occulting body, relative to the current
+           map body
+        theta (float): rotation angle of the map
+
+    Returns:
+        ArrayLike: flux
+    """
+    rT_deg = rT(map.deg)
+
+    # no occulting body
+    if r is None:
+        b_rot = True
+        theta_z = 0.0
+        x = rT_deg
+
+    # occulting body
+    else:
+        b = jnp.sqrt(jnp.square(xo) + jnp.square(yo))
+        b_rot = jnp.logical_or(jnp.greater_equal(b, 1.0 + r), jnp.less_equal(zo, 0.0))
+        b_occ = jnp.logical_not(b_rot)
+        theta_z = jnp.arctan2(xo, yo)
+        sT = solution_vector(map.deg, order=order)(b, r)
+
+        # scipy.sparse.linalg.inv of a sparse matrix[[1]] is a non-sparse [[1]], hence
+        # `from_scipy_sparse`` raises an error (case deg=0)
+        if map.deg > 0:
+            A2 = scipy.sparse.linalg.inv(A2_inv(map.deg))
[…]

### 2024-07-02 workflows/performance/scripts/flux_starry.py

@@ -1,0 +1,37 @@
+import starry
+import numpy as np
+import timeit
+
+starry.config.lazy = False
+
+n = int(snakemake.wildcards.N)
+
+ms = starry.Map(ydeg=1, udeg=1)
+u = (1.0,)
+ms[1:] = u
+
+r = np.linspace(0, 2, n)
+b = np.linspace(0, 2, n)
+
+ro = 0.5
+
+calculated = np.zeros((n, n))
+
+for i, _r in enumerate(r):
+    expected = ms.flux(ro=_r, yo=b, xo=0.0, zo=10.0)
+    calculated[i, :] = expected
+
+
+def timeit_f(strf, number=100):
+    times = (
+        np.array(timeit.repeat(f"{strf}", number=number, globals=globals()))[1:]
+        / number
+    )
+    return np.median(times), times.std()
+
+
+reference_time = timeit_f(f"ms.flux(ro={ro}, yo=b, xo=0.0, zo=10.0)")[0]
+
+np.savez(
+    snakemake.output[0], r=r, b=b, value=calculated, ro=ro, u=u, time=reference_time
+)

### 2024-07-02 workflows/performance/scripts/plot_error_n.py

@@ -1,0 +1,61 @@
+import matplotlib.pyplot as plt
+import matplotlib.colors as colors
+import numpy as np
+
+orders = snakemake.params.orders
+
+data = {
+    "starry": np.load(snakemake.input["starry"]),
+    "jaxoplanet": {
+        orders[i]: np.load(snakemake.input["jaxoplanet"][i]) for i in range(len(orders))
+    },
+}
+
+r = data["jaxoplanet"][orders[0]]["r"]
+b = data["jaxoplanet"][orders[0]]["b"]
+R, B = np.meshgrid(r, b)
+
+
+def colorbar(mappable):
+    from mpl_toolkits.axes_grid1 import make_axes_locatable
+    import matplotlib.pyplot as plt
+
+    last_axes = plt.gca()
+    ax = mappable.axes
+    fig = ax.figure
+    divider = make_axes_locatable(ax)
+    cax = divider.append_axes("right", size="5%", pad=0.05)
+    cbar = fig.colorbar(mappable, cax=cax)
+    plt.sca(last_axes)
+    return cbar
+
+
+fig, axes = plt.subplots(1, len(orders), figsize=(7.5, 2.5))
+norm = colors.LogNorm(vmin=1e-15, vmax=1e-8)
+for i, order in enumerate(orders):
+    ax = axes[i]
+    ax.set_aspect("equal", "box")
+    errors = np.abs(data["starry"]["value"] - data["jaxoplanet"][order]["value"])
+    errors[errors == 0] = 1e-20
+    errors[np.logical_not(np.isfinite(errors))] = 1e-20
+
+    # norm = colors.LogNorm(vmin=1e-15, vmax=np.max(np.abs(errors[i])))
+    pcm = ax.pcolor(R, B, errors.T, norm=norm, cmap="PuBu_r", shading="auto")
+    k = dict(color="k", linestyle="-", linewidth=1)
+    ax.axline((0, 0), slope=1, **k)
+    ax.axline((1, 0), slope=1, **k)
+    ax.axline((0, 1), slope=1, **k)
+    ax.axline((0, 1), slope=-1, **k)
+    if i > 0:
+        ax.set_yticklabels([])
+    else:
+        ax.set_ylabel("b")
+    ax.set_xlabel("r")
+    ax.set_title(f"$n$ = {order}")
+
+# cm_ax = fig.add_axes([1.5, 0.1, 0.02, 0.8])
[…]

### 2024-07-02 latex/paper.tex
starting proper perfomance workflow + gitignore generated data

@@
 \newpage
-\section{starry}\label{starry}
+\section{starry formalism}\label{starry}
 
@@
 
-\subsection{Autodiff?}
-
 \newpage
@@
     \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/performance/relative_error_1.png}
+        \includegraphics[width=\textwidth]{../workflows/performance/figures/error_n.png}
         \caption{}
@@
     \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/performance/relative_error_2.pdf}
+        \includegraphics[width=\textwidth]{../workflows/performance/figures/time_n.pdf}
         \caption{}
@@
 
-\subsection{Speed}
-
 \section{Case study: ...}

### 2024-07-02 workflows/performance/scripts/flux_jaxoplanet.py

@@ -1,4 +1,5 @@
 import jax
+from tqdm import tqdm
 
 jax.config.update("jax_platform_name", "cpu")
 jax.config.update("jax_enable_x64", True)
@@ -6,7 +7,6 @@ jax.config.update("jax_enable_x64", True)
 import numpy as np
 from jaxoplanet.experimental.starry import Surface
 import jax.numpy as jnp
-import timeit
 
 data = dict(np.load(snakemake.input[0]))
 order = int(snakemake.wildcards.order)
@@ -14,6 +14,18 @@ u = data["u"]
 surface = Surface(period=None, u=(u))
 
 from jaxoplanet.experimental.starry.light_curves import *
+
+import timeit
+
+
+def timeit_f(strf, number=2, repeat=2):
+    times = (
+        np.array(
+            timeit.repeat(f"{strf}", repeat=repeat, number=number, globals=globals())
+        )[1:]
+        / number
+    )
+    return np.median(times)
 
 
 # TODO: figure out the sparse matrices (and Pijk) to avoid todense()
@@ -103,16 +115,8 @@ expected = data["value"]
 result = np.zeros_like(expected)
 
 
-def timeit_f(strf, number=100):
-    times = (
-        np.array(timeit.repeat(f"{strf}", number=number, globals=globals()))[1:]
-        / number
-    )
-    return np.median(times)
-
-
 def flux(order):
-    for i, _r in enumerate(r):
+    for i, _r in enumerate(tqdm(r)):
         calc = f(surface, _r, 0.0, b, 10.0, 0.0, order)
         result[i, :] = calc
 

### 2024-07-02 workflows/performance/scripts/flux_starry.py

@@ -1,6 +1,18 @@
+from tqdm import tqdm
 import starry
 import numpy as np
 import timeit
+
+
+def timeit_f(strf, number=2, repeat=1):
+    times = (
+        np.array(
+            timeit.repeat(f"{strf}", repeat=repeat, number=number, globals=globals())
+        )
+        / number
+    )
+    return np.median(times)
+
 
 starry.config.lazy = False
 
@@ -17,20 +29,12 @@ ro = 0.5
 
 calculated = np.zeros((n, n))
 
-for i, _r in enumerate(r):
+for i, _r in enumerate(tqdm(r)):
     expected = ms.flux(ro=_r, yo=b, xo=0.0, zo=10.0)
     calculated[i, :] = expected
 
 
-def timeit_f(strf, number=100):
-    times = (
-        np.array(timeit.repeat(f"{strf}", number=number, globals=globals()))[1:]
-        / number
-    )
-    return np.median(times), times.std()
-
-
-reference_time = timeit_f(f"ms.flux(ro={ro}, yo=b, xo=0.0, zo=10.0)")[0]
+reference_time = timeit_f(f"ms.flux(ro={ro}, yo=b, xo=0.0, zo=10.0)")
 
 np.savez(
     snakemake.output[0], r=r, b=b, value=calculated, ro=ro, u=u, time=reference_time

### 2024-07-02 workflows/performance/scripts/plot_error_n.py

@@ -30,8 +30,8 @@ def colorbar(mappable):
     return cbar
 
 
-fig, axes = plt.subplots(1, len(orders), figsize=(7.5, 2.5))
-norm = colors.LogNorm(vmin=1e-15, vmax=1e-8)
+fig, axes = plt.subplots(1, len(orders), figsize=(8.5, 2.5))
+# norm = colors.LogNorm(vmin=1e-15, vmax=1e-8)
 for i, order in enumerate(orders):
     ax = axes[i]
     ax.set_aspect("equal", "box")
@@ -39,7 +39,7 @@ for i, order in enumerate(orders):
     errors[errors == 0] = 1e-20
     errors[np.logical_not(np.isfinite(errors))] = 1e-20
 
-    # norm = colors.LogNorm(vmin=1e-15, vmax=np.max(np.abs(errors[i])))
+    norm = colors.LogNorm(vmin=1e-15, vmax=np.max(np.abs(errors)))
     pcm = ax.pcolor(R, B, errors.T, norm=norm, cmap="PuBu_r", shading="auto")
     k = dict(color="k", linestyle="-", linewidth=1)
     ax.axline((0, 0), slope=1, **k)
@@ -52,10 +52,10 @@ for i, order in enumerate(orders):
         ax.set_ylabel("b")
     ax.set_xlabel("r")
     ax.set_title(f"$n$ = {order}")
+    colorbar(pcm)
 
 # cm_ax = fig.add_axes([1.5, 0.1, 0.02, 0.8])
 # place colorbar at cm_ax
 
-colorbar(pcm)
 plt.tight_layout()
 plt.savefig(snakemake.output[0], dpi=250)

### 2024-07-02 workflows/performance/scripts/plot_time_NN.py

@@ -1,0 +1,20 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+
+Ns = snakemake.params.Ns
+data = {
+    "starry": [np.load(f)["time"] for f in snakemake.input["starry"]],
+    "jaxoplanet": [np.load(f)["time"] for f in snakemake.input["jaxoplanet"]],
+}
+
+plt.plot(Ns, data["starry"], ".-", label="starry (CPU)")
+plt.plot(Ns, data["jaxoplanet"], ".-", label="jaxoplanet (CPU)")
+plt.xlabel("N")
+plt.ylabel("computation time (s)")
+plt.xscale("log")
+plt.yscale("log")
+
+plt.legend()
+plt.tight_layout()
+plt.savefig(snakemake.output[0])

### 2024-07-02 workflows/performance/scripts/plot_time_n.py

@@ -1,0 +1,80 @@
+import matplotlib.pyplot as plt
+import matplotlib.colors as colors
+import numpy as np
+
+orders = snakemake.params.orders
+
+data = {
+    "starry": {
+        1000: np.load(snakemake.input["starry_1000"]),
+        # 10000: np.load(snakemake.input["starry_10000"]),
+    },
+    "jaxoplanet": {
+        1000: {
+            orders[i]: np.load(snakemake.input["jaxoplanet_1000"][i])
+            for i in range(len(orders))
+        },
+        # 10000: {
+        #     orders[i]: np.load(snakemake.input["jaxoplanet_10000"][i])
+        #     for i in range(len(orders))
+        # },
+    },
+}
+
+
+def get_data(N):
+    precisions = np.array(
+        [
+            np.abs(
+                data["jaxoplanet"][N][order]["value"] - data["starry"][N]["value"]
+            ).max()
+            for order in orders
+        ]
+    )
+
+    times = np.array([data["jaxoplanet"][N][order]["time"] for order in orders])
+
+    reference_precision = np.min(precisions)
+    reference_time = data["starry"][N]["time"]
+    return precisions, times, reference_precision, reference_time
+
+
+plt.figure(figsize=(5, 3.5))
+
+plt.subplot(111)
+precisions, times, reference_precision, reference_time = get_data(1000)
+plt.plot(preci sions, times * 1e6, ".-", c="k", label="jaxoplanet (CPU)")
+plt.plot(reference_precision, reference_time * 1e6, "+", label="starry (CPU)")
+plt.axhline(reference_time * 1e6, c="C0", ls="-", alpha=0.2)
+plt.axvline(reference_precision, c="C0", ls="-", alpha=0.2)
+plt.xscale("log")
+plt.xlim(1e-7, 0.5e-15)
+plt.ylim(0)
+plt.xlabel("relative error")
+plt.ylabel("computation time (µs)")
+
+for i, order in enumerate(orders):
+    plt.text(precisions[i], times[i] * 1e6 + 20, f"{order}", fontsize=9, ha="right")
+
+# plt.title(f"N = 1000")
+plt.legend(loc="upper left")
+
+# plt.subplot(122)
+# precisions, times, reference_precision, reference_time = get_data(10000)
[…]

### 2024-07-17 latex/paper.tex
feat: add precision computations

--- 2024-07-01 latex/paper.tex
@@
 % commands
-\newcommand{\TODO}{\texttt{todo}}
+\newcommand{\todo}[1]{\colorbox{red}{\textcolor{white}{\texttt{TO DO}}}\,}
 \newcommand{\set}[1]{\{\,#1\,\}}
@@
 \section*{Introduction}
+
+\begin{equation}\label{eq:starry}
+    f = \bm{s^TARR'y}
+\end{equation}
 
@@
 
-\subsection{Computation of the solution vectors}
-\subsection{Autodiff?}
+\subsection{Computation of the solution vectors}\label{solution_vectors}
 
-\section{performance}\label{performances}
+\newpage
+\section{Performance}\label{performances}
+Although \textit{starry} occultation light curves have analytical expressions, their practical implementation relies on numerical approximations chosen to balance precision and computation time.
+This is the case, for example, of the solution vector $\bm{s}$, computed using numerical series expansions in \texttt{starry} (see \todo{} in \citealt{starry}) and numerical integration in \texttt{jaxoplanet} (see \autoref{solution_vectors}).
+Other expressions, such as the Wigner-D matrices used in the rotation of the spherical harmonics basis, are computed using reccurence relations prone to the accumulation of numerical errors and instability.\\\\
+In this section, we evaluate the accuracy and speed of the \texttt{jaxoplanet} implementation and compare it with the C++ implementation of \texttt{starry}.
+
+\subsection{Precision}
+We evaluate the precision of the \texttt{starry} and \texttt{jaxoplanet} terms against quantities computed at arbitrary precision, using a ground truth version of the code implemented with the \texttt{mpmath} Python library\footlink{https://mpmath.org/}.
+The precision of the overall flux $f$ can only be understood by considering the precision of the terms involved in \autoref{eq:starry}: the solution vector $\bm{s}$ ($\bm{r}$ if no occultation), the basis matrices $\bm{A}$ and $\bm{B}$, and the Wigner-D rotation tensor $\bm{R}$.
+Although we show in \todo{add an appendix showing errors for these quantities} that the precision of the \texttt{starry} and \texttt{jaxoplanet} implentations differ for many of these terms, we focus this section on the solution vector $\bm{s}$ and the overall flux $f$.\\\\
+Assuming that we perform the numerical integration of the solution vector $\bm{s}$ at a relatively high order ($n=500$; see \autoref{solution_vectors}), \autoref{fig:precision_s} shows that our reimplentation of \texttt{starry} reaches relative errors lower than 1 part per billion for $\bm{s}$ and the overall flux $f$, considering both a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$.
+For reference, \autoref{fig:precision_s_starry} shows comparable errors on the same quantities computed with the C++ implentation of \texttt{starry}.
+In addition, \autoref{fig:precision_degree} shows the evolution of the error on $f$ for increasing degrees of the spherical harmonics, both for \texttt{starry} and \texttt{jaxoplanet}.
+With these results, we validate the precision of \texttt{jaxoplanet} ans show it is on per with the legacy C++ \texttt{starry} implentation\footnote{version \todo{add version number of starry}} described in \cite{starry}.\\\\
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[width=\textwidth]{../workflows/precision/figures/error_jax.pdf}
+        \caption{}
+        \label{fig:precision_s}
+    \end{center}
+\end{figure}
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[width=\textwidth]{../workflows/precision/figures/error_degree.pdf}
+        \caption{}
+        \label{fig:precision_degree}
+    \end{center}
+\end{figure}
+As described in \autoref{solution_vectors}, we compute \textit{starry} light curves using the Gauss Legendre quadrature to approximate the $\mathcal{P}$ integral involved in the expression of the solution vector $\bm{s}$.
+Hence, the precision of the computed flux $f$ depends on the order of the Legendre polynomial which defines the number of points used to approximate $\mathcal{P}$.
+For this reason, users must carefully set the order parameter to reach the precision required for their application, as shown in \autoref{fig:relative_error_1}.\\\\
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[width=\textwidth]{../workflows/speed/figures/error_rb.png}
+        \caption{relative numerical error in the flux $f$ of a linearly limb-darkened star transited by an opaque companion over the $(r, b)$ parameter space.
+The error is evaluated for different orders $n$ of the Gauss-Legendre quadrature and computed against \texttt{starry} for speed and convenience.
+Note that \autoref{fig:precision_s} was produced using $n=500$ (right plot) for which $\bm{s}$ reaches machine precision.}
+        \label{fig:relative_error_1}
+    \end{center}
+\end{figure}
+At a fixed occultor radius, \autoref{fig:relative_error_2} shows the time required to compute the flux of the system for a range of $N$ impact parameters $b$, plotted against the maximum relative error reached, the order $n$ of the Gauss-Legendre approximation, and the hardware used for the computation.
+
+% \begin{figure}[H]
+%     \begin{center}
+%         \includegraphics[width=\textwidth]{../workflows/performance/figures/error_vs_time.pdf}
+%         \caption{}
+%         \label{fig:relative_error_2}
+%     \end{center}
+% \end{figure}
+These computations were run on an apple M1 CPU and an NVIDIA ... GPU. Thanks to the use of JAX, no harware-specific... \\\\
+
+By computing $\mathcal{P}$ numerically, we show that our vectorized implementation reaches machine-precision on a GPU, up to an order of magnitude faster than the current C++ implementation of \texttt{starry} ran on a CPU.\\\\
+
+Plot of the precision versus speed of the computed light curves.
+
+Plot of the solution vectors computation similar to figure 11 and 12 of luger (maybe in the appendix)
+
+\subsection{Speed}\label{speed}
 
 \section{Case study: ...}
+
+\appendix
+
+\section{Performance}
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[width=\textwidth]{../workflows/precision/figures/error_starry.pdf}
+        \caption{}
+        \label{fig:precision_s_starry}
+    \end{center}
+\end{figure}
 

### 2024-07-17 workflows/precision/scripts/concatenate_fs.py

@@ -1,0 +1,29 @@
+import pickle
+import numpy as np
+from tqdm import tqdm
+from jaxoplanet.experimental.starry.mpcore import mp
+
+flux = []
+s = []
+
+bs = np.load(snakemake.input[0], allow_pickle=True)["bs"]
+bs = np.abs(np.concatenate(bs))
+
+for file in tqdm(snakemake.input.i):
+    with open(file, "rb") as f:
+        data = np.load(f, allow_pickle=True)
+        flux.append(data["f"].tolist())
+
+s_data = np.load(snakemake.input.s, allow_pickle=True)
+
+for b in bs:
+    s.append(s_data[b])
+
+s = mp.matrix(np.array(s)[:, :, 0].tolist())
+flux = mp.matrix(flux).T
+
+with open(snakemake.output[0], "wb") as f:
+    pickle.dump(
+        {"f": flux, "s": s},
+        f,
+    )

### 2024-07-17 workflows/precision/scripts/jax_fs.py

@@ -1,0 +1,89 @@
+import jax
+
+jax.config.update("jax_enable_x64", True)
+
+from jaxoplanet.experimental.starry.solution import solution_vector
+from jaxoplanet.experimental.starry.light_curves import *
+import numpy as np
+import jax.numpy as jnp
+from jaxoplanet.experimental.starry.mpcore.solution import rT
+from jaxoplanet.experimental.starry.mpcore.basis import A1, A2_inv
+from jaxoplanet.experimental.starry.mpcore.utils import to_numpy
+
+
+def surface_light_curve(
+    deg,
+    y,
+    inc,
+    obl,
+    r: float = None,
+    xo: float = None,
+    yo: float = None,
+    zo: float = None,
+    theta: float = 0.0,
+    order: int = 20,
+):
+    rT_deg = to_numpy(rT(deg))
+
+    # occulting body
+    if True:
+        b = jnp.sqrt(jnp.square(xo) + jnp.square(yo))
+        b_rot = jnp.logical_or(jnp.greater_equal(b, 1.0 + r), jnp.less_equal(zo, 0.0))
+        b_occ = jnp.logical_not(b_rot)
+        theta_z = jnp.arctan2(xo, yo)
+        sT = solution_vector(deg, order=order)(b, r)
+
+        if deg > 0:
+            A2 = to_numpy(A2_inv(deg) ** -1)
+        else:
+            A2 = jnp.array([1])
+
+        x = jnp.where(b_occ, sT @ A2, rT_deg)
+
+    rotated_y = left_project(deg, inc, obl, theta, theta_z, y)
+
+    A1_val = to_numpy(A1(deg))
+    p_y = Pijk.from_dense(A1_val @ rotated_y, degree=deg)
+
+    b_full_occ = jnp.logical_and(r >= 1.0, b <= r - 1)
+    norm = jnp.where(b_full_occ, 0.0, 1.0)
+
+    return p_y.tosparse() @ x * norm
+
+
+def jax_flux(deg, order=20, inc=np.pi / 2, obl=0.0):
+
+    def impl(y, b, r, theta):
+        return surface_light_curve(
+            deg, y, inc, obl, r, 0.0, b, 10.0, theta=theta, order=order
+        )
+
+    return jax.jit(impl)
+
+
+r = float(snakemake.wildcards.r)
+l_max = snakemake.params.l_max
[…]

### 2024-07-17 workflows/precision/scripts/make_b.py

@@ -1,0 +1,83 @@
+import numpy as np
+
+
+def b_range(r, logdelta=-3, logeps=-6, res=1):
+    delta = 10**logdelta
+    eps = 10**logeps
+    if r > 1:
+        bs = [
+            np.linspace(r - 1 + 1e-8, r - 1 + eps, res),
+            np.linspace(r - 1 + eps, r - 1 + delta, res),
+            np.linspace(r - 1 + delta, r - delta, 3 * res),
+            np.linspace(r - delta, r - eps, res),
+            np.linspace(r - eps, r + eps, res),
+            np.linspace(r + eps, r + delta, res),
+            np.linspace(r + delta, r + 1 - delta, 3 * res),
+            np.linspace(r + 1 - delta, r + 1 - eps, res),
+            np.linspace(r + 1 - eps, r + 1 - 1e-8, res),
+        ]
+        labels = [
+            r"$r - 1$",
+            r"$r - 1 + 10^{%d}$" % logeps,
+            r"$r - 1 + 10^{%d}$" % logdelta,
+            r"$r - 10^{%d}$" % logdelta,
+            r"$r - 10^{%d}$" % logeps,
+            r"$r + 10^{%d}$" % logeps,
+            r"$r + 10^{%d}$" % logdelta,
+            r"$r + 1 - 10^{%d}$" % logdelta,
+            r"$r + 1 - 10^{%d}$" % logeps,
+            r"$r + 1$",
+        ]
+    else:
+        bs = [
+            np.linspace(1e-8, eps, res),
+            np.linspace(eps, delta, res),
+            np.linspace(delta, r - delta, 3 * res),
+            np.linspace(r - delta, r - eps, res),
+            np.linspace(r - eps, r + eps, res),
+            np.linspace(r + eps, r + delta, res),
+            np.linspace(r + delta, 1 - r - delta, 3 * res),
+            np.linspace(1 - r - delta, 1 - r - eps, res),
+            np.linspace(1 - r - eps, 1 - r + eps, res),
+            np.linspace(1 - r + eps, 1 - r + delta, res),
+            np.linspace(1 - r + delta, 1 - delta, 3 * res),
+            np.linspace(1 - delta, 1 - eps, res),
+            np.linspace(1 - eps, 1 + eps, res),
+            np.linspace(1 + eps, 1 + delta, res),
[…]

### 2024-07-17 workflows/precision/scripts/num_fs.py

@@ -1,0 +1,79 @@
+import numpy as np
+
+r = float(snakemake.wildcards.r)
+l_max = snakemake.params.l_max
+bs = np.load(snakemake.input[0], allow_pickle=True)["bs"]
+bs = np.abs(np.concatenate(bs))
+y = snakemake.params.y
+
+num_matrices = np.load(snakemake.input[1], allow_pickle=True)
+ST = np.load(snakemake.input[2], allow_pickle=True)
+
+from jaxoplanet.experimental.starry.mpcore import mp
+from jaxoplanet.experimental.starry.mpcore.rotation import (
+    dot_rotation_matrix,
+    dot_rz,
+)
+
+y = mp.matrix(y.tolist())
+
+
+def flux_function(l_max, inc, obl):
+    _A1 = num_matrices["A1"]
+    _A2 = num_matrices["A2"]
+
+    R_px = num_matrices["R_px"]
+    R_mx = num_matrices["R_mx"]
+    R_obl = num_matrices["R_obl"]
+    R_inc = num_matrices["R_inc"]
+
+    def rotate_y(y, phi):
+        y_rotated = dot_rotation_matrix(l_max, None, None, R_px)(y)
+        y_rotated = dot_rz(l_max, phi)(y_rotated)
+        y_rotated = dot_rotation_matrix(l_max, None, None, R_mx)(y_rotated)
+        y_rotated = dot_rotation_matrix(l_max, None, None, R_obl)(y_rotated)
+        y_rotated = dot_rotation_matrix(l_max, None, None, R_inc)(y_rotated)
+        return y_rotated
+
+    def occ_flux(y, b, r, phi):
+        xo = 0.0
+        yo = b
+        theta_z = mp.atan2(xo, yo)
+        _sT = ST[b]
+        x = _sT.T @ _A2
+        y_rotated = rotate_y(y, phi)
+        y_rotated = dot_rz(l_max, theta_z)(y_rotated)
+        py = (_A1 @ y_rotated).T
+        return (py @ x.T)[0], _sT
+
+    def impl(y, b, r, phi):
+        if abs(b) >= (r + 1):
+            raise ValueError("Impact parameter out of bounds")
+        else:
+            if r > 1 and abs(b) <= (r - 1):
+                raise ValueError("Impact parameter out of bounds")
+            else:
+                return occ_flux(y, b, r, phi)
+
+    return impl
+
+
+def is_occ(b, r):
[…]

### 2024-07-17 workflows/precision/scripts/num_fs_matrices.py

@@ -1,0 +1,33 @@
+l_max = snakemake.params.l_max
+
+from jaxoplanet.experimental.starry.mpcore import mp
+from jaxoplanet.experimental.starry.mpcore.basis import A1, A2_inv
+from jaxoplanet.experimental.starry.mpcore.rotation import R
+
+
+inc = mp.pi / 2
+obl = 0.0
+
+_A1 = A1(l_max)
+_A2 = A2_inv(l_max) ** -1
+
+R_px = R(l_max, (1.0, 0.0, 0.0), -0.5 * mp.pi)
+R_mx = R(l_max, (1.0, 0.0, 0.0), 0.5 * mp.pi)
+R_obl = R(l_max, (0.0, 0.0, 1.0), -obl)
+R_inc = R(l_max, (-mp.cos(obl), -mp.sin(obl), 0.0), (0.5 * mp.pi - inc))
+
+import numpy as np
+import pickle
+
+with open(snakemake.output[0], "wb") as f:
+    pickle.dump(
+        {
+            "A1": _A1,
+            "A2": _A2,
+            "R_px": R_px,
+            "R_mx": R_mx,
+            "R_obl": R_obl,
+            "R_inc": R_inc,
+        },
+        f,
+    )

### 2024-07-17 workflows/precision/scripts/num_sT.py

@@ -1,0 +1,16 @@
+from jaxoplanet.experimental.starry.mpcore.solution import sT
+import numpy as np
+from tqdm import tqdm
+
+l_max = snakemake.params.l_max
+r = float(snakemake.wildcards.r)
+bs = np.load(snakemake.input[0], allow_pickle=True)["bs"]
+bs = np.abs(np.concatenate(bs))
+
+result = {b: sT(l_max, b, r) for b in tqdm(bs)}
+
+with open(snakemake.output[0], "wb") as f:
+    pickle.dump(
+        result,
+        f,
+    )

### 2024-07-17 workflows/precision/scripts/plot_error.py

@@ -1,0 +1,216 @@
+import matplotlib.pyplot as plt
+from matplotlib import gridspec
+import numpy as np
+from jaxoplanet.experimental.starry.ylm import Ylm
+from jaxoplanet.experimental.starry.mpcore import mp, utils
+
+ref = "num"
+comp = snakemake.wildcards.code
+
+data = {
+    "small": {
+        ref: np.load(snakemake.input.ref[0], allow_pickle=True),
+        comp: np.load(snakemake.input.comp[0], allow_pickle=True),
+    },
+    "large": {
+        ref: np.load(snakemake.input.ref[1], allow_pickle=True),
+        comp: np.load(snakemake.input.comp[1], allow_pickle=True),
+    },
+}
+
+S = {
+    "small": {
+        ref: data["small"][ref]["s"],
+        comp: data["small"][comp]["s"],
+    },
+    "large": {
+        ref: data["large"][ref]["s"],
+        comp: data["large"][comp]["s"],
+    },
+}
+
+F = {
+    "small": {
+        ref: data["small"][ref]["f"],
+        comp: data["small"][comp]["f"],
+    },
+    "large": {
+        ref: data["large"][ref]["f"],
+        comp: data["large"][comp]["f"],
+    },
+}
+
+data_b = {
+    "small": np.load(snakemake.input.b[0], allow_pickle=True),
+    "large": np.load(snakemake.input.b[1], allow_pickle=True),
+}
+
+Bs = {"small": data_b["small"]["bs"], "large": data_b["large"]["bs"]}
+
+labels = {
+    "small": data_b["small"]["labels"],
+    "large": data_b["large"]["labels"],
+}
+
+r = {
+    "small": data_b["small"]["r"],
+    "large": data_b["large"]["r"],
+}
+
+
+def mp_diff(M_mp, M_np):
+    return utils.to_numpy(utils.to_mp(M_np) - M_mp)
+
+
+def errors(M_mp, M_np):
+    if isinstance(M_mp, mp.matrix):
+        d = mp_diff(M_mp, M_np)
+    else:
+        d = M_mp - M_np
+    rel = np.abs(d)
+    frac = np.abs(
+        d
[…]

### 2024-07-17 workflows/precision/scripts/plot_error_vs_degree.py

@@ -1,0 +1,118 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from jaxoplanet.experimental.starry.mpcore import mp, utils
+
+l_max = snakemake.params.l_max
+
+data = {
+    "small": {
+        "jax": np.load(snakemake.input.jax[0], allow_pickle=True)["f"],
+        "num": np.load(snakemake.input.num[0], allow_pickle=True)["f"],
+        "starry": np.load(snakemake.input.starry[0], allow_pickle=True)["f"],
+    },
+    "large": {
+        "jax": np.load(snakemake.input.jax[1], allow_pickle=True)["f"],
+        "num": np.load(snakemake.input.num[1], allow_pickle=True)["f"],
+        "starry": np.load(snakemake.input.starry[1], allow_pickle=True)["f"],
+    },
+}
+
+
+def mp_diff(M_mp, M_np):
+    return utils.to_numpy(utils.to_mp(M_np) - M_mp)
+
+
+def errors(M_mp, M_np):
+    if isinstance(M_mp, mp.matrix):
+        d = mp_diff(M_mp, M_np)
+    else:
+        d = M_mp - M_np
+    rel = np.abs(d)
+    frac = np.abs(
+        d
+        / max(1e-9, np.nanmax(np.min([np.abs(utils.to_numpy(M_mp)), np.abs(M_np)], 0)))
+    )
+    frac[rel < 1e-16] = 1e-16
+    return frac
+
+
+from jaxoplanet.experimental.starry.ylm import Ylm
+
+indices = {
+    l: np.array([Ylm.index(l, m) for m in range(-l, l + 1)]) for l in range(l_max + 1)
+}
+
+degree = np.arange(l_max + 1)
+err = {
+    "small": {
+        "jax": errors(data["small"]["num"], data["small"]["jax"]).max(0),
+        "starry": errors(data["small"]["num"], data["small"]["starry"]).max(0),
+    },
+    "large": {
+        "jax": errors(data["large"]["num"], data["large"]["jax"]).max(0),
+        "starry": errors(data["large"]["num"], data["large"]["starry"]).max(0),
+    },
+}
+
+
+def err_per_degree(err):
+    return [err[indices[l]].max() for l in range(l_max + 1)]
+
+
+s_color = "k"
+j_color = "C0"
+
+
+fig = plt.figure(figsize=(8, 3))
[…]

### 2024-07-17 workflows/precision/scripts/starry_fs.py

@@ -1,0 +1,41 @@
+import numpy as np
+import starry
+
+starry.config.lazy = False
+
+
+r = float(snakemake.wildcards.r)
+bs = np.load(snakemake.input[0], allow_pickle=True)["bs"]
+bs = np.abs(np.concatenate(bs))
+l_max = snakemake.params.l_max
+ys = snakemake.params.ys
+
+
+map = starry.Map(ydeg=l_max)
+
+
+S = []
+
+for b in bs:
+    try:
+        val = map.ops.sT([b], [r])[0]
+    except:
+        val = np.nan * np.ones(map.Ny)
+
+    S.append(val)
+
+S = np.array(S)
+
+
+flux = []
+from tqdm import tqdm
+
+for y in tqdm(ys):
+
+    map[:, :] = y
+    f = map.flux(xo=0.0, yo=bs, zo=10.0, ro=r)
+    flux.append(f)
+
+flux = np.array(flux).T
+
+np.savez(snakemake.output[0], f=flux, s=S)

### 2024-07-17 workflows/speed/scripts/flux_jaxoplanet.py

@@ -18,13 +18,8 @@ from jaxoplanet.experimental.starry.light_curves import *
 import timeit
 
 
-def timeit_f(strf, number=2, repeat=2):
-    times = (
-        np.array(
-            timeit.repeat(f"{strf}", repeat=repeat, number=number, globals=globals())
-        )[1:]
-        / number
-    )
+def timeit_f(strf, n=10):
+    times = np.array(timeit.repeat(f"{strf}", number=n, globals=globals()))[1:] / n
     return np.median(times)
 
 

### 2024-07-17 workflows/speed/scripts/flux_starry.py

@@ -4,7 +4,7 @@ import numpy as np
 import timeit
 
 
-def timeit_f(strf, number=2, repeat=1):
+def timeit_f(strf, number=4, repeat=2):
     times = (
         np.array(
             timeit.repeat(f"{strf}", repeat=repeat, number=number, globals=globals())

### 2024-07-17 workflows/speed/scripts/plot_error_rb.py

@@ -51,7 +51,7 @@ for i, order in enumerate(orders):
     else:
         ax.set_ylabel("b")
     ax.set_xlabel("r")
-    ax.set_title(f"$n$ = {order}")
+    ax.set_title(rf"$n$ = {order}")
     colorbar(pcm)
 
 # cm_ax = fig.add_axes([1.5, 0.1, 0.02, 0.8])

### 2024-07-17 workflows/speed/scripts/plot_error_vs_time.py

@@ -23,35 +23,51 @@ data = {
 
 
 def get_data(N):
-    precisions = np.array(
-        [
-            np.abs(
-                data["jaxoplanet"][N][order]["value"] - data["starry"][N]["value"]
-            ).max()
-            for order in orders
-        ]
-    )
+    relative_errors = []
+
+    for order in orders:
+        error = np.abs(
+            data["jaxoplanet"][N][order]["value"] - data["starry"][N]["value"]
+        )
+        norm = np.abs(data["jaxoplanet"][N][order]["value"])
+        norm[norm < 1e-9] = 1e-9
+        error /= norm
+        relative_errors.append(error.max())
+
+    relative_errors = np.array(relative_errors)
 
     times = np.array([data["jaxoplanet"][N][order]["time"] for order in orders])
 
-    reference_precision = np.min(precisions)
+    reference_error = 1e-15
+    norm = np.abs(data["jaxoplanet"][N][order]["value"])
+    norm[norm < 1e-9] = 1e-9
+    reference_relative_error = reference_error / norm
+
     reference_time = data["starry"][N]["time"]
-    return precisions, times, reference_precision, reference_time
+    return relative_errors, times, reference_relative_error.max(), reference_time
 
 
 plt.figure(figsize=(5, 3.5))
 
-plt.subplot(111)
+ax = plt.subplot(111)
 precisions, times, reference_precision, reference_time = get_data(1000)
-plt.plot(preci sions, times * 1e6, ".-", c="k", label="jaxoplanet (CPU)")
+print(precisions.shape)
+plt.plot(precisions, times * 1e6, ".-", c="k", label="jaxoplanet (CPU)")
 plt.plot(reference_precision, reference_time * 1e6, "+", label="starry (CPU)")
 plt.axhline(reference_time * 1e6, c="C0", ls="-", alpha=0.2)
 plt.axvline(reference_precision, c="C0", ls="-", alpha=0.2)
 plt.xscale("log")
-plt.xlim(1e-7, 0.5e-15)
+# plt.yscale("log")
+xlim = plt.xlim()
+plt.xlim(xlim[1], xlim[0])
+# plt.xlim(1e-7, 0.5e-15)
 plt.ylim(0)
 plt.xlabel("relative error")
 plt.ylabel("computation time (µs)")
+y_labels = ax.get_yticks()
+import matplotlib.ticker as ticker
+
[…]

### 2024-07-17 workflows/speed/scripts/plot_points_vs_time.py

@@ -3,6 +3,7 @@ import numpy as np
 
 
 Ns = snakemake.params.Ns
+
 data = {
     "starry": [np.load(f)["time"] for f in snakemake.input["starry"]],
     "jaxoplanet": [np.load(f)["time"] for f in snakemake.input["jaxoplanet"]],

### 2024-07-18 latex/paper.tex
add SAR error computation+ larger b range s and f errors

@@
 \documentclass[modern]{aastex631}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{hyperref}
-\usepackage{bm}
-\usepackage{listings}
-\usepackage{graphicx}
-\usepackage{float}
-\usepackage{soul}
-\usepackage{mathtools}
 
-% commands
-\newcommand{\todo}[1]{\colorbox{red}{\textcolor{white}{\texttt{TO DO}}}\,}
-\newcommand{\set}[1]{\{\,#1\,\}}
-\newcommand{\footlink}[1]{\footnote{\url{#1}}}
-\DeclareMathOperator*{\argmax}{arg\,max}
-\DeclareMathOperator*{\argmin}{arg\,min}
-\let\subsectionautorefname\sectionautorefname
-\let\subsubsectionautorefname\sectionautorefname
-
-% no indent
-\setlength\parindent{0pt}
-
-% code style
-\lstdefinestyle{mystyle}{
-    backgroundcolor=\color{white},
-    commentstyle=\color{gray!70},
-    keywordstyle=\color{Bittersweet},
-    stringstyle=\color{RoyalBlue},
-    basicstyle=\fontsize{8.5}{13}\fontfamily{DejaVuSansMono-TLF}\selectfont,
-    breakatwhitespace=false,         
-    breaklines=true,
-    rulecolor=\color{black!15},
-    numbers=none,
-    numberstyle=\fontsize{7}{11}\fontfamily{DejaVuSansMono-TLF}\selectfont\color{gray!50},
-    framerule=0pt,
-    breakindent=5pt,
-    resetmargins=true,
-    numbersep=10pt,
-    frame=single,
-    aboveskip=1em,
-    belowskip=1em,
-    xleftmargin=6pt,
-    framexleftmargin=4pt
-}
-\lstset{style=mystyle}
+\input{stylez}
 
@@
 Through this implementation, we provide the community with a differentiable model of stellar light curves, and enable the use of powerful machine-learning tools available in the \texttt{jax} ecosystem.
+\gitlink{}
 
@@
 \begin{equation}\label{eq:starry}
-    f = \bm{s^TARR'y}
+    f = \bvec{s}^{\boldsymbol{\mathsf{T}}} \bvec{A} \, \bvec{R'} \, \bvec{R} \, \bvec{y}
 \end{equation}
@@
 \begin{equation}
-    \bm{v} = \frac{1}{\sqrt{1 - \cos^2{\left(\frac{inc}{2} \right)} \cos^2{\left(\frac{obl}{2}\right)}}} \begin{pmatrix}
+    \bvec{v} = \frac{1}{\sqrt{1 - \cos^2{\left(\frac{inc}{2} \right)} \cos^2{\left(\frac{obl}{2}\right)}}} \begin{pmatrix}
         \sin{\left(\frac{inc}{2} \right)} \cos{\left(\frac{obl}{2}\right)}\\
@@
 Although \textit{starry} occultation light curves have analytical expressions, their practical implementation relies on numerical approximations chosen to balance precision and computation time.
-This is the case, for example, of the solution vector $\bm{s}$, computed using numerical series expansions in \texttt{starry} (see \todo{} in \citealt{starry}) and numerical integration in \texttt{jaxoplanet} (see \autoref{solution_vectors}).
+This is the case, for example, of the solution vector $\bvec{s}$, computed using numerical series expansions in \texttt{starry} (\citealt[section D.2.3]{starry}) and numerical integration in \texttt{jaxoplanet} (see \autoref{solution_vectors}).
 Other expressions, such as the Wigner-D matrices used in the rotation of the spherical harmonics basis, are computed using reccurence relations prone to the accumulation of numerical errors and instability.\\\\
@@
 We evaluate the precision of the \texttt{starry} and \texttt{jaxoplanet} terms against quantities computed at arbitrary precision, using a ground truth version of the code implemented with the \texttt{mpmath} Python library\footlink{https://mpmath.org/}.
-The precision of the overall flux $f$ can only be understood by considering the precision of the terms involved in \autoref{eq:starry}: the solution vector $\bm{s}$ ($\bm{r}$ if no occultation), the basis matrices $\bm{A}$ and $\bm{B}$, and the Wigner-D rotation tensor $\bm{R}$.
-Although we show in \todo{add an appendix showing errors for these quantities} that the precision of the \texttt{starry} and \texttt{jaxoplanet} implentations differ for many of these terms, we focus this section on the solution vector $\bm{s}$ and the overall flux $f$.\\\\
-Assuming that we perform the numerical integration of the solution vector $\bm{s}$ at a relatively high order ($n=500$; see \autoref{solution_vectors}), \autoref{fig:precision_s} shows that our reimplentation of \texttt{starry} reaches relative errors lower than 1 part per billion for $\bm{s}$ and the overall flux $f$, considering both a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$.
+The precision of the overall flux $f$ can only be understood by considering the precision of the terms involved in \autoref{eq:starry}: the solution vector $\bvec{s}$ ($\bvec{r}$ if no occultation), the basis matrices $\bvec{A}$ and $\bvec{B}$, and the Wigner-D rotation tensor $\bvec{R}$.
+Although we show in \autoref{fig:precision_SAR} that the precision of the \texttt{starry} and \texttt{jaxoplanet} implentations differ for many of these terms, we focus this section on the solution vector $\bvec{s}$ and the overall flux $f$.\\\\
+Assuming that we perform the numerical integration of the solution vector $\bvec{s}$ at a relatively high order ($n=500$; see \autoref{solution_vectors}), \autoref{fig:precision_s} shows that our reimplentation of \texttt{starry} reaches relative errors lower than 1 part per billion for $\bvec{s}$ and the overall flux $f$, considering both a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$.
 For reference, \autoref{fig:precision_s_starry} shows comparable errors on the same quantities computed with the C++ implentation of \texttt{starry}.
 In addition, \autoref{fig:precision_degree} shows the evolution of the error on $f$ for increasing degrees of the spherical harmonics, both for \texttt{starry} and \texttt{jaxoplanet}.
-With these results, we validate the precision of \texttt{jaxoplanet} ans show it is on per with the legacy C++ \texttt{starry} implentation\footnote{version \todo{add version number of starry}} described in \cite{starry}.\\\\
+With these results, we validate the precision of \texttt{jaxoplanet} ans show it is on per with the legacy C++ \texttt{starry} implentation\footnote{version 1.2.0 \citep{starry_120}} described in \cite{starry}.\\\\
 \begin{figure}[H]
@@
         \includegraphics[width=\textwidth]{../workflows/precision/figures/error_jax.pdf}
-        \caption{}
+        \caption{Error of the solution vector $\bvec{s}$ and the overall flux $f$ for all terms of the spherical harmonics basis up to degree $l=20$.
+Errors are computed against arbitrary-precision computations of $\bvec{s}$ and $f$ for a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$.
+For each $(l, m)$ basis component, errors are scaled to the highest values of $\bvec{s}$ (respectively $f$) over $b$.
+Then, the plotted errors for each degree $l$ are the maximum of the scaled errors over all $m\in [-l, l]$ components over $b$.
+The disks at the top of the figure show the transit configurations of the occultor (light dark) and the star (light gray) for different values of $b$.
+\codelink{workflows/precision/scripts/plot_error}}
         \label{fig:precision_s}
@@
         \includegraphics[width=\textwidth]{../workflows/precision/figures/error_degree.pdf}
-        \caption{}
+        \caption{Maximum errors in the overall flux $f$ for all degrees of the spherical harmonics components.
+Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s}, maxima taken over the range of values $b$.
+\codelink{workflows/precision/scripts/plot_error_vs_degree}}
         \label{fig:precision_degree}
@@
 \end{figure}
-As described in \autoref{solution_vectors}, we compute \textit{starry} light curves using the Gauss Legendre quadrature to approximate the $\mathcal{P}$ integral involved in the expression of the solution vector $\bm{s}$.
+As described in \autoref{solution_vectors}, we compute \textit{starry} light curves using the Gauss Legendre quadrature to approximate the $\mathcal{P}$ integral involved in the expression of the solution vector $\bvec{s}$.
 Hence, the precision of the computed flux $f$ depends on the order of the Legendre polynomial which defines the number of points used to approximate $\mathcal{P}$.
@@
 The error is evaluated for different orders $n$ of the Gauss-Legendre quadrature and computed against \texttt{starry} for speed and convenience.
-Note that \autoref{fig:precision_s} was produced using $n=500$ (right plot) for which $\bm{s}$ reaches machine precision.}
+Note that \autoref{fig:precision_s} was produced using $n=500$ (right plot) for which $\bvec{s}$ reaches machine precision.}
         \label{fig:relative_error_1}
@@
 
-\section{Performance}
+\section{Error budget}
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[width=\textwidth]{../workflows/precision/figures/error_SAR.pdf}
+        \caption{\codelink{workflows/precision/scripts/plot_error_SAR}}
+        \label{fig:precision_SAR}
+    \end{center}
+\end{figure}
+
 \begin{figure}[H]
@@
         \includegraphics[width=\textwidth]{../workflows/precision/figures/error_starry.pdf}
-        \caption{}
+        \caption{Same as \autoref{fig:precision_s} but with $bvec{s}$ and $f$ computed with the C++ implementation of \texttt{starry}. \codelink{workflows/precision/scripts/plot_error}}
         \label{fig:precision_s_starry}

### 2024-07-18 workflows/precision/scripts/SAR_jax_starry.py

@@ -1,0 +1,102 @@
+import jax
+
+jax.config.update("jax_enable_x64", True)
+
+
+import starry
+import numpy as np
+
+starry.config.lazy = False
+
+l_max = snakemake.params.l_max
+
+# R
+
+u = (1.0, 0.0, 0.0)
+theta = 0.1
+
+
+from jaxoplanet.experimental.starry.mpcore.rotation import R
+
+num_R = R(l_max, u, theta)
+
+
+def R_starry(lmax, u, theta):
+    shapes = np.array([2 * l + 1 for l in range(l_max + 1)])
+
+    R = []
+    ms = starry.Map(lmax)
+    y0 = np.zeros((l_max + 1) ** 2)
+
+    for i in range(len(y0)):
+        y = y0.copy()
+        y[i] = 1
+        R.append(ms.ops.dotR(y[None, :], *u, theta)[0])
+
+    R = np.array(R)
+
+    R_blocks = []
+
+    n = 0
+    for j in range(0, len(shapes)):
+        i = shapes[j]
+        a, b = n, n + i
+        n += i
+        R_blocks.append(R[a:b][:, a:b])
+
+    return R_blocks
+
+
+sta_R = R_starry(l_max, u, theta)
+
+from jaxoplanet.experimental.starry.s2fft_rotation import compute_rotation_matrices
+
+jax_R = compute_rotation_matrices(l_max, *u, theta)
+
+# A
+from jaxoplanet.experimental.starry.basis import A1, A2_inv
+import jax.numpy as jnp
+
+
+jax_A1 = A1(l_max).todense()
+jax_A2 = jnp.linalg.inv(A2_inv(l_max).todense())
+jax_A = jax_A2 @ jax_A1
+
+ms = starry.Map(l_max)
+sta_A1 = ms.ops.A1.eval().todense()
+sta_A = ms.ops.A.eval().todense()
+
+# S
+
+from jaxoplanet.experimental.starry.solution import solution_vector
+from jaxoplanet.experimental.starry.mpcore.solution import sT
+
+r = 1.0
+b = 1.0
+
+num_sT = sT(l_max, b, r)
+jax_sT = solution_vector(l_max, 500)(b, r)
+sta_sT = ms.ops.sT([b], [r])[0]
+
+import pickle
+
+results = {
+    "u": u,
+    "theta": theta,
+    "sta_R": sta_R,
+    "jax_R": jax_R,
[…]

### 2024-07-18 workflows/precision/scripts/make_b.py

@@ -1,7 +1,7 @@
 import numpy as np
 
 
-def b_range(r, logdelta=-3, logeps=-6, res=1):
+def b_range(r, logdelta=-3, logeps=-6, res=50):
     delta = 10**logdelta
     eps = 10**logeps
     if r > 1:

### 2024-07-18 workflows/precision/scripts/plot_error.py

@@ -1,3 +1,8 @@
+# Plot the errors of s, f on each of the (l,m) spherical harmonic components
+# The scaled error is computed in the `errors` function and the error for a
+# given degree l is the maximum error across all m \in [-l, l] components of that degree
+# (indices are for each degree are stored in the `indices` dictionary)
+
 import matplotlib.pyplot as plt
 from matplotlib import gridspec
 import numpy as np

### 2024-07-18 workflows/precision/scripts/plot_error_SAR.py

@@ -1,0 +1,98 @@
+import jax
+
+jax.config.update("jax_enable_x64", True)
+
+import numpy as np
+
+num_data = np.load(snakemake.input[0], allow_pickle=True)
+num_A = num_data["A2"] @ num_data["A1"]
+
+other_data = np.load(snakemake.input[1], allow_pickle=True)
+jax_A = other_data["jax_A"]
+sta_A = other_data["sta_A"]
+
+jax_sT = other_data["jax_sT"]
+sta_sT = other_data["sta_sT"]
+num_sT = other_data["num_sT"]
+
+jax_R = other_data["jax_R"]
+sta_R = other_data["sta_R"]
+num_R = other_data["num_R"]
+
+b = other_data["b"]
+r = other_data["r"]
+
+u = other_data["u"]
+theta = other_data["theta"]
+
+from jaxoplanet.experimental.starry.mpcore.utils import to_numpy, to_mp
+
+
+def diff(M1, M2, max_diff=1e-16):
+    if isinstance(M1, np.ndarray) and isinstance(M2, np.ndarray):
+        d = M1 - M2
+    else:
+        if (M1, np.ndarray):
+            _M1 = to_mp(M1)
+        if (M2, np.ndarray):
+            _M2 = to_mp(M2)
+        d = to_numpy(_M1 - _M2)
+
+    # d[d < max_diff] = max_diff
+    return d
+
+
+def diff_R(M1, M2):
+    return np.hstack([np.abs(diff(a, b)).max(0) for a, b in zip(M1, M2)])
+
+
+import matplotlib.pyplot as plt
+
+starry_color = "0.6"
+jax_color = "C0"
+
+fig, subplot = plt.subplots(1, 3, figsize=(8.0, 3), sharey=True, sharex=True)
+ax = subplot[0]
+ax.plot(np.abs(diff(num_sT, jax_sT)).T, label="jaxoplanet", color=jax_color)
+ax.plot(np.abs(diff(num_sT, sta_sT)).T, label="starry", color=starry_color)
+ax.set_title(r"$s^{T}$")
+ax.set_ylabel("Error")
+# annotation in the corner
+ax.annotate(
+    f"$b={b}$\n$r={r}$",
+    xy=(0.05, 0.95),
+    xycoords="axes fraction",
+    textcoords="axes fraction",
+    color="black",
+    va="top",
+)
+ax.legend()
+
[…]

### 2024-07-18 workflows/precision/scripts/plot_error_vs_degree.py

@@ -67,7 +67,9 @@ fig = plt.figure(figsize=(8, 3))
 x = np.arange(0, l_max + 2, 2)
 
 ax1 = plt.subplot(121)
-ax1.plot(degree, err_per_degree(err["small"]["jax"]), ".-", label="jax", c=j_color)
+ax1.plot(
+    degree, err_per_degree(err["small"]["jax"]), ".-", label="jaxoplanet", c=j_color
+)
 ax1.plot(
     degree, err_per_degree(err["small"]["starry"]), ".-", label="starry", c=s_color
 )
@@ -79,7 +81,9 @@ ax1.set_ylim(1e-17, 1e-5)
 ax1.set_xticks(x)
 
 ax2 = plt.subplot(122)
-ax2.plot(degree, err_per_degree(err["large"]["jax"]), ".-", label="jax", c=j_color)
+ax2.plot(
+    degree, err_per_degree(err["large"]["jax"]), ".-", label="jaxoplanet", c=j_color
+)
 ax2.plot(
     degree, err_per_degree(err["large"]["starry"]), ".-", label="starry", c=s_color
 )

### 2024-07-18 workflows/utils/scripts/starry_version.py

@@ -1,0 +1,6 @@
+import starry
+
+starry_version = starry.__version__
+
+with open(snakemake.output[0], "w") as f:
+    f.write(starry_version)

### 2024-08-09 latex/paper.tex
feat: add all error benchmark tex and scripts

@@
 Errors are computed against arbitrary-precision computations of $\bvec{s}$ and $f$ for a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$.
-For each $(l, m)$ basis component, errors are scaled to the highest values of $\bvec{s}$ (respectively $f$) over $b$.
+For each $(l, m)$ basis component all coefficients are set to zero except $y_{l,m}=1$ and errors are scaled to the highest values of $\bvec{s}$ (respectively $f$) over $b$.
 Then, the plotted errors for each degree $l$ are the maximum of the scaled errors over all $m\in [-l, l]$ components over $b$.
@@
 \end{figure}
+For reference, \autoref{fig:precision_limbdark} shows the relative error in the flux of a limb-darkened star occulted by an opaque body for increasing orders of a polynomial limb-darkening law.
+In this figure, we also report the error of the flux computed using the \textsf{exoplanet} Python package (limited to linear and quadratic limb-darkening).
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[width=\textwidth]{../workflows/precision/figures/limbdark_error.PDF}
+        \caption{Maximum errors in the overall flux $f$ of a star occulted by an opaque body for increasing orders of its polynomial limb-darkening law.
+Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s}, taken over the range of values $b$.
+For each order $n$ all limb-darkening coefficients are set to zero except $u_n=1$.
+\codelink{workflows/precision/scripts/plot_error_limbdark}}
+        \label{fig:precision_limbdark}
+    \end{center}
+\end{figure}
+A set of unitary polynomial limb-darkening coefficients of order 20 leads to spherical harmonics coefficients as large as $10^6$.
+This explains the larger errors observed in \autoref{fig:precision_limbdark} (error versus limb-darkening order) compared to the ones shown in \autoref{fig:precision_degree} (error versus degree of the spherical harmonics).
+We not that the errors scale with the values of $u$.\\\\
 As described in \autoref{solution_vectors}, we compute \textit{starry} light curves using the Gauss Legendre quadrature to approximate the $\mathcal{P}$ integral involved in the expression of the solution vector $\bvec{s}$.
@@
     \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/speed/figures/error_rb.png}
-        \caption{relative numerical error in the flux $f$ of a linearly limb-darkened star transited by an opaque companion over the $(r, b)$ parameter space.
+        \includegraphics[width=\textwidth]{../workflows/precision/figures/br_error.png}
+        \caption{relative numerical error in the flux $f$ of a linearly limb-darkened star (degree $l$=1) transited by an opaque companion over the $(r, b)$ parameter space.
 The error is evaluated for different orders $n$ of the Gauss-Legendre quadrature and computed against \texttt{starry} for speed and convenience.
-Note that \autoref{fig:precision_s} was produced using $n=500$ (right plot) for which $\bvec{s}$ reaches machine precision.}
+Note that \autoref{fig:precision_s} was produced using $n=500$ (right plot) for which $\bvec{s}$ reaches machine precision.
+\codelink{workflows/precision/scripts/plot_br_error}}
         \label{fig:relative_error_1}
@@
 \end{figure}
-At a fixed occultor radius, \autoref{fig:relative_error_2} shows the time required to compute the flux of the system for a range of $N$ impact parameters $b$, plotted against the maximum relative error reached, the order $n$ of the Gauss-Legendre approximation, and the hardware used for the computation.
+In \autoref{fig:error_order_gausslegendre}, we show how the error in the solution vector evolves with the order $n$ of the Gauss-Legendre quadrature, and its impact on the error in the overall flux.
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[width=\textwidth]{../workflows/precision/figures/error_order.pdf}
+        \caption{Maximum errors in the solution vector and overall flux $f$ for different occultor radii and orders $n$ of the Gauss-Legendre quadrature used to compute the P integral (see \autoref{solution_vectors}).
+Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s}, maxima taken over the range of values $b$.
+As in \autoref{fig:precision_s} the maximum degree of the spherical harmonics basis is set to $l=20$.}
+        \label{fig:error_order_gausslegendre}
+    \end{center}
+\end{figure}
+The minimum error is always achieved for higher values of $n$, which comes with higher computational cost (see \autoref{speed}).
+Hence users must carefully set the order parameter to balance precision and computational speed.
+As the minimum error reached also depends on the degree of the spherical harmonics basis used, \autoref{fig:error_order_gausslegendre_degree} shows the errors on the flux for each degree of the spherical harmonics basis, for an occultor with relative radii $r=0.01$, $r=1$ and $r=100$.
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[width=\textwidth]{../workflows/precision/figures/error_order_degree.pdf}
+        \caption{Maximum errors in the overall flux $f$ for all degrees of the spherical harmonics components for different occultor radii and orders $n$ of the Gauss-Legendre quadrature used to compute the P integral (see \autoref{solution_vectors}).
+Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s}, maxima taken over the range of values $b$.}
+        \label{fig:error_order_gausslegendre_degree}
+    \end{center}
+\end{figure}
+\subsection{Speed}\label{speed}
 
-% \begin{figure}[H]
-%     \begin{center}
-%         \includegraphics[width=\textwidth]{../workflows/performance/figures/error_vs_time.pdf}
-%         \caption{}
-%         \label{fig:relative_error_2}
-%     \end{center}
-% \end{figure}
-These computations were run on an apple M1 CPU and an NVIDIA ... GPU. Thanks to the use of JAX, no harware-specific... \\\\
+The main advantage of the \texttt{jax} library is its ability to perform hardware-accelerated computations on CPUs and GPUs.
+In this section, we evaluate the speed of the \texttt{jaxoplanet} implementation and compare it with the C++ implementation of \texttt{starry}.
+For quadratically limb-darkened light curve, we also compare our implementation with the one provided by the \textsf{exoplanet} Python package.\\\\
+In \autoref{fig:speed_order_gausslegendre}, we show the evolution of the computing time required to compute an occultation light-curve for a limb-darkened star and a non-uniform star described by spherical harmonics up to order $l=20$.
+This figure include the value of the order $n$ of the Gauss-Legendre quadrature required to reach a given precision, and illustrate well the trade-off between precision and speed that users must consider when setting the order parameter.\\\\
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[width=\textwidth]{../workflows/speed/figures/speed_vs_order.pdf}
+        \caption{Computation time of the occultation light curve of a quadts}
+        \label{fig:speed_order_gausslegendre}
+    \end{center}
+\end{figure}
 
-By computing $\mathcal{P}$ numerically, we show that our vectorized implementation reaches machine-precision on a GPU, up to an order of magnitude faster than the current C++ implementation of \texttt{starry} ran on a CPU.\\\\
-
-Plot of the precision versus speed of the computed light curves.
-
-Plot of the solution vectors computation similar to figure 11 and 12 of luger (maybe in the appendix)
-
-\subsection{Speed}\label{speed}
+\todo{} shows
 

### 2024-08-09 workflows/precision/scripts/SAR_jax_starry.py

@@ -90,6 +90,8 @@ results = {
     "r": r,
     "sta_A": sta_A,
     "jax_A": jax_A,
+    "jax_A1": jax_A1,
+    "jax_A2": jax_A2,
     "sta_sT": sta_sT,
     "jax_sT": jax_sT,
     "num_sT": num_sT,

### 2024-08-09 workflows/precision/scripts/concatenate_errors_GL.py

@@ -1,0 +1,59 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from jaxoplanet.experimental.starry.mpcore import utils, mp
+from tqdm import tqdm
+
+rs = snakemake.params.rs
+orders = snakemake.params.orders
+
+all_errors_s = []
+all_errors_f = []
+
+
+def errors(M_mp, M_np):
+    if isinstance(M_mp, mp.matrix):
+        d = utils.diff_mp(M_mp, M_np)
+    else:
+        d = M_mp - M_np
+
+    rel = np.abs(d)
+    frac = np.abs(
+        d
+        / max(1e-9, np.nanmax(np.min([np.abs(utils.to_numpy(M_mp)), np.abs(M_np)], 0)))
+    )
+    frac[rel < 1e-16] = 1e-16
+    return frac
+
+
+for i in tqdm(range(len(rs))):
+    num_data = np.load(snakemake.input.num[i], allow_pickle=True)
+    all_errors_f.append(
+        [
+            np.abs(
+                errors(
+                    num_data["f"],
+                    np.load(snakemake.input[f"jax_{order}"][i], allow_pickle=True)["f"],
+                )
+            )
+            for order in orders
+        ]
+    )
+    all_errors_s.append(
+        [
+            np.abs(
+                errors(
+                    num_data["s"],
+                    np.load(snakemake.input[f"jax_{order}"][i], allow_pickle=True)["s"],
+                )
+            )
+            for order in orders
+        ]
+    )
+
+np.savez(
+    snakemake.output[0],
+    errors_s=all_errors_s,
+    errors_f=all_errors_f,
+    orders=orders,
+    rs=rs,
+)

### 2024-08-09 workflows/precision/scripts/exoplanet_f_limbdark.py

@@ -1,0 +1,18 @@
+from exoplanet.light_curves import LimbDarkLightCurve
+import numpy as np
+from tqdm import tqdm
+
+r = float(snakemake.wildcards.r)
+degree = snakemake.params.degree
+bs = np.load(snakemake.input[0], allow_pickle=True)["bs"]
+b = np.abs(np.concatenate(bs))
+
+
+def flux(u):
+    return (
+        LimbDarkLightCurve(*u)._compute_light_curve(b, np.ones_like(b) * r).eval()
+    ) + 1.0
+
+
+fluxes = np.array([flux(u) for u in [(0.0, 0.0), (1.0, 0.0), (0.0, 1.0)]])
+np.savez(snakemake.output[0], f=fluxes)

### 2024-08-09 workflows/precision/scripts/jax_br_flux.py

@@ -1,0 +1,34 @@
+import jax
+from tqdm import tqdm
+
+jax.config.update("jax_platform_name", "cpu")
+jax.config.update("jax_enable_x64", True)
+
+import numpy as np
+from jaxoplanet.experimental.starry import Surface
+from jaxoplanet.experimental.starry.light_curves import surface_light_curve
+
+b, r = np.load(snakemake.input[0]).values()
+order = int(snakemake.wildcards.order)
+
+surface = Surface(period=None, u=(1,))
+
+f = jax.jit(
+    jax.vmap(surface_light_curve, in_axes=(None, None, None, 0, None, None, None)),
+    static_argnames=("order",),
+)
+
+result = np.zeros((len(r), len(b)))
+
+
+def flux(order):
+    for i, _r in enumerate(tqdm(r)):
+        calc = f(surface, _r, 0.0, b, 10.0, 0.0, order)
+        result[i, :] = calc
+
+    return result
+
+
+calc = flux(order)
+
+np.savez(snakemake.output[0], flux=calc)

### 2024-08-09 workflows/precision/scripts/jax_f_limbdark.py

@@ -1,0 +1,35 @@
+import jax
+
+jax.config.update("jax_enable_x64", True)
+
+from jaxoplanet.experimental.starry import Surface
+from jaxoplanet.experimental.starry.light_curves import surface_light_curve
+import numpy as np
+from tqdm import tqdm
+
+
+r = float(snakemake.wildcards.r)
+degree = snakemake.params.degree
+bs = np.load(snakemake.input[0], allow_pickle=True)["bs"]
+b = np.abs(np.concatenate(bs))
+order = snakemake.params.order
+
+
+def u(deg):
+    u = np.zeros(degree)
+    if degree == 0:
+        return u
+    else:
+        u[deg - 1] = 1
+        return u
+
+
+function = jax.vmap(
+    lambda deg, b: surface_light_curve(
+        Surface(u=u(deg)), y=b, z=10.0, r=r, order=order
+    ),
+    (None, 0),
+)
+
+fluxes = np.array([function(deg, b) for deg in tqdm(range(degree + 1))])
+np.savez(snakemake.output[0], f=fluxes)

### 2024-08-09 workflows/precision/scripts/jax_fs_GL.py

@@ -1,3 +1,4 @@
+import numpy as np
 import jax
 
 jax.config.update("jax_enable_x64", True)
@@ -6,8 +7,9 @@ from jaxoplanet.experimental.starry.solution import solution_vector
 from jaxoplanet.experimental.starry.light_curves import *
 import numpy as np
 import jax.numpy as jnp
-from jaxoplanet.experimental.starry.mpcore.solution import rT
-from jaxoplanet.experimental.starry.mpcore.basis import A1, A2_inv
+from jaxoplanet.experimental.starry.solution import rT
+from jaxoplanet.experimental.starry.mpcore.basis import A1
+from jaxoplanet.experimental.starry.basis import A2_inv
 from jaxoplanet.experimental.starry.mpcore.utils import to_numpy
 
 
@@ -23,7 +25,7 @@ def surface_light_curve(
     theta: float = 0.0,
     order: int = 20,
 ):
-    rT_deg = to_numpy(rT(deg))
+    rT_deg = rT(deg)
 
     # occulting body
     if True:
@@ -34,7 +36,8 @@ def surface_light_curve(
         sT = solution_vector(deg, order=order)(b, r)
 
         if deg > 0:
-            A2 = to_numpy(A2_inv(deg) ** -1)
+            A2 = scipy.sparse.linalg.inv(A2_inv(deg))
+            A2 = jax.experimental.sparse.BCOO.from_scipy_sparse(A2)
         else:
             A2 = jnp.array([1])
 
@@ -61,30 +64,35 @@ def jax_flux(deg, order=20, inc=np.pi / 2, obl=0.0):
     return jax.jit(impl)
 
 
+data = np.load(snakemake.input[0])
+b = np.hstack(data["bs"])
+orders = data["orders"]
+ys = snakemake.params.ys
 r = float(snakemake.wildcards.r)
 l_max = snakemake.params.l_max
-bs = np.load(snakemake.input[0], allow_pickle=True)["bs"]
-bs = np.abs(np.concatenate(bs))
-ys = snakemake.params.ys
-order = int(snakemake.wildcards.order)
 
-jax_s_function = jax.jit(
-    jax.vmap(jax.jit(solution_vector(l_max, order)), in_axes=(0, None))
-)
-
-jax_f_function = jax.jit(
-    jax.vmap(jax.jit(jax_flux(l_max, order=order)), in_axes=(None, 0, None, None))
[…]

### 2024-08-09 workflows/precision/scripts/make_b_orders_GL.py

@@ -1,0 +1,12 @@
+import numpy as np
+
+r = float(snakemake.wildcards.r)
+
+if r > 1:
+    b = np.linspace(r - 1 - 1e-6, r, 10)
+else:
+    b = np.linspace(1e-6, 1, 10)
+
+orders = np.array([20, 50, 100, 150, 200, 250, 300, 350, 400, 450, 500]).astype(int)
+
+np.savez(snakemake.output[0], bs=[b], orders=orders)

### 2024-08-09 workflows/precision/scripts/make_br_dense.py

@@ -1,0 +1,8 @@
+import numpy as np
+
+N = int(snakemake.wildcards.N)
+
+r = np.linspace(0, 2, N)
+b = np.linspace(0, 2, N)
+
+np.savez(snakemake.output[0], r=r, b=b)

### 2024-08-09 workflows/precision/scripts/num_f_limbdark.py

@@ -1,0 +1,64 @@
+import numpy as np
+import pickle
+from jaxoplanet.experimental.starry.mpcore.utils import mp
+from jaxoplanet.experimental.starry import basis
+from jaxoplanet.experimental.starry.mpcore import solution
+from tqdm import tqdm
+
+# some params
+r = snakemake.wildcards.r
+degree = snakemake.params.degree
+
+# load the fs for each y such as
+# y0 = (1, 0, 0, 0, 0, ...)
+# y1 = (1, 1, 0, 0, 0, ...)
+# y2 = (1, 1, 1, 0, 0, ...)
+# etc.
+fs = np.load(snakemake.input.fs, allow_pickle=True)["f"].T
+
+fs = mp.matrix(
+    [
+        fs[0, :].tolist()[0],
+        *[(fs[i, :] - fs[0, :]).tolist()[0] for i in range(1, len(fs))],
+    ]
+)
+
+# load b
+bs = np.load(snakemake.input.b, allow_pickle=True)["bs"]
+b = np.hstack(bs)
+
+# load A1
+print("inverting A1")
+num_matrices = np.load(snakemake.input.matrices, allow_pickle=True)
+
+# TODO: transfer that into the saved matrices
+# not done cause everything will rerun...
+A1inv = num_matrices["A1"] ** -1
+
+
+# the function to compute the flux
+def num_flux(u):
+    deg = len(u)
+    n = (deg + 1) ** 2
+    _u = mp.matrix([[1, *u]])
+    U = mp.matrix(basis.U(deg))
+    pu = _u @ U
+    yu = A1inv[0:n, 0:n] @ pu.T
+    norm = (pu @ solution.rT(deg))[0]
+    yufs = mp.matrix([(yu[i] * fs[i, :]).tolist()[0] for i in range((deg + 1) ** 2)])
+    return mp.matrix([sum(yufs[:, i]) for i in range(len(yufs.T))]) / norm
+
+
+def u(deg):
+    u = np.zeros(degree)
+    if degree == 0:
+        return u
+    else:
+        u[deg - 1] = 1
+        return u
+
+
+print("running num_flux")
+fluxes = mp.matrix([num_flux(u(deg)).T.tolist()[0] for deg in tqdm(range(degree + 1))])
+with open(snakemake.output[0], "wb") as f:
+    pickle.dump({"f": fluxes}, f)

### 2024-08-09 workflows/precision/scripts/num_fs_GL.py

@@ -1,0 +1,79 @@
+import numpy as np
+
+r = float(snakemake.wildcards.r)
+l_max = snakemake.params.l_max
+data = np.load(snakemake.input[0])
+b = data["b"]
+y = snakemake.params.y
+
+num_matrices = np.load(snakemake.input[1], allow_pickle=True)
+ST = np.load(snakemake.input[2], allow_pickle=True)
+
+from jaxoplanet.experimental.starry.mpcore import mp
+from jaxoplanet.experimental.starry.mpcore.rotation import (
+    dot_rotation_matrix,
+    dot_rz,
+)
+
+y = mp.matrix(y.tolist())
+
+
+def flux_function(l_max, inc, obl):
+    _A1 = num_matrices["A1"]
+    _A2 = num_matrices["A2"]
+
+    R_px = num_matrices["R_px"]
+    R_mx = num_matrices["R_mx"]
+    R_obl = num_matrices["R_obl"]
+    R_inc = num_matrices["R_inc"]
+
+    def rotate_y(y, phi):
+        y_rotated = dot_rotation_matrix(l_max, None, None, R_px)(y)
+        y_rotated = dot_rz(l_max, phi)(y_rotated)
+        y_rotated = dot_rotation_matrix(l_max, None, None, R_mx)(y_rotated)
+        y_rotated = dot_rotation_matrix(l_max, None, None, R_obl)(y_rotated)
+        y_rotated = dot_rotation_matrix(l_max, None, None, R_inc)(y_rotated)
+        return y_rotated
+
+    def occ_flux(y, b, r, phi):
+        xo = 0.0
+        yo = b
+        theta_z = mp.atan2(xo, yo)
+        _sT = ST[b]
+        x = _sT.T @ _A2
+        y_rotated = rotate_y(y, phi)
+        y_rotated = dot_rz(l_max, theta_z)(y_rotated)
+        py = (_A1 @ y_rotated).T
+        return (py @ x.T)[0], _sT
+
+    def impl(y, b, r, phi):
+        if abs(b) >= (r + 1):
+            raise ValueError("Impact parameter out of bounds")
+        else:
+            if r > 1 and abs(b) <= (r - 1):
+                raise ValueError("Impact parameter out of bounds")
+            else:
+                return occ_flux(y, b, r, phi)
+
+    return impl
+
+
+def is_occ(b, r):
[…]

### 2024-08-09 workflows/precision/scripts/plot_br_error.py

@@ -1,0 +1,62 @@
+import matplotlib.pyplot as plt
+import matplotlib.colors as colors
+import numpy as np
+
+orders = snakemake.params.orders
+br = np.load(snakemake.input["br"])
+r = br["r"]
+b = br["b"]
+data = {
+    "starry": np.load(snakemake.input["starry"]),
+    "jaxoplanet": {
+        orders[i]: np.load(snakemake.input["jaxoplanet"][i]) for i in range(len(orders))
+    },
+}
+
+
+R, B = np.meshgrid(r, b)
+
+
+def colorbar(mappable):
+    from mpl_toolkits.axes_grid1 import make_axes_locatable
+    import matplotlib.pyplot as plt
+
+    last_axes = plt.gca()
+    ax = mappable.axes
+    fig = ax.figure
+    divider = make_axes_locatable(ax)
+    cax = divider.append_axes("right", size="5%", pad=0.05)
+    cbar = fig.colorbar(mappable, cax=cax)
+    plt.sca(last_axes)
+    return cbar
+
+
+fig, axes = plt.subplots(1, len(orders), figsize=(8.5, 2.5))
+# norm = colors.LogNorm(vmin=1e-15, vmax=1e-8)
+for i, order in enumerate(orders):
+    ax = axes[i]
+    ax.set_aspect("equal", "box")
+    errors = np.abs(data["starry"]["flux"] - data["jaxoplanet"][order]["flux"])
+    errors[errors == 0] = 1e-20
+    errors[np.logical_not(np.isfinite(errors))] = 1e-20
+
+    norm = colors.LogNorm(vmin=1e-15, vmax=np.max(np.abs(errors)))
+    pcm = ax.pcolor(R, B, errors.T, norm=norm, cmap="PuBu_r", shading="auto")
+    k = dict(color="k", linestyle="-", linewidth=1)
+    ax.axline((0, 0), slope=1, **k)
+    ax.axline((1, 0), slope=1, **k)
+    ax.axline((0, 1), slope=1, **k)
+    ax.axline((0, 1), slope=-1, **k)
+    if i > 0:
+        ax.set_yticklabels([])
+    else:
+        ax.set_ylabel("b")
+    ax.set_xlabel("r")
+    ax.set_title(rf"$n$ = {order}")
+    colorbar(pcm)
+
+# cm_ax = fig.add_axes([1.5, 0.1, 0.02, 0.8])
+# place colorbar at cm_ax
+
[…]

### 2024-08-09 workflows/precision/scripts/plot_error.py

@@ -10,7 +10,7 @@ from jaxoplanet.experimental.starry.ylm import Ylm
 from jaxoplanet.experimental.starry.mpcore import mp, utils
 
 ref = "num"
-comp = snakemake.wildcards.code
+comp = "comp"
 
 data = {
     "small": {

### 2024-08-09 workflows/precision/scripts/plot_error_GL.py

@@ -1,0 +1,58 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+data = np.load(snakemake.input[0], allow_pickle=True)
+
+errors_f = np.array([[i.max() for i in err] for err in data["errors_f"]])
+errors_s = np.array([[i.max() for i in err] for err in data["errors_s"]])
+orders = np.array(data["orders"])
+radii = np.array(data["rs"])
+
+plt.figure(figsize=(8, 3))
+fig, axes = plt.subplots(1, 2, figsize=(8, 3), sharey=True)
+ax1, ax2 = axes
+cmap = plt.get_cmap("plasma")
+
+for i, r in enumerate(radii):
+    color = cmap(i / len(radii))
+    axes[0].plot(orders, errors_s[i], ".-", label=f"r={r}", color=color)
+    axes[1].plot(orders, errors_f[i], ".-", label=f"r={r}", color=color)
+
+for ax in axes:
+    ax.set_yscale("log")
+    ax.set_xlabel("order")
+    ax.set_ylim(1e-16, 1)
+
+axes[0].set_ylabel("max relative error")
+axes[0].set_title("solution vector")
+axes[1].set_title("flux")
+axes[0].legend()
+
+plt.tight_layout()
+
+for axis in axes:
+    axis.axhline(1e-6, ls="--", lw=1, color="k", alpha=0.2)
+    axis.axhline(1e-9, ls="--", lw=1, color="k", alpha=0.2)
+    axis.annotate(
+        "ppm",
+        xy=(0, 1e-6),
+        xycoords="data",
+        xytext=(3, -3),
+        textcoords="offset points",
+        ha="left",
+        va="top",
+        alpha=0.75,
+    )
+    axis.annotate(
+        "ppb",
+        xy=(0, 1e-9),
+        xycoords="data",
+        xytext=(3, -3),
+        textcoords="offset points",
+        ha="left",
+        va="top",
+        alpha=0.75,
+    )
+
+plt.tight_layout()
+plt.savefig(snakemake.output[0])

### 2024-08-09 workflows/precision/scripts/plot_error_limbdark.py

@@ -1,0 +1,99 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from jaxoplanet.experimental.starry.mpcore import mp, utils
+
+degree = snakemake.params.degree
+
+data = {
+    "small": {
+        "jax": np.load(snakemake.input.jax[0], allow_pickle=True)["f"],
+        "num": np.load(snakemake.input.num[0], allow_pickle=True)["f"],
+        "starry": np.load(snakemake.input.starry[0], allow_pickle=True)["f"],
+        "exo": np.load(snakemake.input.exo[0], allow_pickle=True)["f"],
+    },
+    "large": {
+        "jax": np.load(snakemake.input.jax[1], allow_pickle=True)["f"],
+        "num": np.load(snakemake.input.num[1], allow_pickle=True)["f"],
+        "starry": np.load(snakemake.input.starry[1], allow_pickle=True)["f"],
+        "exo": np.load(snakemake.input.exo[1], allow_pickle=True)["f"],
+    },
+}
+
+
+def errors(M1, M2):
+    return np.abs(utils.diff_mp(M1, M2))
+
+
+err = {
+    "small": {
+        "jax": errors(data["small"]["jax"], data["small"]["num"]).max(1)[1:],
+        "starry": errors(data["small"]["starry"], data["small"]["num"]).max(1)[1:],
+        "exo": errors(data["small"]["exo"], data["small"]["num"][0:3, :]).max(1)[1:],
+    },
+    "large": {
+        "jax": errors(data["large"]["jax"], data["large"]["num"]).max(1)[1:],
+        "starry": errors(data["large"]["starry"], data["large"]["num"]).max(1)[1:],
+        "exo": errors(data["large"]["exo"], data["large"]["num"][0:3, :]).max(1)[1:],
+    },
+}
+color = {
+    "jax": "C0",
+    "starry": "k",
+    "exo": "C4",
+}
+
+
+fig = plt.figure(figsize=(8, 3))
+x = np.arange(0, degree + 2, 2)
+
+degrees = np.arange(1, degree + 1)
+
+ax1 = plt.subplot(121)
+ax1.plot(degrees, err["small"]["jax"], ".-", label="jaxoplanet", c=color["jax"])
+ax1.plot(degrees, err["small"]["starry"], ".-", label="starry", c=color["starry"])
[…]

### 2024-08-09 workflows/precision/scripts/plot_error_several_r_GL_degree_SH.py

@@ -1,0 +1,51 @@
+"""Plot of the errors on the flux for different radii, orders of the
+Gauss-Legendre quadrature, and spherical harmonic degrees.
+"""
+
+from jaxoplanet.experimental.starry import Ylm
+import matplotlib.pyplot as plt
+import numpy as np
+
+data = np.load(snakemake.input[0], allow_pickle=True)
+l_max = snakemake.params.l_max
+cmap = plt.get_cmap("plasma")
+
+orders = np.array(data["orders"])
+radii = np.array(data["rs"])
+
+indices = {
+    l: np.array([Ylm.index(l, m) for m in range(-l, l + 1)]) for l in range(l_max + 1)
+}
+
+errors_f = np.array([[i.max(0) for i in err] for err in data["errors_f"]])
+errors_f = np.array([errors_f[:, :, indices[l]].max(-1) for l in range(l_max + 1)])
+
+degrees = np.arange(l_max + 1)
+
+fig, axes = plt.subplots(1, 3, figsize=(8, 3), sharey=True)
+
+for axi, ri in enumerate([0, 2, 4]):
+    ax = axes[axi]
+    for i, error in enumerate(errors_f[:, ri, :]):
+        color = cmap(i / l_max)
+        ax.plot(orders, error, ".-", color=color)
+        ax.set_title(f"$r={radii[ri]}$")
+        ax.set_yscale("log")
+
+ax.set_ylim(1e-16, 1)
+axes[1].set_xlabel("order $n$ of the Gauss-Legendre quadrature")
+axes[0].set_ylabel("max relative error on the flux")
+
+l_arr = [0, 5, 10, 20]
+
+# Dummy curves & a legend
+lines = [None for l in l_arr]
+leg_labels = ["%d" % l for l in l_arr]
+for i, l in enumerate(l_arr):
+    (lines[i],) = ax.plot((0, 1), (1e-20, 1e-20), color=cmap(l / (l_max + 2)), lw=2)
+leg = fig.legend(lines, leg_labels, title="Degree")
+
+plt.tight_layout()
+
+plt.tight_layout()
+plt.savefig(snakemake.output[0])

### 2024-08-09 workflows/precision/scripts/plot_error_vs_degree.py

@@ -73,7 +73,7 @@ ax1.plot(
 ax1.plot(
     degree, err_per_degree(err["small"]["starry"]), ".-", label="starry", c=s_color
 )
-ax1.set_xlabel("degree")
+ax1.set_xlabel("degree of spherical harmonics")
 ax1.set_ylabel("relative error")
 ax1.set_title("$r=0.01$")
 ax1.set_yscale("log")
@@ -88,6 +88,7 @@ ax2.plot(
     degree, err_per_degree(err["large"]["starry"]), ".-", label="starry", c=s_color
 )
 ax2.legend()
+ax2.set_xlabel("degree of spherical harmonics")
 ax2.set_title("$r=100$")
 ax2.set_yscale("log")
 ax2.set_ylim(1e-17, 1e-5)

### 2024-08-09 workflows/precision/scripts/starry_br_flux.py

@@ -1,0 +1,19 @@
+from tqdm import tqdm
+import starry
+import numpy as np
+
+starry.config.lazy = False
+
+ms = starry.Map(ydeg=1, udeg=1)
+u = (1.0,)
+ms[1:] = u
+
+b, r = np.load(snakemake.input[0]).values()
+
+calculated = np.zeros((len(r), len(b)))
+
+for i, _r in enumerate(tqdm(r)):
+    expected = ms.flux(ro=_r, yo=b, xo=0.0, zo=10.0)
+    calculated[i, :] = expected
+
+np.savez(snakemake.output[0], flux=calculated)

### 2024-08-09 workflows/precision/scripts/starry_f_limbdark.py

@@ -1,0 +1,35 @@
+import numpy as np
+import starry
+
+starry.config.lazy = False
+
+
+r = float(snakemake.wildcards.r)
+bs = np.load(snakemake.input[0], allow_pickle=True)["bs"]
+bs = np.abs(np.concatenate(bs))
+degree = snakemake.params.degree
+
+
+map = starry.Map(ydeg=degree, udeg=degree)
+
+flux = []
+from tqdm import tqdm
+
+
+def u(deg):
+    u = np.zeros(degree)
+    if degree == 0:
+        return u
+    else:
+        u[deg - 1] = 1
+        return u
+
+
+for deg in tqdm(range(degree + 1)):
+    map[1:] = u(deg)
+    f = map.flux(xo=0.0, yo=bs, zo=10.0, ro=r)
+    flux.append(f)
+
+flux = np.array(flux)
+
+np.savez(snakemake.output[0], f=flux)

### 2024-08-09 workflows/speed/scripts/flux_exoplanet_quadratic.py

@@ -1,5 +1,4 @@
-from tqdm import tqdm
-import starry
+from exoplanet.light_curves import LimbDarkLightCurve
 import numpy as np
 import timeit
 
@@ -17,18 +16,8 @@ def timeit_f(strf, n=10):
     return times
 
 
-starry.config.lazy = False
-
-if u is not None:
-    ms = starry.Map(ydeg=len(u), udeg=len(u))
-    ms[1:] = u
-else:
-    l_max = 20
-    ms = starry.Map(ydeg=l_max)
-    y = np.ones((l_max + 1) ** 2)
-    ms[:, :] = y
-
-
-times = timeit_f(f"ms.flux(ro={radius}, yo=b, xo=0.0, zo=10.0)")
+lc_object = LimbDarkLightCurve(*u)
+radii = np.ones_like(b) * radius
+times = timeit_f(f"lc_object._compute_light_curve(b, radii)")
 
 np.savez(snakemake.output[0], time=times, N=N, b=b)

### 2024-08-09 workflows/speed/scripts/flux_jaxoplanet.py

@@ -4,121 +4,41 @@ from tqdm import tqdm
 jax.config.update("jax_platform_name", "cpu")
 jax.config.update("jax_enable_x64", True)
 
+import timeit
 import numpy as np
-from jaxoplanet.experimental.starry import Surface
+from jaxoplanet.experimental.starry import Surface, Ylm
+from jaxoplanet.experimental.starry.light_curves import surface_light_curve
 import jax.numpy as jnp
 
-data = dict(np.load(snakemake.input[0]))
+radius = float(snakemake.params.radius)
+u = snakemake.params.u
+N = int(snakemake.wildcards.N)
+assert radius < 1
+b = np.linspace(0, 1 + radius, N)
 order = int(snakemake.wildcards.order)
-u = data["u"]
-surface = Surface(period=None, u=(u))
-
-from jaxoplanet.experimental.starry.light_curves import *
-
-import timeit
 
 
-def timeit_f(strf, n=10):
-    times = np.array(timeit.repeat(f"{strf}", number=n, globals=globals()))[1:] / n
-    return np.median(times)
+def timeit_f(strf, n):
+    times = np.array(timeit.repeat(f"{strf}", number=n, globals=globals()))[1:] / (
+        n - 1
+    )
+    return times
 
 
-# TODO: figure out the sparse matrices (and Pijk) to avoid todense()
-def surface_light_curve(
-    map,
-    r: float = None,
-    xo: float = None,
-    yo: float = None,
-    zo: float = None,
-    theta: float = 0.0,
-    order: int = 20,
-):
-    """Light curve of an occulted map.
-
-    Args:
-        map (Map): map object
-        r (float or None): radius of the occulting body, relative to the current map
-           body
-        xo (float or None): x position of the occulting body, relative to the current
-           map body
-        yo (float or None): y position of the occulting body, relative to the current
-           map body
-        zo (float or None): z position of the occulting body, relative to the current
-           map body
-        theta (float): rotation angle of the map
-
-    Returns:
-        ArrayLike: flux
-    """
-    rT_deg = rT(map.deg)
-
-    # no occulting body
-    if r is None:
-        b_rot = True
[…]

### 2024-08-09 workflows/speed/scripts/plot_error_vs_order.py

@@ -1,0 +1,51 @@
+import numpy as np
+
+orders = snakemake.params.orders
+
+jax_times_quadratic = np.array(
+    [np.median(np.load(f)["time"]) for f in snakemake.input.jaxoplanet_quadratic]
+)
+jax_times_quadratic_gpu = np.array(
+    [np.median(np.load(f)["time"]) for f in snakemake.input.jaxoplanet_quadratic_gpu]
+)
+starry_time_quadratic = np.median(np.load(snakemake.input.starry_quadratic)["time"])
+exoplanet_time_quadratic = np.median(
+    np.load(snakemake.input.exoplanet_quadratic)["time"]
+)
+
+jax_times_l20 = np.array(
+    [np.median(np.load(f)["time"]) for f in snakemake.input.jaxoplanet_l20]
+)
+
+jax_times_l20_gpu = np.array(
+    [np.median(np.load(f)["time"]) for f in snakemake.input.jaxoplanet_l20_gpu]
+)
+
+starry_times_l20 = np.median(np.load(snakemake.input.starry_l20)["time"])
+
+import matplotlib.pyplot as plt
+
+plt.figure(figsize=(8, 3.5))
+
+plt.subplot(121)
+plt.plot(orders, jax_times_quadratic * 1e3, ".-", label="jaxoplanet (CPU)")
+plt.plot(orders, jax_times_quadratic_gpu * 1e3, ".-", label="jaxoplanet (GPU)")
+plt.axhline(starry_time_quadratic * 1e3, color="k", label="starry")
+plt.axhline(exoplanet_time_quadratic * 1e3, color="C4", label="exoplanet")
+plt.legend()
+plt.xlabel("order of Gauss-Legendre quadrature")
+plt.ylabel("time (ms)")
+plt.title("Quadratic limb-darkened map ($l=2$)")
+plt.yscale("log")
+
+plt.subplot(122)
+plt.plot(orders, jax_times_l20 * 1e3, ".-", label="jaxoplanet (CPU)")
+plt.plot(orders, jax_times_l20_gpu * 1e3, ".-", label="jaxoplanet (GPU)")
+plt.axhline(starry_times_l20 * 1e3, color="k", label="starry")
+plt.legend()
+plt.xlabel("order of Gauss-Legendre quadrature")
+plt.title("Non-uniform map ($l=20$)")
+plt.yscale("log")
+
+plt.tight_layout()
+plt.savefig(snakemake.output[0])

### 2024-08-13 latex/paper.tex
feat: update speed benchmark

@@
 In order to infer these properties, a fast, accurate and robust model of stellar light curves is needed, one that accounts for the non-uniform surface intensity of spherical bodies and their mutual occultations.
-The analytical framework known as \texttt{starry} satisfies these requirements, and was successfully applied over the years following its development.
+The analytical framework known as \textsf{starry} satisfies these requirements, and was successfully applied over the years following its development.
 However, as datasets and model complexity grow, the tractable inference of these parameters becomes increasingly challenging, and often motivates biased approximations.
-In this paper, we present a new implementation of the \texttt{starry} framework in \texttt{jax}, a high-performance machine-learning library designed for large-scale computations.
-We describe the changes made to the original version of \texttt{starry} and evaluate the performance of its new implementation that we release in the open-source Python package \texttt{jaxoplanet}.
+In this paper, we present a new implementation of the \textsf{starry} framework in \texttt{jax}, a high-performance machine-learning library designed for large-scale computations.
+We describe the changes made to the original version of \textsf{starry} and evaluate the performance of its new implementation that we release in the open-source Python package \textsf{jaxoplanet}.
 Through this implementation, we provide the community with a differentiable model of stellar light curves, and enable the use of powerful machine-learning tools available in the \texttt{jax} ecosystem.
@@
 Hence, knowing how to decompose the full spherical harmonics rotation using pre-computed rotations is essential to achieve optimal performance in the \textit{starry} framework.
-\autoref{fig:rotation_starry} shows the 6 elementary rotations currently used in the \texttt{starry} implementation.
+\autoref{fig:rotation_starry} shows the 6 elementary rotations currently used in the \textsf{starry} implementation.
 \begin{figure}[H]
@@
         \includegraphics[width=\textwidth]{../workflows/rotations/figures/rotation_starry.pdf}
-        \caption{Consecutive rotations of the spherical harmonics basis in the \texttt{starry} implementation.
+        \caption{Consecutive rotations of the spherical harmonics basis in the \textsf{starry} implementation.
 Each schematic shows a sphere rotated from a previous orientation around an axis displayed as a black vector and an angle shown on top.}
@@
 This explains the current decomposition of the complete rotation in six separate steps.\\\\
-In \texttt{jaxoplanet}, we compute the Wigner-D-matrices by employing the Risbo recursion relations \citep{Risbo1996} implemented in \texttt{JAX} as part of the \texttt{s2fft} Python package \citep{price:s2fft}.
+In \textsf{jaxoplanet}, we compute the Wigner-D-matrices by employing the Risbo recursion relations \citep{Risbo1996} implemented in \texttt{JAX} as part of the \texttt{s2fft} Python package \citep{price:s2fft}.
 In addition, we merge rotations 3, 4 and 5 (see \autoref{fig:rotation_jaxoplanet}) into a single compound rotation of axis
@@
 \begin{equation}
+    \label{eq:combined_angle}
     \omega = 2 \cos^{-1}{\left(\cos{\left(\frac{inc}{2} \right)} \cos{\left(\frac{obl}{2}\right)} \right)}.
@@
         \includegraphics[width=\textwidth]{../workflows/rotations/figures/rotation_jaxoplanet_1.pdf}
-        \caption{Consecutive rotations of the spherical harmonics basis in the \texttt{jaxoplanet} implementation.
+        \caption{Consecutive rotations of the spherical harmonics basis in the \textsf{jaxoplanet} implementation.
 Each schematic shows a sphere rotated from a previous orientation around an axis displayed as a black vector and an angle shown on top.}
@@
 Although \textit{starry} occultation light curves have analytical expressions, their practical implementation relies on numerical approximations chosen to balance precision and computation time.
-This is the case, for example, of the solution vector $\bvec{s}$, computed using numerical series expansions in \texttt{starry} (\citealt[section D.2.3]{starry}) and numerical integration in \texttt{jaxoplanet} (see \autoref{solution_vectors}).
+This is the case, for example, of the solution vector $\bvec{s}$, computed using numerical series expansions in \textsf{starry} (\citealt[section D.2.3]{starry}) and numerical integration in \textsf{jaxoplanet} (see \autoref{solution_vectors}).
 Other expressions, such as the Wigner-D matrices used in the rotation of the spherical harmonics basis, are computed using reccurence relations prone to the accumulation of numerical errors and instability.\\\\
-In this section, we evaluate the accuracy and speed of the \texttt{jaxoplanet} implementation and compare it with the C++ implementation of \texttt{starry}.
+In this section, we evaluate the accuracy and speed of the \textsf{jaxoplanet} implementation and compare it with the C++ implementation of \textsf{starry}.
 
 \subsection{Precision}
-We evaluate the precision of the \texttt{starry} and \texttt{jaxoplanet} terms against quantities computed at arbitrary precision, using a ground truth version of the code implemented with the \texttt{mpmath} Python library\footlink{https://mpmath.org/}.
+We evaluate the precision of the \textsf{starry} and \textsf{jaxoplanet} terms against quantities computed at arbitrary precision, using a ground truth version of the code implemented with the \texttt{mpmath} Python library\footlink{https://mpmath.org/}.
 The precision of the overall flux $f$ can only be understood by considering the precision of the terms involved in \autoref{eq:starry}: the solution vector $\bvec{s}$ ($\bvec{r}$ if no occultation), the basis matrices $\bvec{A}$ and $\bvec{B}$, and the Wigner-D rotation tensor $\bvec{R}$.
-Although we show in \autoref{fig:precision_SAR} that the precision of the \texttt{starry} and \texttt{jaxoplanet} implentations differ for many of these terms, we focus this section on the solution vector $\bvec{s}$ and the overall flux $f$.\\\\
-Assuming that we perform the numerical integration of the solution vector $\bvec{s}$ at a relatively high order ($n=500$; see \autoref{solution_vectors}), \autoref{fig:precision_s} shows that our reimplentation of \texttt{starry} reaches relative errors lower than 1 part per billion for $\bvec{s}$ and the overall flux $f$, considering both a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$.
-For reference, \autoref{fig:precision_s_starry} shows comparable errors on the same quantities computed with the C++ implentation of \texttt{starry}.
-In addition, \autoref{fig:precision_degree} shows the evolution of the error on $f$ for increasing degrees of the spherical harmonics, both for \texttt{starry} and \texttt{jaxoplanet}.
-With these results, we validate the precision of \texttt{jaxoplanet} ans show it is on per with the legacy C++ \texttt{starry} implentation\footnote{version 1.2.0 \citep{starry_120}} described in \cite{starry}.\\\\
+Although we show in \autoref{fig:precision_SAR} that the precision of the \textsf{starry} and \textsf{jaxoplanet} implentations differ for many of these terms, we focus this section on the solution vector $\bvec{s}$ and the overall flux $f$.\\\\
+Assuming that we perform the numerical integration of the solution vector $\bvec{s}$ at a relatively high order ($n=500$; see \autoref{solution_vectors}), \autoref{fig:precision_s} shows that our reimplentation of \textsf{starry} reaches relative errors lower than 1 part per billion for $\bvec{s}$ and the overall flux $f$, considering both a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$.
+For reference, \autoref{fig:precision_s_starry} shows comparable errors on the same quantities computed with the C++ implentation of \textsf{starry}.
+In addition, \autoref{fig:precision_degree} shows the evolution of the error on $f$ for increasing degrees of the spherical harmonics, both for \textsf{starry} and \textsf{jaxoplanet}.
+With these results, we validate the precision of \textsf{jaxoplanet} ans show it is on per with the legacy C++ \textsf{starry} implentation\footnote{version 1.2.0 \citep{starry_120}} described in \cite{starry}.\\\\
 \begin{figure}[H]
@@
         \caption{relative numerical error in the flux $f$ of a linearly limb-darkened star (degree $l$=1) transited by an opaque companion over the $(r, b)$ parameter space.
-The error is evaluated for different orders $n$ of the Gauss-Legendre quadrature and computed against \texttt{starry} for speed and convenience.
+The error is evaluated for different orders $n$ of the Gauss-Legendre quadrature and computed against \textsf{starry} for speed and convenience.
 Note that \autoref{fig:precision_s} was produced using $n=500$ (right plot) for which $\bvec{s}$ reaches machine precision.
@@
 \end{figure}
-In \autoref{fig:error_order_gausslegendre}, we show how the error in the solution vector evolves with the order $n$ of the Gauss-Legendre quadrature, and its impact on the error in the overall flux.
+In \autoref{fig:precision_order_gausslegendre}, we show how the error in the solution vector evolves with the order $n$ of the Gauss-Legendre quadrature, and its impact on the error in the overall flux.
 \begin{figure}[H]
@@
 As in \autoref{fig:precision_s} the maximum degree of the spherical harmonics basis is set to $l=20$.}
-        \label{fig:error_order_gausslegendre}
+        \label{fig:precision_order_gausslegendre}
     \end{center}
@@
 Hence users must carefully set the order parameter to balance precision and computational speed.
-As the minimum error reached also depends on the degree of the spherical harmonics basis used, \autoref{fig:error_order_gausslegendre_degree} shows the errors on the flux for each degree of the spherical harmonics basis, for an occultor with relative radii $r=0.01$, $r=1$ and $r=100$.
+As the minimum error reached also depends on the degree of the spherical harmonics basis used, \autoref{fig:precision_order_gausslegendre_degree} shows the errors on the flux for each degree of the spherical harmonics basis, for an occultor with relative radii $r=0.01$, $r=1$ and $r=100$.
 \begin{figure}[H]
@@
 Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s}, maxima taken over the range of values $b$.}
-        \label{fig:error_order_gausslegendre_degree}
+        \label{fig:precision_order_gausslegendre_degree}
     \end{center}
@@
 
-The main advantage of the \texttt{jax} library is its ability to perform hardware-accelerated computations on CPUs and GPUs.
-In this section, we evaluate the speed of the \texttt{jaxoplanet} implementation and compare it with the C++ implementation of \texttt{starry}.
+One of the advantage of the \texttt{JAX} library is its ability to perform hardware-accelerated computations on CPUs and GPUs.
+In this section, we evaluate the speed of the \textsf{jaxoplanet} implementation and compare it with the C++ implementation of \textsf{starry}.
 For quadratically limb-darkened light curve, we also compare our implementation with the one provided by the \textsf{exoplanet} Python package.\\\\
-In \autoref{fig:speed_order_gausslegendre}, we show the evolution of the computing time required to compute an occultation light-curve for a limb-darkened star and a non-uniform star described by spherical harmonics up to order $l=20$.
-This figure include the value of the order $n$ of the Gauss-Legendre quadrature required to reach a given precision, and illustrate well the trade-off between precision and speed that users must consider when setting the order parameter.\\\\
+In \autoref{fig:speed_order_gausslegendre}, we show the evolution of the computing time required to compute an occultation light-curve of a limb-darkened star and that of a non-uniform star described by spherical harmonics with a maximum degrees $l=20$, depending on the order $n$ of the Gauss-Legendre quadrature used to compute the $\mathcal{P}$ integral numerically (see \autoref{solution_vectors}).\\\\
 \begin{figure}[H]
@@
         \includegraphics[width=\textwidth]{../workflows/speed/figures/speed_vs_order.pdf}
-        \caption{Computation time of the occultation light curve of a quadts}
+        \caption{Computation time of the occultation light curve of a limb-darkened star (left) and a non-uniform star described by spherical harmonics up to order $l=20$ (right) depending on the order $n$ of the Gauss-Legendre quadrature used to compute the P integral numerically (see \autoref{solution_vectors}).
+The processing time reported for \textsf{exoplanet} and \textsf{starry} is independent of $n$ as these two codes provide closed-form solutions for the integral $\mathcal{P}$.
+Light curves are computed for $10\,000$ points in transit and an occultor radius $r=0.1$.}
         \label{fig:speed_order_gausslegendre}
@@
 \end{figure}
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[width=\textwidth]{../workflows/speed/figures/speed_vs_N.pdf}
+        \caption{Computation time of the occultation light curve of a limb-darkened star (left) and a non-uniform star described by spherical harmonics up to order $l=20$ (right) depending on the number of points in the light curve.
+Light curves are computed at order $n=200$ and an occultor radius $r=0.1$.}
+        \label{fig:speed_N}
+    \end{center}
+\end{figure}
+\autoref{fig:speed_order_gausslegendre}, in combination with \autoref{fig:precision_order_gausslegendre_degree}, can be used to determine a value of $n$ balancing precision and computation time.\\\\
+Finally, \autoref{fig:speed_N} shows the evaluation time of \textsf{jaxoplanet} as a function of the number of points in the light curve, compared to \textsf{starry} for an $l=20$ surface map, and \textsf{exoplanet} for a quadratically limb-darkened star.
+As in the previous figure, we show that \textsf{jaxoplanet} is competitive against state-of-the-art codes on CPUs and offer a clear advantage on GPUs.
 
-\todo{} shows
+\section{Case studies: ...}
 
-\section{Case study: ...}
+\subsection{}
+
+\bibliography{ref}
 
@@
         \includegraphics[width=\textwidth]{../workflows/precision/figures/error_starry.pdf}
-        \caption{Same as \autoref{fig:precision_s} but with $bvec{s}$ and $f$ computed with the C++ implementation of \texttt{starry}. \codelink{workflows/precision/scripts/plot_error}}
+        \caption{Same as \autoref{fig:precision_s} but with $bvec{s}$ and $f$ computed with the C++ implementation of \textsf{starry}. \codelink{workflows/precision/scripts/plot_error}}
         \label{fig:precision_s_starry}
@@
 
-\bibliography{ref}
+% Table of symbols
+
+\clearpage
+\begin{center}
+\renewcommand*{\arraystretch}{1.08}
+\begin{longtable}{cll}
+\caption{Symbols used in this paper} \label{tab:symbols} \\
+%
+\toprule
+\multicolumn{1}{c}{\textbf{Symbol}} &
+\multicolumn{1}{c}{\textbf{Definition}} &
+\multicolumn{1}{c}{\textbf{Reference}} \\
+\midrule
+\endfirsthead
+%
+\multicolumn{3}{c}%
+{{\bfseries \tablename\ \thetable{} --} continued from previous page} \\
+\toprule
+\multicolumn{1}{c}{\textbf{Symbol}} &
+\multicolumn{1}{c}{\textbf{Definition}} &
+\multicolumn{1}{c}{\textbf{Reference}} \\
+\midrule
+\endhead
+\bottomrule
+%
+\endfoot
+%
+\bottomrule
+\endlastfoot
+%
+$\omega$        & Rotation angle of the combined rotation       & \autoref{eq:combined_angle} \\
+$n$ & order of the Gauss-Legendre approximation & \\
+$r$ & occultor radius in units of occulted body's radius & \\
+%
+\end{longtable}
+\end{center}
 

### 2024-08-13 workflows/precision/scripts/plot_error_GL.py

@@ -33,26 +33,27 @@ plt.tight_layout()
 for axis in axes:
     axis.axhline(1e-6, ls="--", lw=1, color="k", alpha=0.2)
     axis.axhline(1e-9, ls="--", lw=1, color="k", alpha=0.2)
-    axis.annotate(
-        "ppm",
-        xy=(0, 1e-6),
-        xycoords="data",
-        xytext=(3, -3),
-        textcoords="offset points",
-        ha="left",
-        va="top",
-        alpha=0.75,
-    )
-    axis.annotate(
-        "ppb",
-        xy=(0, 1e-9),
-        xycoords="data",
-        xytext=(3, -3),
-        textcoords="offset points",
-        ha="left",
-        va="top",
-        alpha=0.75,
-    )
+
+axes[0].annotate(
+    "ppm",
+    xy=(0, 1e-6),
+    xycoords="data",
+    xytext=(3, 3),
+    textcoords="offset points",
+    ha="left",
+    va="bottom",
+    alpha=0.75,
+)
+axes[0].annotate(
+    "ppb",
+    xy=(80, 1e-9),
+    xycoords="data",
+    xytext=(3, 3),
+    textcoords="offset points",
+    ha="left",
+    va="bottom",
+    alpha=0.75,
+)
 
 plt.tight_layout()
 plt.savefig(snakemake.output[0])

### 2024-08-13 workflows/precision/scripts/plot_error_several_r_GL_degree_SH.py

@@ -45,7 +45,31 @@ for i, l in enumerate(l_arr):
     (lines[i],) = ax.plot((0, 1), (1e-20, 1e-20), color=cmap(l / (l_max + 2)), lw=2)
 leg = fig.legend(lines, leg_labels, title="Degree")
 
-plt.tight_layout()
+for axis in axes:
+    axis.axhline(1e-6, ls="--", lw=1, color="k", alpha=0.2)
+    axis.axhline(1e-9, ls="--", lw=1, color="k", alpha=0.2)
+
+
+axes[0].annotate(
+    "ppm",
+    xy=(0, 1e-6),
+    xycoords="data",
+    xytext=(3, 3),
+    textcoords="offset points",
+    ha="left",
+    va="bottom",
+    alpha=0.75,
+)
+axes[0].annotate(
+    "ppb",
+    xy=(0, 1e-9),
+    xycoords="data",
+    xytext=(3, 3),
+    textcoords="offset points",
+    ha="left",
+    va="bottom",
+    alpha=0.75,
+)
 
 plt.tight_layout()
 plt.savefig(snakemake.output[0])

### 2024-08-13 workflows/speed/scripts/flux_exoplanet_quadratic.py

@@ -18,6 +18,6 @@ def timeit_f(strf, n=10):
 
 lc_object = LimbDarkLightCurve(*u)
 radii = np.ones_like(b) * radius
-times = timeit_f(f"lc_object._compute_light_curve(b, radii)")
+times = timeit_f(f"lc_object._compute_light_curve(b, radii).eval()")
 
 np.savez(snakemake.output[0], time=times, N=N, b=b)

### 2024-08-13 workflows/speed/scripts/flux_jaxoplanet_quadratic.py

@@ -1,0 +1,33 @@
+import jax
+from tqdm import tqdm
+
+jax.config.update("jax_platform_name", "cpu")
+jax.config.update("jax_enable_x64", True)
+
+import timeit
+import numpy as np
+from jaxoplanet.experimental.starry import Surface, Ylm
+from jaxoplanet.core.limb_dark import light_curve
+from jaxoplanet.experimental.starry.light_curves import surface_light_curve
+import jax.numpy as jnp
+
+radius = float(snakemake.params.radius)
+u = snakemake.params.u
+N = int(snakemake.wildcards.N)
+assert radius < 1
+b = np.linspace(0, 1 + radius, N)
+order = int(snakemake.wildcards.order)
+
+
+def timeit_f(strf, n):
+    times = np.array(timeit.repeat(f"{strf}", number=n, globals=globals()))[1:] / (
+        n - 1
+    )
+    return times
+
+
+flux_function = jax.jit(jax.vmap(lambda b: light_curve(u, b, radius, order=order)))
+times = timeit_f(f"flux_function(b).block_until_ready()", n=50)
+
+
+np.savez(snakemake.output[0], time=times, N=N, b=b)

### 2024-08-13 workflows/speed/scripts/plot_error_vs_order.py

@@ -28,10 +28,10 @@ import matplotlib.pyplot as plt
 plt.figure(figsize=(8, 3.5))
 
 plt.subplot(121)
-plt.plot(orders, jax_times_quadratic * 1e3, ".-", label="jaxoplanet (CPU)")
-plt.plot(orders, jax_times_quadratic_gpu * 1e3, ".-", label="jaxoplanet (GPU)")
 plt.axhline(starry_time_quadratic * 1e3, color="k", label="starry")
 plt.axhline(exoplanet_time_quadratic * 1e3, color="C4", label="exoplanet")
+plt.plot(orders, jax_times_quadratic * 1e3, ".-", label="jaxoplanet (CPU)", c="C0")
+plt.plot(orders, jax_times_quadratic_gpu * 1e3, ".-", label="jaxoplanet (GPU)", c="C3")
 plt.legend()
 plt.xlabel("order of Gauss-Legendre quadrature")
 plt.ylabel("time (ms)")
@@ -39,9 +39,9 @@ plt.title("Quadratic limb-darkened map ($l=2$)")
 plt.yscale("log")
 
 plt.subplot(122)
-plt.plot(orders, jax_times_l20 * 1e3, ".-", label="jaxoplanet (CPU)")
-plt.plot(orders, jax_times_l20_gpu * 1e3, ".-", label="jaxoplanet (GPU)")
 plt.axhline(starry_times_l20 * 1e3, color="k", label="starry")
+plt.plot(orders, jax_times_l20 * 1e3, ".-", label="jaxoplanet (CPU)")
+plt.plot(orders, jax_times_l20_gpu * 1e3, ".-", label="jaxoplanet (GPU)", c="C3")
 plt.legend()
 plt.xlabel("order of Gauss-Legendre quadrature")
 plt.title("Non-uniform map ($l=20$)")

### 2024-08-13 workflows/speed/scripts/plot_speed_vs_N.py

@@ -1,0 +1,57 @@
+import numpy as np
+
+Ns = snakemake.params.Ns
+
+jax_times_quadratic = np.array(
+    [np.median(np.load(f)["time"]) for f in snakemake.input.jaxoplanet_quadratic]
+)
+starry_time_quadratic = np.array(
+    [np.median(np.load(f)["time"]) for f in snakemake.input.starry_quadratic]
+)
+exoplanet_time_quadratic = np.array(
+    [np.median(np.load(f)["time"]) for f in snakemake.input.exoplanet_quadratic]
+)
+
+jax_times_l20 = np.array(
+    [np.median(np.load(f)["time"]) for f in snakemake.input.jaxoplanet_l20]
+)
+starry_time_l20 = np.array(
+    [np.median(np.load(f)["time"]) for f in snakemake.input.starry_l20]
+)
+
+jax_times_l20_gpu = np.array(
+    [np.median(np.load(f)["time"]) for f in snakemake.input.jaxoplanet_l20_gpu]
+)
+
+jax_times_quadratic_gpu = np.array(
+    [np.median(np.load(f)["time"]) for f in snakemake.input.jaxoplanet_quadratic_gpu]
+)
+
+import matplotlib.pyplot as plt
+
+plt.figure(figsize=(8, 3.5))
+
+plt.subplot(121)
+plt.plot(Ns, starry_time_quadratic * 1e3, ".-", label="starry", c="k")
+plt.plot(Ns, exoplanet_time_quadratic * 1e3, ".-", label="exoplanet", c="C4")
+plt.plot(Ns, jax_times_quadratic * 1e3, ".-", label="jaxoplanet (CPU)", c="C0")
+plt.plot(Ns, jax_times_quadratic_gpu * 1e3, ".-", label="jaxoplanet (GPU)", c="C3")
+plt.legend()
+plt.xlabel("number of points")
+plt.ylabel("time (ms)")
+plt.title("Quadratic limb-darkened map ($l=2$)")
+plt.yscale("log")
+plt.xscale("log")
+
+plt.subplot(122)
+plt.plot(Ns, starry_time_l20 * 1e3, ".-", label="starry", c="k")
+plt.plot(Ns, jax_times_l20 * 1e3, ".-", label="jaxoplanet (CPU)", c="C0")
+plt.plot(Ns, jax_times_l20_gpu * 1e3, ".-", label="jaxoplanet (GPU)", c="C3")
+plt.legend()
[…]

### 2024-09-23 latex/paper.tex
feat: add equations + light curve starry models

@@
 
-\title{Hardware-Accelerated Computation of Stellar Light Curves}
+\title{\texttt{jaxoplanet}: Hardware-Accelerated Computation of Stellar Light Curves}
 
@@
 Transit and occultation light curves, for example, are used to characterize the orbital parameters of stellar systems, and infer the atmospheric properties of their transiting bodies.
-Phase curves, on the other hand, can be used to study the thermal emission of exoplanets, and the atmospheric properties of stars.
+Phase curves, on the other hand, can be used to study the thermal emission of exoplanets, and the atmospheric circulation of stars.
 In order to infer these properties, a fast, accurate and robust model of stellar light curves is needed, one that accounts for the non-uniform surface intensity of spherical bodies and their mutual occultations.
@@
 \newpage
-\section{starry formalism}\label{starry}
+\section{Light curve models}\label{starry}
+\subsection{Spherical harmonics}
+
+The surface of a spherical body can be naturally described by spherical harmonics: a set of special functions that can be used as orthonormal basis to represent any spherical surface map.
+If $\bvec{\tilde y}$ denotes the spherical harmonics basis of degree $l_{max}$ (defined in \citealt[section X]{starry}) and $\bvec{y}$ a vector that contains the coefficients describing the surface in this basis, the intensity $I$ of the map projected onto the sky-projected $(x, y)$ plane of the observer is
+$$I(x, y) = \bvec{\tilde y}^\mathsf{T}(x, y) \, \bvec{y}.$$
+In order to orient this surface, for example with a given inclination, obliquity and phase around its rotation axis, we define a rotation matrix $\bvec{R}$ that acts on the spherical harmonics coefficients $\bvec{y}$, such that the intensity of the map projected onto the sky of the observer is
+$$I(x, y) = \bvec{\tilde y}^\mathsf{T}(x, y) \, \bvec{R} \, \bvec{y}.$$\\
+\subsection{Rotation light curve}
+Using this expression, the integrated intensity of the rotated map is
+$$F = \oiint I(x,y)\,dS = \oiint \bvec{\tilde y}^\mathsf{T}\bvec{R}\bvec{y}\,dS$$
+In order to simplify this integral, \citealt{starry} introduce a change of basis from the spherical harmonics basis $\bvec{\tilde y}$ to a simple polynomial basis $\bvec{\tilde p}$ characterized by the change of basis matrix $\bvec{A_1}$.
+In this new basis, $\bvec{y}$ can be expressed as
+$$\bvec{p} = \bvec{A_1}\bvec{y}.$$
+Hence, the integrated intensity becomes
+$$F = \bvec{r^\mathsf{T}}\bvec{A_1}\bvec{R}\bvec{y},$$
+with
+$$\bvec{r^\mathsf{T}} \equiv \oiint \mathbf{\tilde{p}}^\top\,dS.$$
+As $\bvec{\tilde{p}}$ is simply made of polynomials of the coordinates $x$, $y$ and $z$, the elements of the \textit{rotation vector} $\bvec{r}$ can be computed analytically and only once for a given map degree $l_{max}$ (see \citealt[Eq.
+20]{starry}).\\\\
+\subsection{Occultation light curve}
+The same formalism can be used to compute the integrated intensity of a surface map occulted by another spherical body.
+In this case, the intensity of the map must be integrated only over the unocculted surface $S$ of the disk, such that
+$$ F = \oiint_S I(x,y)\,dS = \left(\oiint_S \bvec{\tilde p^\mathsf{T}}\,dS\right)\bvec{A_1}\bvec{R}\bvec{y}.$$
+To simplify this expression, \citealt{pal2012} note that any two-dimensional integral of the unocculted disk can be reduced to a one-dimensional integral over the contour $C$ of the occulting body using Green's theorem\footnote{Also known as Stokes' theorem in three-dimensional vector calculus.}.
+This can be done if the integrand of the integral over S can be described by a function $f$ such that
+$$f(x, y) = \frac{\delta G_{y}}{\delta x}-\frac{\delta G_{x}}{\delta y},$$
+where $G$ is a vector field defined and differentiable over $S$.\\\\
+By applying a last change of basis matrix $\bvec{A_2}$ to the polynomial basis $\bvec{\tilde p}$, \citealt{starry} transforms $\bvec{\tilde p}$ to a basis $\bvec{\tilde g}$ that satisfies this property, meaning that the surface integral of each of its components $\tilde{g_n}$ can be transformed to the line integral
+$$s_n \equiv \oiint_S \tilde{g_n}(x, y)\,dS = \oint_C \bvec{G}_n(x, y)\,d\bvec{r},$$
+with $\bvec{G}_n$ given in \citealt[Equation 34]{starry}.
+To simplify the bounds of this integral, the surface map can be aligned with the occulting body through a final rotation matrix $\bvec{R'}.$\\\\
+Using these expressions, the integrated intensity of the occulted map follows the idiomatic linear expression 
+$$F = \bvec{s^\mathsf{T}}\bvec{A}\,\bvec{R'}\,\bvec{R}\,\bvec{y},$$
+with $\bvec{A} = \bvec{A_2A_1}$ and the integrals $s_n$ arranged in the \textit{solution vector} $\bvec{s^\mathsf{T}}$.\\\\
+
+\subsection{Limb-darkened surfaces}
 
@@
 
+\begingroup\makeatletter\def\f@size{9}\check@mathfonts
+$$
+    \mbox{\normalsize$\mathcal{P}(\bvec{G}_n)$} =
+    \begin{dcases}
+        %
+        2(2r)^{l+2}\int\displaylimits_{-\kappa/2}^{\kappa/2}
+            (s_\varphi^2 -s_\varphi^4)^{\frac{\mu+4}{4}}
+            (\delta + s_\varphi^2)^{\frac{\nu}{2}}
+            \, \text{d}\varphi
+            %
+            & \qquad \frac{\mu}{2} \, \mathrm{even}
+        \\[1em]
+        %
+        \mathcal{F}\int\displaylimits_{-\kappa/2}^{\kappa/2}
+            (s_\varphi^2-s_\varphi^4)^{\tfrac{l-2}{2}}
+            (k^2- s_\varphi^2)^{\frac{3}{2}}
+            (1-2s_\varphi^2)
+            \, \text{d}\varphi
+            %
+            & \qquad \mu = 1, \,
+                     l \, \mathrm{even}
+        \\[1em]
+        %
+        \mathcal{F}\int\displaylimits_{-\kappa/2}^{\kappa/2}
+            (s_\varphi^2-s_\varphi^4)^{\tfrac{l-3}{2}}
+            (\delta + s_\varphi^2)
+            (k^2- s_\varphi^2)^{\frac{3}{2}}
+            (1-2 s_\varphi^2)
+            \, \text{d}\varphi
+            %
+            & \qquad \mu = 1, \, l \ne 1, \,
+                     l \, \mathrm{odd}
+        \\[1em]
+        %
+        2\mathcal{F}\int\displaylimits_{-\kappa/2}^{\kappa/2}
+            (s_\varphi^2-s_\varphi^4)^{\frac{\mu-1}{4}}
+            (\delta + s_\varphi^2)^{\frac{\nu-1}{2}}
+            (k^2 - s_\varphi^2)^{\frac{3}{2}}
+            \, \text{d}\varphi
+            & \qquad \frac{\mu-1}{2} \, \mathrm{even}, \, l \ne 1
+        \\[1em]
+        %
+        \int\displaylimits_{-\kappa/2}^{\kappa/2}
+        r(r-b\cos{2\varphi})\;\frac{1-z^3}{3(1-z^2)}\text{d}\varphi & \qquad \mu=1, \, l=1
+        \\[1em]
+        %
+        0 & \qquad \mathrm{otherwise},
+    \end{dcases}
+$$
+\endgroup
+
+\begingroup\makeatletter\def\f@size{9}\check@mathfonts
+$$
+    \mbox{\normalsize$s_n$} =
+    \begin{dcases}
+        %
+        \pi - \kappa_1 - r^2\kappa_0 + \textit{A}_{kite} & \qquad n = 0
+        \\[1em]
+        \int\displaylimits_{-\kappa_0/2}^{\kappa_0/2}
+        r(r-b\cos{2\varphi})\;\frac{1-z^3}{3(1-z^2)}\text{d}\varphi - \frac{2}{3}(\kappa_1 - \pi)& \qquad n = 1
+        \\[1em]
+        2s_0 + 4\pi\eta - 2\pi & \qquad n = 2
+        \\[1em]
+        2r(4br)^{\frac{n}{2}}
+        \int\displaylimits_{-\kappa_0/2}^{\kappa_0/2}
+        (k^2 - s_\varphi^2)^{\frac{n}{2}}
+        (r - b + 2bs_\varphi^2) \, \text{d}\varphi & \qquad \mathrm{n\ge3},
+    \end{dcases}
+$$
+\endgroup
+
 \newpage
@@
 Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s}, taken over the range of values $b$.
-For each order $n$ all limb-darkening coefficients are set to zero except $u_n=1$.
+For each order $n$ all limb-darkening coefficients are set to zero except the $n$-th coefficient $u_n=1$.
 \codelink{workflows/precision/scripts/plot_error_limbdark}}

### 2024-09-23 latex/paper.tex
feat: bits on the limb-darkened case.

@@
 where $G$ is a vector field defined and differentiable over $S$.\\\\
-By applying a last change of basis matrix $\bvec{A_2}$ to the polynomial basis $\bvec{\tilde p}$, \citealt{starry} transforms $\bvec{\tilde p}$ to a basis $\bvec{\tilde g}$ that satisfies this property, meaning that the surface integral of each of its components $\tilde{g_n}$ can be transformed to the line integral
+By applying a last change of basis matrix $\bvec{A_2}$ to the polynomial basis $\bvec{\tilde p}$, \citealt{starry} transforms $\bvec{\tilde p}$ to the \textit{Green's basis} $\bvec{\tilde g}$ that satisfies this property, meaning that the surface integral of each of its components $\tilde{g_n}$ can be transformed to the line integral
 $$s_n \equiv \oiint_S \tilde{g_n}(x, y)\,dS = \oint_C \bvec{G}_n(x, y)\,d\bvec{r},$$
@@
 \subsection{Limb-darkened surfaces}
+We notice that a simpler situation is encountered when the surface of the spherical body can be described by a radial intensity profile $I(r)$, such as a limb-darkened star.
+In this case, the vector of spherical harmonics coefficients $\bvec{y}$ is very sparse (because all $m\neq0$ components are zero) and the surface map doesn't need to be rotated.
+Contemporary to \cite{starry}, \cite{Agol2020} explore this case and provide simple expressions for the integrated intensity of a star whose surface can be described by a polynomial limb-darkening law occulted by an opaque body.
+In this case:
+\begin{enumerate}
+    \item The surface is simply described by $\bvec{u}$, a vector of polynomial limb-darkening coefficients.
+    \item The Green's basis takes a simpler form (see \citealt[Eq. X]{Agol2020}).
+    \item Matrices involved in the computation of the integrated flux have $l_{max}$ columns, versus $(l_{max}+1)^2$ in the general case, reducing drastically the computational cost of the model.
+\end{enumerate}
+For these reasons, and because of its wide application, our implementation treats radially symmetric maps as a separate case.\\\\
 

### 2024-09-24 latex/paper.tex
fix: some implementation details

--- 2024-08-13 latex/paper.tex
@@
 \section*{Introduction}
-
-\begin{equation}\label{eq:starry}
-    f = \bvec{s}^{\boldsymbol{\mathsf{T}}} \bvec{A} \, \bvec{R'} \, \bvec{R} \, \bvec{y}
-\end{equation}
 
@@
 \newpage
-\section{starry formalism}\label{starry}
+\section{Light curve models}\label{starry}
 
-\section{optimization}\label{optimization}
+Inspired by the idea of \cite{pal2012}, \cite{starry} and \cite{Agol2020} present a general framework to compute the light curves of spherical bodies with non-uniform surface intensities.
+This framework, known as \textit{starry}, allows for precise and efficient computation of the light curves of stars and exoplanets, and has been successfully applied to a wide variety of science cases.
+In this section, we describe the mathematical formalism behind the \textit{starry} model.
+
+\subsection{Spherical harmonics}
+
+The surface of a spherical body can be naturally described by spherical harmonics: a set of special functions defined on the sphere that can serve as orthonormal basis to represent any spherical surface map.
+If $\bvec{\tilde y}$ denotes the spherical harmonics basis of degree $l_{max}$ (defined in \citealt[section X]{starry}) and $\bvec{y}$ a vector that contains the coefficients describing the surface in this basis, the intensity $I$ of the map projected onto the sky-projected $(x, y)$ plane of the observer is
+\begin{equation}I(x, y) = \bvec{\tilde y^\mathsf{T}}(x, y) \, \bvec{y}.\end{equation}
+In order to orient this surface, for example with a given inclination, obliquity and phase around its rotation axis, we define a rotation matrix $\bvec{R}$ that acts on the spherical harmonics coefficients $\bvec{y}$, such that the intensity of the map projected onto the sky of the observer is
+\begin{equation}I(x, y) = \bvec{\tilde y}^\mathsf{T}(x, y) \, \bvec{R} \, \bvec{y}.\end{equation}\\
+\subsection{Rotation light curve}
+Using this expression, the integrated intensity of the rotated map is
+\begin{equation}F = \iint I(x,y)\,\text{d}S = \iint \bvec{\tilde y^\mathsf{T}}\bvec{R}\bvec{y}\,\text{d}S\end{equation}
+In order to simplify this integral, \citealt{starry} introduce a change of basis from the spherical harmonics basis $\bvec{\tilde y}$ to a simple polynomial basis $\bvec{\tilde p}$ characterized by the change of basis matrix $\bvec{A_1}$.
+In this new basis, $\bvec{y}$ can be expressed as
+\begin{equation}\bvec{p} = \bvec{A_1}\bvec{y},\end{equation}
+such that 
+\begin{equation}I(x, y) = \bvec{\tilde p^\mathsf{T}}(x, y)\,\bvec{A_1}\bvec{R}\bvec{y}.\end{equation}
+Hence,
+\begin{equation}F = \iint \bvec{\tilde p^\mathsf{T}}\bvec{A_1}\bvec{R}\bvec{y}\,\text{d}S = \bvec{r^\mathsf{T}}\bvec{A_1}\bvec{R}\bvec{y},\end{equation}
+with
+\begin{equation}\bvec{r^\mathsf{T}} \equiv \iint \mathbf{\tilde{p}}^\top\,\text{d}S,\end{equation}
+since $\bvec{A_1}$, $\bvec{R}$ and $\bvec{y}$ are independent of the coordinates on the sphere.\\\\
+As $\bvec{\tilde{p}}$ is simply made of polynomials of the coordinates $x$, $y$ and $z$, the elements of the \textit{rotation vector} $\bvec{r}$ can be computed analytically and only once for a given $l_{max}$ (see \citealt[Eq.
+20]{starry}).\\\\
+\subsection{Occultation light curve}
+The same formalism can be used to compute the integrated intensity of a surface map occulted by another spherical body.
+In this case, the intensity of the map must be integrated only over the unocculted surface $S$ of the disk, such that
+\begin{equation} F = \iint_S I(x,y)\,\text{d}S = \left(\iint_S \bvec{\tilde p^\mathsf{T}}\,\text{d}S\right)\bvec{A_1}\bvec{R}\bvec{y}.\end{equation}
+To simplify this expression, \citealt{pal2012} note that any two-dimensional integral of the unocculted disk can be reduced to a one-dimensional integral over the contour $C$ of the occulting body using Green's theorem\footnote{Also known as Stokes' theorem in three-dimensional vector calculus.}.
+This can be done if the integrand of the integral over S can be described by a function $f$ such that
+\begin{equation}f(x, y) = \frac{\delta G_{y}}{\delta x}-\frac{\delta G_{x}}{\delta y},\end{equation}
+where $G$ is a vector field defined and differentiable over $S$.\\\\
+By applying a last change of basis matrix $\bvec{A_2}$ to the polynomial basis $\bvec{\tilde p}$, \citealt{starry} transforms $\bvec{\tilde p}$ to the \textit{Green's basis} $\bm{\tilde g}$ that satisfies this property, meaning that the surface integral of each of its components $\tilde{g_n}$ can be transformed to the line integral
+\begin{equation}
+    \label{eq:sn_gn}
+    s_n \equiv \iint_S \tilde{g_n}(x, y)\,\text{d}S = \oint_C \bvec{G}_n(x, y)\,\text{d}\bvec{r},
+\end{equation}
+with $\bvec{G}_n$ given in \citealt[Equation 34]{starry}.
+These functions require the occulting body to be aligned with the $y$-axis of the reference frame, which can be achieved by applying a final rotation matrix $\bvec{R'}$ to the surface map.\\\\
+Using these expressions, the integrated intensity of the occulted map follows the idiomatic linear expression 
+\begin{equation}F = \bvec{s^\mathsf{T}}\bvec{A}\,\bvec{R'}\,\bvec{R}\,\bvec{y},\end{equation}
+with $\bvec{A} = \bvec{A_2A_1}$ and the integrals $s_n$ arranged in the \textit{solution vector} $\bvec{s^\mathsf{T}}$.\\\\
+
+\subsection{Limb-darkened surfaces}
+We notice that a simpler situation is encountered when the surface of the spherical body can be described by a radial intensity profile $I(r)$, such as a limb-darkened star.
+In this case, the vector of spherical harmonics coefficients $\bvec{y}$ is very sparse (because all $m\neq0$ components are zero) and the surface map doesn't need to be rotated.
+Contemporary to \cite{starry}, \cite{Agol2020} explore this case and provide simple expressions for the integrated intensity of a star whose surface can be described by a polynomial limb-darkening law occulted by an opaque body.
+In this case:
+\begin{enumerate}
+    \item The surface is simply described by $\bvec{u}$, a vector of polynomial limb-darkening coefficients.
+    \item The Green's basis takes a simpler form (see \citealt[Eq. X]{Agol2020}).
+    \item Matrices involved in the computation of the integrated flux have $l_{max}$ columns, versus $(l_{max}+1)^2$ in the general case, reducing drastically the computational cost of the model.
+\end{enumerate}
+For these reasons, and because of its wide application, our implementation treats radially-symmetric maps separately, leading to highly optimize computations of limb-darkened light curves.\\\\
+
+\newpage
+\section{Implementation details}\label{optimization}
 \subsection{Limb-darkening multiplicative maps}
@@
 
+In the general case, we described in \ref{model_occultation} how the integrated intensity of a spherical body occulted by another can be computed using the solution vector $\bvec{s}$.
+\cite{pal2012} shows that, knowing $\bvec{G}_n$, \autoref{eq:sn_gn} can be decomposed into
+\begin{equation}
+    s_n = \mathcal{Q}(\bvec{G}_n) - \mathcal{P}(\bvec{G}_n)
+\end{equation}
+where $\mathcal{Q}(\bvec{G}_n)$ and $\mathcal{P}(\bvec{G}_n)$ are line integrals over the contour of the occulting body parametrized by the angle $\varphi$ shown in \autoref{}\\\\
+
+\begin{itemize}
+    \item $\mathcal{Q}(\bvec{G}_n)$ is given by \citealt{starry} (Equation D28-29) in the general case, and is conveniently equal to $0$ for all $n > 0$ in the polynomial limb-darkening case.
+    \item $\mathcal{P}(\bvec{G}_n)$ is given by \citealt{starry} (Equation D32), for the general case, and by \citealt{Agol2020} (Equation 71) in the polynomial limb-darkening case.
+\end{itemize}
+$\mathcal{P}$ and $\mathcal{Q}$ are given by \citealt[Equation D32]{starry}, for the general case, and \citealt[Equation 71]{Agol2020}, in the polynomial limb-darkening case
+
+\begingroup\makeatletter\def\f@size{9}\check@mathfonts
+$$
+    \mbox{\normalsize$\mathcal{P}(\bvec{G}_n)$} =
+    \begin{dcases}
+        %
+        2(2r)^{l+2}\int\displaylimits_{-\kappa/2}^{\kappa/2}
+            (s_\varphi^2 -s_\varphi^4)^{\frac{\mu+4}{4}}
+            (\delta + s_\varphi^2)^{\frac{\nu}{2}}
+            \, \text{d}\varphi
+            %
+            & \qquad \frac{\mu}{2} \, \mathrm{even}
+        \\[1em]
+        %
+        \mathcal{F}\int\displaylimits_{-\kappa/2}^{\kappa/2}
+            (s_\varphi^2-s_\varphi^4)^{\tfrac{l-2}{2}}
+            (k^2- s_\varphi^2)^{\frac{3}{2}}
+            (1-2s_\varphi^2)
+            \, \text{d}\varphi
+            %
+            & \qquad \mu = 1, \,
+                     l \, \mathrm{even}
+        \\[1em]
+        %
+        \mathcal{F}\int\displaylimits_{-\kappa/2}^{\kappa/2}
+            (s_\varphi^2-s_\varphi^4)^{\tfrac{l-3}{2}}
+            (\delta + s_\varphi^2)
+            (k^2- s_\varphi^2)^{\frac{3}{2}}
+            (1-2 s_\varphi^2)
+            \, \text{d}\varphi
+            %
+            & \qquad \mu = 1, \, l \ne 1, \,
+                     l \, \mathrm{odd}
+        \\[1em]
+        %
+        2\mathcal{F}\int\displaylimits_{-\kappa/2}^{\kappa/2}
+            (s_\varphi^2-s_\varphi^4)^{\frac{\mu-1}{4}}
+            (\delta + s_\varphi^2)^{\frac{\nu-1}{2}}
+            (k^2 - s_\varphi^2)^{\frac{3}{2}}
+            \, \text{d}\varphi
+            & \qquad \frac{\mu-1}{2} \, \mathrm{even}, \, l \ne 1
+        \\[1em]
+        %
+        \int\displaylimits_{-\kappa/2}^{\kappa/2}
+        r(r-b\cos{2\varphi})\;\frac{1-z^3}{3(1-z^2)}\text{d}\varphi & \qquad \mu=1, \, l=1
+        \\[1em]
+        %
+        0 & \qquad \mathrm{otherwise},
+    \end{dcases}
+$$
+\endgroup
+
+\begingroup\makeatletter\def\f@size{9}\check@mathfonts
+$$
+    \mbox{\normalsize$s_n$} =
+    \begin{dcases}
+        %
+        \pi - \kappa_1 - r^2\kappa_0 + \textit{A}_{kite} & \qquad n = 0
+        \\[1em]
+        \int\displaylimits_{-\kappa_0/2}^{\kappa_0/2}
+        r(r-b\cos{2\varphi})\;\frac{1-z^3}{3(1-z^2)}\text{d}\varphi - \frac{2}{3}(\kappa_1 - \pi)& \qquad n = 1
+        \\[1em]
+        2s_0 + 4\pi\eta - 2\pi & \qquad n = 2
+        \\[1em]
+        2r(4br)^{\frac{n}{2}}
+        \int\displaylimits_{-\kappa_0/2}^{\kappa_0/2}
+        (k^2 - s_\varphi^2)^{\frac{n}{2}}
+        (r - b + 2bs_\varphi^2) \, \text{d}\varphi & \qquad \mathrm{n\ge3},
+    \end{dcases}
+$$
+\endgroup
+
 \newpage

### 2024-09-26 latex/paper.tex
feat: solution vector

@@
 20]{starry}).\\\\
-\subsection{Occultation light curve}
+\subsection{Occultation light curve}\label{occultation}
 The same formalism can be used to compute the integrated intensity of a surface map occulted by another spherical body.
@@
 \section{Implementation details}\label{optimization}
+
+\subsection{Limb-darkening normalization}
+In \textsf{starry} version 0.2.2 \citep{starry}, the normalization of a limb-darkened light curve was done by assuming that adding limb-darkening to a non-uniform surface does not modify its total luminosity.
+Hence, the total flux received from a star during its complete rotation should be the same with or without limb-darkening.
+In older versions of starry ($\le$0.2.2), this was done by actually computing the light curve of the star with and without limb darkening, and using the ratio of the two to normalize the limb-darkened light curve.
+As discussed in the \textsf{starry} documentation\footlink{https://starry.readthedocs.io/en/latest/notebooks/LDNormalization/} this approach is wrong and leads to an unphysical normalization.\\\\
+Instead, one can assume that the total flux of the star should be similar for all observers, independently of the surface map of the star.
+By noticing that all components of the spherical harmonics basis integrate to 0 (except the constant component $Y_{0,0}$) we deduce that the mean flux of the star is that of a static surface as seen by different observers, which corresponds to a purely limb-darkened surface (with no other source of non-uniformity).
+In the spherical harmonics basis, this surface can be represented by the vector
+\begin{equation}
+    \bvec{y_u} = \bvec{U}\bvec{u}
+\end{equation}
+where $\bvec{U}$ is the change of basis matrix from the polynomial limb-darkening coefficients to the spherical harmonics coefficients. The flux of this purely limb-darkened surface is
+\begin{equation}
+    f_u = \bvec{r}^\mathsf{T}\bvec{A_1}\bvec{y_u}
+\end{equation}
+which can serve as a normalization factor. In practice we use $\pi/f_u$ to be this factor, so that the total flux for a uniform map with no limb darkening is unchanged.
+
 \subsection{Limb-darkening multiplicative maps}
@@
 
-\subsection{Computation of the solution vectors}\label{solution_vectors}
+\newpage
+\subsection{Computation of the solution vectors}\label{solution_vector}
 
-In the general case, we described in \ref{model_occultation} how the integrated intensity of a spherical body occulted by another can be computed using the solution vector $\bvec{s}$.
-\cite{pal2012} shows that, knowing $\bvec{G}_n$, \autoref{eq:sn_gn} can be decomposed into
+We described in \autoref{occultation} how the integrated intensity of a spherical body occulted by another can be computed using the solution vector $\bvec{s}^\mathsf{T}$.
+\cite{pal2012} emphasizes that, knowing $\bvec{G}_n$, each component $s_n$ of the solution vector defined in \autoref{eq:sn_gn} can be decomposed into
 \begin{equation}
@@
 \end{equation}
-where $\mathcal{Q}(\bvec{G}_n)$ and $\mathcal{P}(\bvec{G}_n)$ are line integrals over the contour of the occulting body parametrized by the angle $\varphi$ shown in \autoref{}\\\\
+where $\mathcal{Q}(\bvec{G}_n)$ and $\mathcal{P}(\bvec{G}_n)$ are line integrals over the contours of the occulting body and the primary's surface, as shown in \autoref{fig:geometry}.
+%
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[width=\textwidth]{../workflows/geometry/figures/occultation_geometry.pdf}
+        \caption{Geometric configuration and angles definitions for the two-body occultation.
+On the left figure, the larger light grey disk corresponds to the primary body being occulted by the darker secondary body.
+On the right figure, the area $S$ on which the intensity is integrated is shown in light blue.
+By choosing an appropriate transform, the integration over $S$ can be reduced to a line integral over the contour of the occulting body, which can be decomposed into the contour of the occulting body within the area of the primary body (in cyan) and the contour of the primary body everywhere else (in blue).}
+        \label{fig:geometry}
+    \end{center}
+\end{figure}
+%
+In the general case, the $\mathcal{Q}$ integral takes a simple analytical form and can be computed using the recurrence relations from \cite{starry} (Equation D29).
+The case of a limb-darkened surface is even simpler as $\mathcal{Q}(\bvec{G}_n) = 0$ for all $n > 0$.
+On the other hand, the $\mathcal{P}$ integral is harder to evaluate and requires the use of recurrence relations and truncated series expansions, with a scheme that depends on the geometric configuration of the system (see \citealt{starry}, Section D.2).\\\\
+In order to translate the \textsf{starry}  implementation to \textsf{JAX}, we evaluate the $\mathcal{P}$ integrals numerically using Gauss-Legendre integrations, requiring the evaluation of the integrated functions only over a fixed number of points.\\\\
 
-\begin{itemize}
-    \item $\mathcal{Q}(\bvec{G}_n)$ is given by \citealt{starry} (Equation D28-29) in the general case, and is conveniently equal to $0$ for all $n > 0$ in the polynomial limb-darkening case.
-    \item $\mathcal{P}(\bvec{G}_n)$ is given by \citealt{starry} (Equation D32), for the general case, and by \citealt{Agol2020} (Equation 71) in the polynomial limb-darkening case.
-\end{itemize}
-$\mathcal{P}$ and $\mathcal{Q}$ are given by \citealt[Equation D32]{starry}, for the general case, and \citealt[Equation 71]{Agol2020}, in the polynomial limb-darkening case
+\subsubsection{Polynomial limb-darkening}
+For a surface described by a polynomial limb-darkening law of order $n$, the $\mathcal{P}$ integral is given in \cite{Agol2020} by
+\begingroup\makeatletter\def\f@size{9}\check@mathfonts
+$$
+    \mbox{\normalsize$\mathcal{P}(\bvec{G}_n)$} =
+    \begin{dcases}
+        %
+        \pi - \kappa_1 - r^2\kappa_0 + \textit{A}_{kite} & \qquad n = 0
+        \\[1em]
+        2s_0 + 4\pi\eta - 2\pi & \qquad n = 2
+        \\[1em]
+        2r(4br)^{\frac{n}{2}}
+        \int\displaylimits_{-\kappa_0/2}^{\kappa_0/2}
+        (k^2 - \sin^2{\varphi '})^{\frac{n}{2}}
+        (r - b + 2b\sin^2\varphi ') \, \text{d}\varphi ' & \qquad \mathrm{n\ge3},
+    \end{dcases}
+$$
+\endgroup
+where $\kappa_0$, $\kappa_1$ and $A_{kite}$ are the angles and area shown in \autoref{fig:geometry} and computed following \cite{Agol2020} (Equations 31 and 32).
+For $n=2$, $\eta$ is defined in \cite{Agol2020} (Equation 53) and depends on the elliptic parameter
+%
+\begin{equation}
+    k = \frac{1-r^2 - b^2 + 2br}{4br}.
+\end{equation}\\
+%
+For $n=1$, we use Equation 31, 32 and 34 from \cite{starry} to write
+%
+\begin{equation}
+    \mathcal{P}(\bvec{G}_1) = \int\displaylimits_{-\kappa_0/2}^{\kappa_0/2}
+        r(r-b\cos{2\varphi '})\;\frac{1-z^3}{3(1-z^2)}\text{d}\varphi ' \quad \text{and} \quad \mathcal{Q}(\bvec{G}_1) = \frac{2}{3}(\pi-\kappa_1),
+\end{equation}\\
+%
+where $1 - z^2 = b + r - 2\cos{\varphi '}$ \footnote{note that $\varphi '$ is different from the angle $\varphi$ shown in \autoref{fig:geometry} and was introduced through the change of variable $\varphi ' = \frac{1}{2}(\varphi -\frac{3\pi}{2})$.}.
+\subsubsection{Non-uniform surface}
+In the general case, for a star with a non-radially symmetric surface intensity, $n$ denotes the index of the spherical harmonics component such that
+\begin{equation}
+    n = l^2 + l + m,
+\end{equation}
+where $l$ is the degree of the spherical harmonic and $m$ its order. For each component, the expression of $\mathcal{P}(\bvec{G}_n)$ depends on $\mu$ and $\nu$ defined as
+\begin{equation}
+    \begin{aligned}
+    \mu &= l - m,\\
+    \nu &= l + m.
+    \end{aligned}
+\end{equation}
 
+The $\mathcal{P}$ integral is then given by Equation D32 from \cite{starry} as
+%
 \begingroup\makeatletter\def\f@size{9}\check@mathfonts
@@
         2(2r)^{l+2}\int\displaylimits_{-\kappa/2}^{\kappa/2}
-            (s_\varphi^2 -s_\varphi^4)^{\frac{\mu+4}{4}}
-            (\delta + s_\varphi^2)^{\frac{\nu}{2}}
-            \, \text{d}\varphi
+            (s_{\varphi '}^2 -s_{\varphi '}^4)^{\frac{\mu+4}{4}}
+            (\delta + s_{\varphi '}^2)^{\frac{\nu}{2}}
+            \, \text{d}\varphi '
             %
@@
         \mathcal{F}\int\displaylimits_{-\kappa/2}^{\kappa/2}
-            (s_\varphi^2-s_\varphi^4)^{\tfrac{l-2}{2}}
-            (k^2- s_\varphi^2)^{\frac{3}{2}}
-            (1-2s_\varphi^2)
-            \, \text{d}\varphi
+            (s_{\varphi '}^2-s_{\varphi '}^4)^{\tfrac{l-2}{2}}
+            (k^2- s_{\varphi '}^2)^{\frac{3}{2}}
+            (1-2s_{\varphi '}^2)
+            \, \text{d}\varphi '
             %
@@
         \mathcal{F}\int\displaylimits_{-\kappa/2}^{\kappa/2}
-            (s_\varphi^2-s_\varphi^4)^{\tfrac{l-3}{2}}
-            (\delta + s_\varphi^2)
-            (k^2- s_\varphi^2)^{\frac{3}{2}}
-            (1-2 s_\varphi^2)
-            \, \text{d}\varphi
+            (s_{\varphi '}^2-s_{\varphi '}^4)^{\tfrac{l-3}{2}}
+            (\delta + s_{\varphi '}^2)
+            (k^2- s_{\varphi '}^2)^{\frac{3}{2}}
+            (1-2 s_{\varphi '}^2)
+            \, \text{d}\varphi '
             %
@@
         2\mathcal{F}\int\displaylimits_{-\kappa/2}^{\kappa/2}
-            (s_\varphi^2-s_\varphi^4)^{\frac{\mu-1}{4}}
-            (\delta + s_\varphi^2)^{\frac{\nu-1}{2}}
-            (k^2 - s_\varphi^2)^{\frac{3}{2}}
-            \, \text{d}\varphi
+            (s_{\varphi '}^2-s_{\varphi '}^4)^{\frac{\mu-1}{4}}
+            (\delta + s_{\varphi '}^2)^{\frac{\nu-1}{2}}
+            (k^2 - s_{\varphi '}^2)^{\frac{3}{2}}
+            \, \text{d}\varphi '
             & \qquad \frac{\mu-1}{2} \, \mathrm{even}, \, l \ne 1
@@
         \int\displaylimits_{-\kappa/2}^{\kappa/2}
-        r(r-b\cos{2\varphi})\;\frac{1-z^3}{3(1-z^2)}\text{d}\varphi & \qquad \mu=1, \, l=1
+        r(r-b\cos{2\varphi '})\;\frac{1-z^3}{3(1-z^2)}\text{d}\varphi ' & \qquad \mu=1, \, l=1
         \\[1em]
@@
 \endgroup
-
-\begingroup\makeatletter\def\f@size{9}\check@mathfonts
-$$
-    \mbox{\normalsize$s_n$} =
-    \begin{dcases}
 %
-        \pi - \kappa_1 - r^2\kappa_0 + \textit{A}_{kite} & \qquad n = 0
-        \\[1em]
-        \int\displaylimits_{-\kappa_0/2}^{\kappa_0/2}
-        r(r-b\cos{2\varphi})\;\frac{1-z^3}{3(1-z^2)}\text{d}\varphi - \frac{2}{3}(\kappa_1 - \pi)& \qquad n = 1
-        \\[1em]
-        2s_0 + 4\pi\eta - 2\pi & \qquad n = 2
-        \\[1em]
-        2r(4br)^{\frac{n}{2}}
-        \int\displaylimits_{-\kappa_0/2}^{\kappa_0/2}
-        (k^2 - s_\varphi^2)^{\frac{n}{2}}
-        (r - b + 2bs_\varphi^2) \, \text{d}\varphi & \qquad \mathrm{n\ge3},
-    \end{dcases}
-$$
-\endgroup
+where $s_{\varphi '} = \sin{\varphi '}$, $\kappa = \varphi + \frac{\pi}{2}$ (see \citealt[Equation D31]{starry}) and $\delta$ is defined for numerical stability as
+\begin{equation}
+    \delta=\frac{b-r}{2r}.
+\end{equation}
+Note how, unlike in the \textsf{starry} implementation, the term $\mu=1$, $l=1$ is derived exactly like in the polynomial limb-darkening case, instead of analytically which would require the evaluation of elliptic integrals.\\\\
+Because of its numerical integration, the precision of the $\mathcal{P}$ integral is limited and depends on the order $q$ of the Gauss-Legendre quadrature, which corresponds to the number of point for which the integrand is evaluated.
+To minimize this number, we note that all integrands are symmetrical functions, except for $\mu=1$ and $l=1$.
+Hence, all integrals can be computed over the interval $[0, \kappa/2]$ and multiplied by 2.
+We also note that all terms for which $\frac{\mu}{2}$ is even reach machine precision for an integration order as low as $q=20$, that we hold fix.
+With these optimizations, we benchmark the precision and speed of our final implementation for the solution vector in \autoref{performance}.
 
 \newpage
-\section{Performance}\label{performances}
+\section{Performance}\label{performance}
 Although \textit{starry} occultation light curves have analytical expressions, their practical implementation relies on numerical approximations chosen to balance precision and computation time.

### 2024-09-27 latex/paper.tex
feat: redoing figures + limb darkened maps

@@
 In this case, the vector of spherical harmonics coefficients $\bvec{y}$ is very sparse (because all $m\neq0$ components are zero) and the surface map doesn't need to be rotated.
-Contemporary to \cite{starry}, \cite{Agol2020} explore this case and provide simple expressions for the integrated intensity of a star whose surface can be described by a polynomial limb-darkening law occulted by an opaque body.
+Contemporary to \cite{starry}, \cite{Agol2020} explore this case and provide simple expressions for the integrated intensity of a star whose surface can be described by a polynomial limb darkening law occulted by an opaque body.
 In this case:
 \begin{enumerate}
-    \item The surface is simply described by $\bvec{u}$, a vector of polynomial limb-darkening coefficients.
+    \item The surface is simply described by $\bvec{u}$, a vector of polynomial limb darkening coefficients.
     \item The Green's basis takes a simpler form (see \citealt[Eq. X]{Agol2020}).
@@
 
-\subsection{Limb-darkening normalization}
-In \textsf{starry} version 0.2.2 \citep{starry}, the normalization of a limb-darkened light curve was done by assuming that adding limb-darkening to a non-uniform surface does not modify its total luminosity.
-Hence, the total flux received from a star during its complete rotation should be the same with or without limb-darkening.
+\subsection{Limb darkened maps}
+Limb darkened maps are obtained by multiplying a base map, devoid of limb darkening, with a pure limb darkened map.
+If $\bvec{U}$ is the change of basis matrix from the polynomial limb darkening coefficients $\bvec{u}$ to the spherical harmonics basis, the pure limb darkened map is
+\begin{equation}
+    \label{eq:limb_darkening_map}
+    \bvec{y}_u = \bvec{U}\bvec{u}.
+\end{equation}
+The multiplication of the base map represented by $\bvec{y}$ with the limb darkening map represented by $\bvec{y}_u$ can be seen as the multiplication of two polynomial functions, only with special coefficients (called Clebsch–Gordan coefficients) inherent to the properties of spherical harmonics.
+To simplify this operation, we transform both maps into the polynomial basis, so that their product is the actual product of two polynomials.
+
+\subsection{Limb darkening normalization}
+In \textsf{starry} version 0.2.2 \citep{starry}, the normalization of a limb-darkened light curve was done by assuming that adding limb darkening to a non-uniform surface does not modify its total luminosity.
+Hence, the total flux received from a star during its complete rotation should be the same with or without limb darkening.
 In older versions of starry ($\le$0.2.2), this was done by actually computing the light curve of the star with and without limb darkening, and using the ratio of the two to normalize the limb-darkened light curve.
@@
 Instead, one can assume that the total flux of the star should be similar for all observers, independently of the surface map of the star.
-By noticing that all components of the spherical harmonics basis integrate to 0 (except the constant component $Y_{0,0}$) we deduce that the mean flux of the star is that of a static surface as seen by different observers, which corresponds to a purely limb-darkened surface (with no other source of non-uniformity).
-In the spherical harmonics basis, this surface can be represented by the vector
+By noticing that all components of the spherical harmonics basis integrate to 0 (except the constant component $Y_{0,0}$) we deduce that the mean flux of the star must be coming from the surface feature seen by any observer, no matter his viewpoint, which corresponds to a purely limb-darkened surface (with no other source of non-uniformity).\\\\
+The flux of this purely limb-darkened surface is
 \begin{equation}
-    \bvec{y_u} = \bvec{U}\bvec{u}
+    f_u = \bvec{r}^\mathsf{T}\bvec{A_1}\bvec{y}_u
 \end{equation}
-where $\bvec{U}$ is the change of basis matrix from the polynomial limb-darkening coefficients to the spherical harmonics coefficients. The flux of this purely limb-darkened surface is
-\begin{equation}
-    f_u = \bvec{r}^\mathsf{T}\bvec{A_1}\bvec{y_u}
-\end{equation}
-which can serve as a normalization factor. In practice we use $\pi/f_u$ to be this factor, so that the total flux for a uniform map with no limb darkening is unchanged.
+with $\bvec{y}_u$ defined in \autoref{eq:limb_darkening_map}. Hence, we normalize limb-darkened maps by $\pi/f_u$, so that the total flux for a uniform map with no limb darkening is unchanged.
 
-\subsection{Limb-darkening multiplicative maps}
 \subsection{Spherical harmonics rotations}
@@
 
-\newpage
 \subsection{Computation of the solution vectors}\label{solution_vector}
@@
 
-\subsubsection{Polynomial limb-darkening}
-For a surface described by a polynomial limb-darkening law of order $n$, the $\mathcal{P}$ integral is given in \cite{Agol2020} by
+\subsubsection{Polynomial limb darkening}
+For a surface described by a polynomial limb darkening law of order $n$, the $\mathcal{P}$ integral is given in \cite{Agol2020} by
 \begingroup\makeatletter\def\f@size{9}\check@mathfonts
@@
     \begin{aligned}
-    \mu &= l - m,\\
-    \nu &= l + m.
+    \mu &\equiv l - m,\\
+    \nu &\equiv l + m.
     \end{aligned}
@@
 \end{equation}
-Note how, unlike in the \textsf{starry} implementation, the term $\mu=1$, $l=1$ is derived exactly like in the polynomial limb-darkening case, instead of analytically which would require the evaluation of elliptic integrals.\\\\
+Note how, unlike in the \textsf{starry} implementation, the term $\mu=1$, $l=1$ is derived exactly like in the polynomial limb darkening case, instead of analytically which would require the evaluation of elliptic integrals.\\\\
 Because of its numerical integration, the precision of the $\mathcal{P}$ integral is limited and depends on the order $q$ of the Gauss-Legendre quadrature, which corresponds to the number of point for which the integrand is evaluated.
 To minimize this number, we note that all integrands are symmetrical functions, except for $\mu=1$ and $l=1$.
-Hence, all integrals can be computed over the interval $[0, \kappa/2]$ and multiplied by 2.
+Hence, all integrals can be computed over the interval $\left[0, \frac{\kappa}{2}\right]$, instead of $\left[-\frac{\kappa}{2}, \frac{\kappa}{2}\right]$, and multiplied by 2.
 We also note that all terms for which $\frac{\mu}{2}$ is even reach machine precision for an integration order as low as $q=20$, that we hold fix.
@@
 \end{figure}
-For reference, \autoref{fig:precision_limbdark} shows the relative error in the flux of a limb-darkened star occulted by an opaque body for increasing orders of a polynomial limb-darkening law.
-In this figure, we also report the error of the flux computed using the \textsf{exoplanet} Python package (limited to linear and quadratic limb-darkening).
+For reference, \autoref{fig:precision_limbdark} shows the relative error in the flux of a limb-darkened star occulted by an opaque body for increasing orders of a polynomial limb darkening law.
+In this figure, we also report the error of the flux computed using the \textsf{exoplanet} Python package (limited to linear and quadratic limb darkening).
 \begin{figure}[H]
@@
         \includegraphics[width=\textwidth]{../workflows/precision/figures/limbdark_error.PDF}
-        \caption{Maximum errors in the overall flux $f$ of a star occulted by an opaque body for increasing orders of its polynomial limb-darkening law.
+        \caption{Maximum errors in the overall flux $f$ of a star occulted by an opaque body for increasing orders of its polynomial limb darkening law.
 Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s}, taken over the range of values $b$.
-For each order $n$ all limb-darkening coefficients are set to zero except the $n$-th coefficient $u_n=1$.
+For each order $n$ all limb darkening coefficients are set to zero except the $n$-th coefficient $u_n=1$.
 \codelink{workflows/precision/scripts/plot_error_limbdark}}
@@
 \end{figure}
-A set of unitary polynomial limb-darkening coefficients of order 20 leads to spherical harmonics coefficients as large as $10^6$.
-This explains the larger errors observed in \autoref{fig:precision_limbdark} (error versus limb-darkening order) compared to the ones shown in \autoref{fig:precision_degree} (error versus degree of the spherical harmonics).
+A set of unitary polynomial limb darkening coefficients of order 20 leads to spherical harmonics coefficients as large as $10^6$.
+This explains the larger errors observed in \autoref{fig:precision_limbdark} (error versus limb darkening order) compared to the ones shown in \autoref{fig:precision_degree} (error versus degree of the spherical harmonics).
 We not that the errors scale with the values of $u$.\\\\

### 2024-09-27 workflows/precision/scripts/jax_f_nonuniform_limbdark.py

@@ -24,18 +24,10 @@ def u(deg):
         return u
 
 
-# function = jax.vmap(
-#     lambda deg, b: surface_light_curve(
-#         Surface(u=u(deg)), y=b, z=10.0, r=r, order=order
-#     ),
-#     (None, 0),
-# )
-
-# USING LIMB-DARK
-from jaxoplanet.core.limb_dark import light_curve
-
 function = jax.vmap(
-    lambda deg, b: light_curve(u(deg), b, r, order=100) + 1,
+    lambda deg, b: surface_light_curve(
+        Surface(u=u(deg)), y=b, z=10.0, r=r, order=order
+    ),
     (None, 0),
 )
 

### 2024-09-27 workflows/precision/scripts/make_b.py

@@ -1,7 +1,7 @@
 import numpy as np
 
 
-def b_range(r, logdelta=-3, logeps=-6, res=2):
+def b_range(r, logdelta=-3, logeps=-6, res=50):
     delta = 10**logdelta
     eps = 10**logeps
     if r > 1:

### 2024-09-27 workflows/precision/scripts/num_f_limbdark.py

@@ -1,8 +1,8 @@
 import numpy as np
 import pickle
-from jaxoplanet.experimental.starry.mpcore.utils import mp
+from jaxoplanet.experimental.starry.multiprecision.utils import mp
 from jaxoplanet.experimental.starry import basis
-from jaxoplanet.experimental.starry.mpcore import solution
+from jaxoplanet.experimental.starry.multiprecision import solution
 from tqdm import tqdm
 
 # some params
@@ -28,12 +28,8 @@ bs = np.load(snakemake.input.b, allow_pickle=True)["bs"]
 b = np.hstack(bs)
 
 # load A1
-print("inverting A1")
 num_matrices = np.load(snakemake.input.matrices, allow_pickle=True)
-
-# TODO: transfer that into the saved matrices
-# not done cause everything will rerun...
-A1inv = num_matrices["A1"] ** -1
+A1inv = num_matrices[degree]["A1inv"]
 
 
 # the function to compute the flux

### 2024-09-27 workflows/precision/scripts/num_fs_matrices.py

@@ -24,6 +24,9 @@ _ = get_R("R_mx", l_max, cache=matrices)
 _ = get_R("R_inc", l_max, inc=inc, obl=obl, cache=matrices)
 _ = get_R("R_obl", l_max, inc=inc, obl=obl, cache=matrices)
 
+print("Computing A1inv...")
+matrices[l_max]["A1inv"] = matrices[l_max]["A1"] ** -1
+
 import pickle
 
 with open(snakemake.output[0], "wb") as f:

### 2024-09-27 workflows/precision/scripts/plot_error_limbdark.py

@@ -1,18 +1,20 @@
 import matplotlib.pyplot as plt
 import numpy as np
-from jaxoplanet.experimental.starry.mpcore import mp, utils
+from jaxoplanet.experimental.starry.multiprecision import mp, utils
 
 degree = snakemake.params.degree
 
 data = {
     "small": {
         "jax": np.load(snakemake.input.jax[0], allow_pickle=True)["f"],
+        "jax_nu": np.load(snakemake.input.jax_nu[0], allow_pickle=True)["f"],
         "num": np.load(snakemake.input.num[0], allow_pickle=True)["f"],
         "starry": np.load(snakemake.input.starry[0], allow_pickle=True)["f"],
         "exo": np.load(snakemake.input.exo[0], allow_pickle=True)["f"],
     },
     "large": {
         "jax": np.load(snakemake.input.jax[1], allow_pickle=True)["f"],
+        "jax_nu": np.load(snakemake.input.jax_nu[0], allow_pickle=True)["f"],
         "num": np.load(snakemake.input.num[1], allow_pickle=True)["f"],
         "starry": np.load(snakemake.input.starry[1], allow_pickle=True)["f"],
         "exo": np.load(snakemake.input.exo[1], allow_pickle=True)["f"],
@@ -27,11 +29,13 @@ def errors(M1, M2):
 err = {
     "small": {
         "jax": errors(data["small"]["jax"], data["small"]["num"]).max(1)[1:],
+        "jax_nu": errors(data["small"]["jax_nu"], data["small"]["num"]).max(1)[1:],
         "starry": errors(data["small"]["starry"], data["small"]["num"]).max(1)[1:],
         "exo": errors(data["small"]["exo"], data["small"]["num"][0:3, :]).max(1)[1:],
     },
     "large": {
         "jax": errors(data["large"]["jax"], data["large"]["num"]).max(1)[1:],
+        "jax_nu": errors(data["small"]["jax_nu"], data["small"]["num"]).max(1)[1:],
         "starry": errors(data["large"]["starry"], data["large"]["num"]).max(1)[1:],
         "exo": errors(data["large"]["exo"], data["large"]["num"][0:3, :]).max(1)[1:],
     },
@@ -49,7 +53,8 @@ x = np.arange(0, degree + 2, 2)
[…]

### 2024-10-08 latex/paper.tex
add intro + other workflow stuff

@@
 
-\title{\texttt{jaxoplanet}: Hardware-Accelerated Computation of Stellar Light Curves}
+\title{Hardware-Accelerated Computation of Stellar Light Curves}
 
@@
 We describe the changes made to the original version of \textsf{starry} and evaluate the performance of its new implementation that we release in the open-source Python package \textsf{jaxoplanet}.
-Through this implementation, we provide the community with a differentiable model of stellar light curves, and enable the use of powerful machine-learning tools available in the \texttt{jax} ecosystem.
+Through this implementation, we provide the community with a differentiable model of stellar light curves, and enable the use of powerful machine-learning tools available in the \texttt{JAX} ecosystem.
 \gitlink{}
@@
 
-\section*{Introduction}
+\section{Introduction}
+
+During millenia, planets of the solar system were mostly known and studied through their orbits.
+Starting in the 16th century, Earth-based telescopes progressively enabled the direct imaging of their surface.
+But despite these new instruments, some minor planets were too small and too distant to be imaged, and remained unresolved until the 20th century, when more capable space-based observatories were launched.
+This is the case of Pluto, for example, whose surface was first studied through the light it reflects during its revolution and orbit around the Sun.
+Using point source light curves obtained from 1945 to 1986, \cite{Drish1995} tentatively reconstructed the map of the planet, only later compared to the first \textit{Hubble Space Telescope} low-resolution images obtained by \cite{Stern1997}.
+This mapping effort was gloriously completed with the extremely detailed pictures obtained during the close fly-by of the \textit{New Horizons} spacrecraft \citep{}.\\\\
+The study of extrasolar planets is taking a path similar to that of Pluto, starting with the study of their unresolved light curves \citep{} while direct imaging instruments are being developed \citep{}.
+The same applies to the study of unresolved stellar atmospheres, whose properties are inferred from the light they emit.
+In this context, developments of accurate and robust models of extrasolar systems light curves are crucial.
+As these models are being compared to large datasets in probabilistic ways, their posterior likelihood must be evaluated millions of times, which require them to be fast.
+This inference can be highly facilitated by the advent of gradient-based optimization and inference tools, so that the differentiability of these models offer a clear advantage.\\\\
+In this context, \cite{starry} presented an algorithm to analytically compute the light curves of spherical bodies with non-uniform surface intensities, particularly optimized for occultation light curves.
+This algorithm, known as \textit{starry}, has been successfully applied to a wide variety of science cases, from the study of exoplanets atmospheres \citep{} to the study of stellar photospheres \citep{}.
+However, as datasets and model complexity grow, the tractable inference of these parameters becomes increasingly challenging, and often motivates biased approximations.
+This is case, for example, in transmission spectroscopy measurements conducted with the \textit{James Webb Space Telescope}, for which ...
+\todo{(So, give an example of large number of spectral bins and required approximation) (Hatori et al.
+in prep)}.\\\\
+In this paper, we present a new implementation of the \textit{starry} framework in \texttt{JAX}, a high-performance machine-learning library designed for large-scale computations.
+We start in \autoref{starry} with a summarry of the formalism introduced by \cite{starry}.
+In \autoref{optimization}, we describe the changes made to the original version of \textit{starry}.
+Finally, in \autoref{performance}, we evaluate the performance of our new implementation that we release in the open-source Python package \textsf{jaxoplanet}.
+Through this implementation, we provide the community with a differentiable model of stellar light curves, and enable the use of powerful machine-learning tools available in the \texttt{JAX} ecosystem.\\\\
 
@@
 since $\bvec{A_1}$, $\bvec{R}$ and $\bvec{y}$ are independent of the coordinates on the sphere.\\\\
-As $\bvec{\tilde{p}}$ is simply made of polynomials of the coordinates $x$, $y$ and $z$, the elements of the \textit{rotation vector} $\bvec{r}$ can be computed analytically and only once for a given $l_{max}$ (see \citealt[Eq.
+As $\bvec{\tilde{p}}$ is simply made of polynomials of the coordinates $x$, $y$ and $z$, the elements of the \textit{rotation vector} $\bvec{r}^\mathsf{T}$ can be computed analytically and only once for a given $l_{max}$ (see \citealt[Eq.
 20]{starry}).\\\\
@@
 Using these expressions, the integrated intensity of the occulted map follows the idiomatic linear expression 
-\begin{equation}F = \bvec{s^\mathsf{T}}\bvec{A}\,\bvec{R'}\,\bvec{R}\,\bvec{y},\end{equation}
+\begin{equation}
+    \label{eq:starry}
+    F = \bvec{s^\mathsf{T}}\bvec{A}\,\bvec{R'}\,\bvec{R}\,\bvec{y},
+\end{equation}
 with $\bvec{A} = \bvec{A_2A_1}$ and the integrals $s_n$ arranged in the \textit{solution vector} $\bvec{s^\mathsf{T}}$.\\\\
@@
 Limb darkened maps are obtained by multiplying a base map, devoid of limb darkening, with a pure limb darkened map.
-If $\bvec{U}$ is the change of basis matrix from the polynomial limb darkening coefficients $\bvec{u}$ to the spherical harmonics basis, the pure limb darkened map is
+If $\bvec{U}$ is the change of basis matrix that transforms the polynomial limb darkening coefficients $\bvec{u}$ to the spherical harmonics coefficients $\bvec{y}_u$, the pure limb darkened map is
 \begin{equation}
@@
 \end{equation}
-The multiplication of the base map represented by $\bvec{y}$ with the limb darkening map represented by $\bvec{y}_u$ can be seen as the multiplication of two polynomial functions, only with special coefficients (called Clebsch–Gordan coefficients) inherent to the properties of spherical harmonics.
-To simplify this operation, we transform both maps into the polynomial basis, so that their product is the actual product of two polynomials.
+The multiplication of the base map represented by $\bvec{y}$ with the limb darkening map represented by $\bvec{y}_u$ can be seen as the multiplication of two polynomials, only with special coefficients (called Clebsch–Gordan coefficients) inherent to the properties of spherical harmonics.
+To simplify this operation, we transform both maps into the polynomial basis, so that their product is the actual product of two polynomials which doesn't need the computation of special coefficients.
 
@@
 Hence, the total flux received from a star during its complete rotation should be the same with or without limb darkening.
-In older versions of starry ($\le$0.2.2), this was done by actually computing the light curve of the star with and without limb darkening, and using the ratio of the two to normalize the limb-darkened light curve.
+In older versions of starry ($\le$0.2.2), this was done by actually computing the light curve of the star with and without limb darkening, and using the ratio of the two to normalize the limb darkened light curve.
 As discussed in the \textsf{starry} documentation\footlink{https://starry.readthedocs.io/en/latest/notebooks/LDNormalization/} this approach is wrong and leads to an unphysical normalization.\\\\
 Instead, one can assume that the total flux of the star should be similar for all observers, independently of the surface map of the star.
-By noticing that all components of the spherical harmonics basis integrate to 0 (except the constant component $Y_{0,0}$) we deduce that the mean flux of the star must be coming from the surface feature seen by any observer, no matter his viewpoint, which corresponds to a purely limb-darkened surface (with no other source of non-uniformity).\\\\
+In addition to the fact that all components of the spherical harmonics basis integrate to 0 (except the constant component $Y_{0,0}$) we deduce that the mean flux of the star must be coming from a static surface seen by any observer, no matter his viewpoint, which corresponds to a purely limb-darkened surface (with no other source of non-uniformity).\\\\
 The flux of this purely limb-darkened surface is
@@
 
-With \textit{starry}, every light curve computation at a given time $t$ involves a rotation of the spherical harmonics basis, from the rest-frame of the star (\autoref{fig:rotation_basis}, left) to its sky-projected orientation (\autoref{fig:rotation_basis}, right), with a final rotation relative to the position of the occulting body (see Figure 2.
+As shown in \autoref{starry}, computing the flux of the surface at a given time $t$ involves a rotation of the spherical harmonics basis, from the rest-frame of the star (\autoref{fig:rotation_basis}, left) to its sky-projected orientation (\autoref{fig:rotation_basis}, right), with a final rotation relative to the position of the occulting body (see Figure 2.
 from \citealt{starry}).
@@
         \includegraphics[width=0.6\textwidth]{../workflows/rotations/figures/rotation_basis.pdf}
-        \caption{Rotation of the spherical harmonics basis to place the star in the sky-projected frame of the observer.}
+        \caption{Rotation of a primary body's surface to place it in the sky-projected frame of the observer.}
         \label{fig:rotation_basis}
@@
 The rotation of spherical harmonics involves the computation of Wigner-D-matrices, commonly obtained through robust recursion relations in both degree $l$ and order $m$, but leading to costly computations.
-Hence, knowing how to decompose the full spherical harmonics rotation using pre-computed rotations is essential to achieve optimal performance in the \textit{starry} framework.
-\autoref{fig:rotation_starry} shows the 6 elementary rotations currently used in the \textsf{starry} implementation.
+Hence, knowing how to decompose the full spherical harmonics rotation using pre-computed rotations is essential to achieve optimal performance in this framework.
+\autoref{fig:rotation_starry} shows the 6 elementary rotations currently used in \textsf{starry}.
 \begin{figure}[H]
@@
         \includegraphics[width=\textwidth]{../workflows/rotations/figures/rotation_starry.pdf}
-        \caption{Consecutive rotations of the spherical harmonics basis in the \textsf{starry} implementation.
+        \caption{Decomposition of the full rotation of the surface into elementary rotations.
 Each schematic shows a sphere rotated from a previous orientation around an axis displayed as a black vector and an angle shown on top.}
@@
 \end{figure}
-In this figure, rotations 1 to 3 correspond to a rotation of the star around its rotation axis, and rotations 3 to 5 place the star on a sky-projected frame.
-Finally, rotation 6 aligns the spherical harmonics map to the occulting body on the $y$ axis in order to apply Green's theorem in the appropriate basis (Figure 2 from \citealt{starry}).
+In this figure, the first 3 rotations have the effect to rotate the star around its rotation axis, and rotations 3 to 5 place the star on a sky-projected frame.
+Finally, rotation 6 aligns the spherical harmonics map to the occulting body on the $y$ axis in order to apply Green's theorem in the appropriate basis (Figure 2 from \citealt{starry}).\\\\
 The first thing to notice is that rotations 1 to 3 could be simply reduced to a rotation of angle $\theta$ about the y-axis, turning steps 1 to 3 into one.
@@
 \end{equation}
-Note how, unlike in the \textsf{starry} implementation, the term $\mu=1$, $l=1$ is derived exactly like in the polynomial limb darkening case, instead of analytically which would require the evaluation of elliptic integrals.\\\\
+Note how, unlike in the \textsf{starry} implementation, the term $\mu=1$, $l=1$ is derived exactly like in the polynomial limb darkening case, instead of analytically which would require the evaluation of elliptic integrals (see \citealt{starry}, Section D.2.3).\\\\
 Because of its numerical integration, the precision of the $\mathcal{P}$ integral is limited and depends on the order $q$ of the Gauss-Legendre quadrature, which corresponds to the number of point for which the integrand is evaluated.
@@
 \section{Performance}\label{performance}
-Although \textit{starry} occultation light curves have analytical expressions, their practical implementation relies on numerical approximations chosen to balance precision and computation time.
-This is the case, for example, of the solution vector $\bvec{s}$, computed using numerical series expansions in \textsf{starry} (\citealt[section D.2.3]{starry}) and numerical integration in \textsf{jaxoplanet} (see \autoref{solution_vectors}).
+Although occultation light curves have analytical expressions, their practical implementation relies on numerical approximations chosen to balance precision and computation time.
+This is the case, for example, of the solution vector $\bvec{s}^\mathsf{T}$, computed using numerical series expansions in \textsf{starry} (\citealt[section D.2.3]{starry}) and numerical integration in \textsf{jaxoplanet} (see \autoref{solution_vector}).
 Other expressions, such as the Wigner-D matrices used in the rotation of the spherical harmonics basis, are computed using reccurence relations prone to the accumulation of numerical errors and instability.\\\\
@@
 We evaluate the precision of the \textsf{starry} and \textsf{jaxoplanet} terms against quantities computed at arbitrary precision, using a ground truth version of the code implemented with the \texttt{mpmath} Python library\footlink{https://mpmath.org/}.
-The precision of the overall flux $f$ can only be understood by considering the precision of the terms involved in \autoref{eq:starry}: the solution vector $\bvec{s}$ ($\bvec{r}$ if no occultation), the basis matrices $\bvec{A}$ and $\bvec{B}$, and the Wigner-D rotation tensor $\bvec{R}$.
-Although we show in \autoref{fig:precision_SAR} that the precision of the \textsf{starry} and \textsf{jaxoplanet} implentations differ for many of these terms, we focus this section on the solution vector $\bvec{s}$ and the overall flux $f$.\\\\
-Assuming that we perform the numerical integration of the solution vector $\bvec{s}$ at a relatively high order ($n=500$; see \autoref{solution_vectors}), \autoref{fig:precision_s} shows that our reimplentation of \textsf{starry} reaches relative errors lower than 1 part per billion for $\bvec{s}$ and the overall flux $f$, considering both a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$.
+The precision of the overall flux $f$ can only be understood by considering the precision of the terms involved in \autoref{eq:starry}: the solution vector $\bvec{s}^\mathsf{T}$ ($\bvec{r}^\mathsf{T}$ if no occultation), the change of basis matrices $\bvec{A_1}$ and $\bvec{A_2}$, and the Wigner-D rotation matrix $\bvec{R}$.
+Although we show in \autoref{fig:precision_SAR} that the precision of the \textsf{starry} and \textsf{jaxoplanet} implentations slightly differ for many of these terms, we focus this section on the solution vector $\bvec{s}^\mathsf{T}$ and the overall flux $f$.\\\\
+Assuming that we perform the numerical integration described in \autoref{solution_vector} at a relatively high order ($q=500$; see \autoref{solution_vector}), \autoref{fig:precision_s} shows that our reimplentation of \textsf{starry} reaches relative errors lower than 1 part per billion for $\bvec{s}^\mathsf{T}$ and the overall flux $f$, considering both a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$.
 For reference, \autoref{fig:precision_s_starry} shows comparable errors on the same quantities computed with the C++ implentation of \textsf{starry}.
 In addition, \autoref{fig:precision_degree} shows the evolution of the error on $f$ for increasing degrees of the spherical harmonics, both for \textsf{starry} and \textsf{jaxoplanet}.
-With these results, we validate the precision of \textsf{jaxoplanet} ans show it is on per with the legacy C++ \textsf{starry} implentation\footnote{version 1.2.0 \citep{starry_120}} described in \cite{starry}.\\\\
+We note that the coefficients associated to spherical harmonics of degrees $l$ usually take values of the order of magnitude $10^{-l}$.
+Given that the numerical errors in computing \autoref{eq:starry} scale lineraly with the values of the coefficients $\bvec{y}$, the results shown in \autoref{fig:precision_degree}, for spherical harmonics coefficients equal to 1, translate to machine precision in most practical applications.
+With these results, we validate the precision of \textsf{jaxoplanet} and show it is on per with the legacy C++ \textsf{starry} implentation\footnote{version 1.2.0 \citep{starry_120}} described in \cite{starry}.\\\\
 \begin{figure}[H]
@@
         \includegraphics[width=\textwidth]{../workflows/precision/figures/error_jax.pdf}
-        \caption{Error of the solution vector $\bvec{s}$ and the overall flux $f$ for all terms of the spherical harmonics basis up to degree $l=20$.
-Errors are computed against arbitrary-precision computations of $\bvec{s}$ and $f$ for a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$.
-For each $(l, m)$ basis component all coefficients are set to zero except $y_{l,m}=1$ and errors are scaled to the highest values of $\bvec{s}$ (respectively $f$) over $b$.
+        \caption{Error of the solution vector $\bvec{s}^\mathsf{T}$ and the overall flux $f$ for all terms of the spherical harmonics basis up to degree $l=20$.
+Errors are computed against arbitrary-precision computations of the solution vector and the flux for a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$.
+For each basis component $i$, all coefficients are set to zero except $y_i=1$ and errors are scaled to the highest values of $\bvec{s}$ (respectively $f$) over $b$.
 Then, the plotted errors for each degree $l$ are the maximum of the scaled errors over all $m\in [-l, l]$ components over $b$.
-The disks at the top of the figure show the transit configurations of the occultor (light dark) and the star (light gray) for different values of $b$.
+The disks at the top of the figure show the transit configurations of the occultor (light dark) and the primary surface (light gray) for different values of $b$.
 \codelink{workflows/precision/scripts/plot_error}}
@@
 For reference, \autoref{fig:precision_limbdark} shows the relative error in the flux of a limb-darkened star occulted by an opaque body for increasing orders of a polynomial limb darkening law.
-In this figure, we also report the error of the flux computed using the \textsf{exoplanet} Python package (limited to linear and quadratic limb darkening).
+In this figure, we also report the error of the flux computed using the \textsf{exoplanet} Python package (using the same formalism but limited to linear and quadratic limb darkening).
 \begin{figure}[H]
@@
 This explains the larger errors observed in \autoref{fig:precision_limbdark} (error versus limb darkening order) compared to the ones shown in \autoref{fig:precision_degree} (error versus degree of the spherical harmonics).
-We not that the errors scale with the values of $u$.\\\\
-As described in \autoref{solution_vectors}, we compute \textit{starry} light curves using the Gauss Legendre quadrature to approximate the $\mathcal{P}$ integral involved in the expression of the solution vector $\bvec{s}$.
+We note that, here again, the errors scale with the values of $\bvec{u}$.\\\\
+As described in \autoref{solution_vector}, we compute occultation light curves using the Gauss Legendre quadrature to approximate the $\mathcal{P}$ integral involved in the expression of the solution vector $\bvec{s}^\mathsf{T}$.
 Hence, the precision of the computed flux $f$ depends on the order of the Legendre polynomial which defines the number of points used to approximate $\mathcal{P}$.
-For this reason, users must carefully set the order parameter to reach the precision required for their application, as shown in \autoref{fig:relative_error_1}.\\\\
+For this reason, users must carefully set the order parameter $q$ to reach the precision required for their application, as shown in \autoref{fig:relative_error_1}.\\\\
 \begin{figure}[H]
@@
         \caption{relative numerical error in the flux $f$ of a linearly limb-darkened star (degree $l$=1) transited by an opaque companion over the $(r, b)$ parameter space.
-The error is evaluated for different orders $n$ of the Gauss-Legendre quadrature and computed against \textsf{starry} for speed and convenience.
-Note that \autoref{fig:precision_s} was produced using $n=500$ (right plot) for which $\bvec{s}$ reaches machine precision.
+The error is evaluated for different orders $q$ of the Gauss-Legendre quadrature and computed against \textsf{starry} for speed and convenience.
+Note that \autoref{fig:precision_s} was produced using $q=500$ (right plot) for which $\bvec{s}$ reaches machine precision.
 \codelink{workflows/precision/scripts/plot_br_error}}
@@
 \end{figure}
-In \autoref{fig:precision_order_gausslegendre}, we show how the error in the solution vector evolves with the order $n$ of the Gauss-Legendre quadrature, and its impact on the error in the overall flux.
-\begin{figure}[H]
-    \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/precision/figures/error_order.pdf}
-        \caption{Maximum errors in the solution vector and overall flux $f$ for different occultor radii and orders $n$ of the Gauss-Legendre quadrature used to compute the P integral (see \autoref{solution_vectors}).
+% In \autoref{fig:precision_order_gausslegendre}, we show how the error in the solution vector evolves with the order $n$ of the Gauss-Legendre quadrature, and its impact on the error in the overall flux.
+% \begin{figure}[H]
+%     \begin{center}
+%         \includegraphics[width=\textwidth]{../workflows/precision/figures/error_order.pdf}
+%         \caption{Maximum errors in the solution vector and overall flux $f$ for different occultor radii and orders $n$ of the Gauss-Legendre quadrature used to compute the P integral (see \autoref{solution_vector}).
 Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s}, maxima taken over the range of values $b$.
 As in \autoref{fig:precision_s} the maximum degree of the spherical harmonics basis is set to $l=20$.}
-        \label{fig:precision_order_gausslegendre}
-    \end{center}
-\end{figure}
-The minimum error is always achieved for higher values of $n$, which comes with higher computational cost (see \autoref{speed}).
+%         \label{fig:precision_order_gausslegendre}
+%     \end{center}
+% \end{figure}
+The minimum error is always achieved for higher values of $q$, which comes with increased computation time (see \autoref{speed}).
 Hence users must carefully set the order parameter to balance precision and computational speed.
@@
         \includegraphics[width=\textwidth]{../workflows/precision/figures/error_order_degree.pdf}
-        \caption{Maximum errors in the overall flux $f$ for all degrees of the spherical harmonics components for different occultor radii and orders $n$ of the Gauss-Legendre quadrature used to compute the P integral (see \autoref{solution_vectors}).
+        \caption{Maximum errors in the overall flux $f$ for all degrees of the spherical harmonics components for different occultor radii and orders $q$ of the Gauss-Legendre quadrature used to compute the P integral (see \autoref{solution_vector}).
 Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s}, maxima taken over the range of values $b$.}
@@
 In this section, we evaluate the speed of the \textsf{jaxoplanet} implementation and compare it with the C++ implementation of \textsf{starry}.
-For quadratically limb-darkened light curve, we also compare our implementation with the one provided by the \textsf{exoplanet} Python package.\\\\
-In \autoref{fig:speed_order_gausslegendre}, we show the evolution of the computing time required to compute an occultation light-curve of a limb-darkened star and that of a non-uniform star described by spherical harmonics with a maximum degrees $l=20$, depending on the order $n$ of the Gauss-Legendre quadrature used to compute the $\mathcal{P}$ integral numerically (see \autoref{solution_vectors}).\\\\
+For quadratically limb darkened light curves, we also compare our implementation with the one provided by the \textsf{exoplanet} Python package.\\\\
+In \autoref{fig:speed_order_gausslegendre}, we show the evolution of the computing time required to compute an occultation light-curve of a limb-darkened star and that of a non-uniform star described by spherical harmonics with a maximum degrees $l=20$, depending on the order $q$ of the numerical integration.\\\\
 \begin{figure}[H]
@@
         \includegraphics[width=\textwidth]{../workflows/speed/figures/speed_vs_order.pdf}
-        \caption{Computation time of the occultation light curve of a limb-darkened star (left) and a non-uniform star described by spherical harmonics up to order $l=20$ (right) depending on the order $n$ of the Gauss-Legendre quadrature used to compute the P integral numerically (see \autoref{solution_vectors}).
-The processing time reported for \textsf{exoplanet} and \textsf{starry} is independent of $n$ as these two codes provide closed-form solutions for the integral $\mathcal{P}$.
+        \caption{Computation time of the occultation light curve of a limb-darkened star (left) and a non-uniform star described by spherical harmonics up to order $l=20$ (right) depending on the order $q$ of the Gauss-Legendre quadrature used to compute the P integral numerically (see \autoref{solution_vector}).
+The processing time reported for \textsf{exoplanet} and \textsf{starry} is independent of $q$ as these two codes provide closed-form solutions for the integral $\mathcal{P}$.
 Light curves are computed for $10\,000$ points in transit and an occultor radius $r=0.1$.}
@@
 
-\section{Case studies: ...}
-
-\subsection{}
+\section{Conclusion}
 

### 2024-10-08 workflows/precision/scripts/jax_f_limbdark.py

@@ -35,7 +35,7 @@ def u(deg):
 from jaxoplanet.core.limb_dark import light_curve
 
 function = jax.vmap(
-    lambda deg, b: light_curve(u(deg), b, r, order=100) + 1,
+    lambda deg, b: light_curve(u(deg), b, r, order=order) + 1,
     (None, 0),
 )
 

### 2024-10-08 workflows/precision/scripts/jax_fs.py

@@ -33,7 +33,6 @@ jax_s_function = jax.jit(
 
 jax_f_function = jax.vmap(jax_flux, in_axes=(None, 0))
 
-
 jax_f = []
 jax_s = []
 

### 2024-10-08 workflows/precision/scripts/plot_br_error.py

@@ -52,7 +52,7 @@ for i, order in enumerate(orders):
     else:
         ax.set_ylabel("b")
     ax.set_xlabel("r")
-    ax.set_title(rf"$n$ = {order}")
+    ax.set_title(rf"$q$ = {order}")
     colorbar(pcm)
 
 # cm_ax = fig.add_axes([1.5, 0.1, 0.02, 0.8])

### 2024-10-08 workflows/precision/scripts/plot_error.py

@@ -3,6 +3,9 @@
 # given degree l is the maximum error across all m \in [-l, l] components of that degree
 # (indices are for each degree are stored in the `indices` dictionary)
 
+import jax
+
+jax.config.update("jax_enable_x64", True)
 import matplotlib.pyplot as plt
 from matplotlib import gridspec
 import numpy as np

### 2024-10-08 workflows/precision/scripts/plot_error_limbdark.py

@@ -1,8 +1,13 @@
+import jax
+
+jax.config.update("jax_enable_x64", True)
+
 import matplotlib.pyplot as plt
 import numpy as np
 from jaxoplanet.experimental.starry.multiprecision import mp, utils
 
 degree = snakemake.params.degree
+q_order = int(snakemake.params.order)
 
 data = {
     "small": {
@@ -63,6 +68,13 @@ ax1.set_title("$r=0.01$")
 ax1.set_yscale("log")
 ax1.set_ylim(1e-17, 1e-4)
 ax1.set_xticks(x)
+ax1.annotate(
+    f"q={q_order}",
+    xy=(1 - 0.02, 0.05),
+    xycoords="axes fraction",
+    fontsize=10,
+    ha="right",
+)
 
 ax2 = plt.subplot(122)
 ax2.plot(degrees, err["large"]["starry"], ".-", label="starry", c=color["starry"])

### 2024-10-08 workflows/precision/scripts/plot_error_several_r_GL_degree_SH.py

@@ -33,7 +33,7 @@ for axi, ri in enumerate([0, 2, 4]):
         ax.set_yscale("log")
 
 ax.set_ylim(1e-16, 1)
-axes[1].set_xlabel("order $n$ of the Gauss-Legendre quadrature")
+axes[1].set_xlabel("order $q$ of the Gauss-Legendre quadrature")
 axes[0].set_ylabel("max relative error on the flux")
 
 l_arr = [0, 5, 10, 20]

### 2024-10-08 workflows/precision/scripts/plot_error_vs_degree.py

@@ -1,8 +1,13 @@
+import jax
+
+jax.config.update("jax_enable_x64", True)
 import matplotlib.pyplot as plt
 import numpy as np
 from jaxoplanet.experimental.starry.multiprecision import mp, utils
 
 l_max = snakemake.params.l_max
+
+q_order = int(eval(snakemake.input.jax[1].split("order=")[1].split(".")[0]))
 
 data = {
     "small": {
@@ -77,8 +82,15 @@ ax1.set_xlabel("degree of spherical harmonics")
 ax1.set_ylabel("relative error")
 ax1.set_title("$r=0.01$")
 ax1.set_yscale("log")
-ax1.set_ylim(1e-17, 1e-5)
+ax1.set_ylim(1e-17, 1e-4)
 ax1.set_xticks(x)
+ax1.annotate(
+    f"q={q_order}",
+    xy=(1 - 0.02, 0.05),
+    xycoords="axes fraction",
+    fontsize=10,
+    ha="right",
+)
 
 ax2 = plt.subplot(122)
 ax2.plot(
@@ -91,7 +103,7 @@ ax2.legend()
 ax2.set_xlabel("degree of spherical harmonics")
 ax2.set_title("$r=100$")
 ax2.set_yscale("log")
-ax2.set_ylim(1e-17, 1e-5)
+ax2.set_ylim(1e-17, 1e-4)
 ax2.set_yticklabels([])
 ax2.set_xticks(x)
 

### 2024-10-11 latex/paper.tex
big update

@@
 
-\title{Hardware-Accelerated Computation of Stellar Light Curves}
+\title{\textsf{jaxoplanet}: Hardware-Accelerated Computation of Stellar Light Curves}
 
@@
 \begin{abstract}
-    The measurement of stellar fluxes over time serves a wide variety of science cases.
-Transit and occultation light curves, for example, are used to characterize the orbital parameters of stellar systems, and infer the atmospheric properties of their transiting bodies.
+    Modeling light curves of extrasolar systems serves a wide variety of science cases.
+Transit and occultation light curves, for example, are used to characterize their orbital configuration, and infer the atmospheric properties of their transiting bodies.
 Phase curves, on the other hand, can be used to study the thermal emission of exoplanets, and the atmospheric circulation of stars.
 In order to infer these properties, a fast, accurate and robust model of stellar light curves is needed, one that accounts for the non-uniform surface intensity of spherical bodies and their mutual occultations.
-The analytical framework known as \textsf{starry} satisfies these requirements, and was successfully applied over the years following its development.
-However, as datasets and model complexity grow, the tractable inference of these parameters becomes increasingly challenging, and often motivates biased approximations.
+In this context, the analytical framework known as \textsf{starry} provided the community with a powerful tool, instrumental in ....
+However, as datasets and models complexity grow, the tractable inference of these parameters becomes increasingly challenging, and often motivates biased approximations.
 In this paper, we present a new implementation of the \textsf{starry} framework in \texttt{jax}, a high-performance machine-learning library designed for large-scale computations.
 We describe the changes made to the original version of \textsf{starry} and evaluate the performance of its new implementation that we release in the open-source Python package \textsf{jaxoplanet}.
-Through this implementation, we provide the community with a differentiable model of stellar light curves, and enable the use of powerful machine-learning tools available in the \texttt{JAX} ecosystem.
+Through this implementation, we provide the community with a differentiable model of stellar light curves, and enable the use of powerful machine-learning tools available in the \textsf{JAX} ecosystem.
 \gitlink{}
@@
 
-During millenia, planets of the solar system were mostly known and studied through their orbits.
-Starting in the 16th century, Earth-based telescopes progressively enabled the direct imaging of their surface.
-But despite these new instruments, some minor planets were too small and too distant to be imaged, and remained unresolved until the 20th century, when more capable space-based observatories were launched.
-This is the case of Pluto, for example, whose surface was first studied through the light it reflects during its revolution and orbit around the Sun.
-Using point source light curves obtained from 1945 to 1986, \cite{Drish1995} tentatively reconstructed the map of the planet, only later compared to the first \textit{Hubble Space Telescope} low-resolution images obtained by \cite{Stern1997}.
-This mapping effort was gloriously completed with the extremely detailed pictures obtained during the close fly-by of the \textit{New Horizons} spacrecraft \citep{}.\\\\
-The study of extrasolar planets is taking a path similar to that of Pluto, starting with the study of their unresolved light curves \citep{} while direct imaging instruments are being developed \citep{}.
-The same applies to the study of unresolved stellar atmospheres, whose properties are inferred from the light they emit.
-In this context, developments of accurate and robust models of extrasolar systems light curves are crucial.
-As these models are being compared to large datasets in probabilistic ways, their posterior likelihood must be evaluated millions of times, which require them to be fast.
-This inference can be highly facilitated by the advent of gradient-based optimization and inference tools, so that the differentiability of these models offer a clear advantage.\\\\
-In this context, \cite{starry} presented an algorithm to analytically compute the light curves of spherical bodies with non-uniform surface intensities, particularly optimized for occultation light curves.
-This algorithm, known as \textit{starry}, has been successfully applied to a wide variety of science cases, from the study of exoplanets atmospheres \citep{} to the study of stellar photospheres \citep{}.
-However, as datasets and model complexity grow, the tractable inference of these parameters becomes increasingly challenging, and often motivates biased approximations.
-This is case, for example, in transmission spectroscopy measurements conducted with the \textit{James Webb Space Telescope}, for which ...
-\todo{(So, give an example of large number of spectral bins and required approximation) (Hatori et al.
-in prep)}.\\\\
-In this paper, we present a new implementation of the \textit{starry} framework in \texttt{JAX}, a high-performance machine-learning library designed for large-scale computations.
-We start in \autoref{starry} with a summarry of the formalism introduced by \cite{starry}.
-In \autoref{optimization}, we describe the changes made to the original version of \textit{starry}.
+Planets of the solar system used to be known and studied through their position in the sky.
+It is only starting in the 16th century that Earth-based telescopes progressively enabled the direct imaging of their surface.
+But despite these new instruments, some minor planets were too small and too distant to be imaged, and remained unresolved until the end of the 20th century, when more capable space-based observatories were launched.
+This is the case of Pluto, for example, whose surface was first studied thanks to the light it reflects while rotating and orbiting the Sun.
+Using point source observations obtained from 1945 to 1986, \cite{Drish1995} tentatively reconstructed the map of the planet, and only later compared it to the first \textit{Hubble Space Telescope} low-resolution images obtained by \cite{Stern1997}.
+This mapping effort was gloriously completed with the extremely detailed pictures obtained by the \textit{New Horizons} spacrecraft during its fly-by of the planet \citep{Stern2015}.\\\\
+Although imaging the surface of an exoplanet might be several decades away, their study follows that of Pluto, by first employing indirect methods and unresolved light curves (see \citealt{Mayor1995} and \citealt{Charbonneau2000})  while waiting for the development of more powerful direct imaging instruments \cite{}.
+Similarly, stellar photospheres can currently be mapped using interferometry and Doppler imaging, but these techniques are costly and can only be applied to a handful of objects, so that time-resolved spectrophotometry remains our main way to study them.
+In this context, the development of accurate and robust models of stellar light curves is crucial.\\\\
+In the age of wide-field surveys (such as) and high-precision spectroscopic instruments, these models are being compared to large datasets in probabilistic ways, requiring fast and robust evaluations of their posterior likelihood over larger and larger parameter spaces.
+These inferences can be highly facilitated by gradient-based optimization and sampling algorithms, so that being able to differentiate these models with respect to their parameters offers clear advantages.\\\\
+In this context, \cite{starry} and \cite{Agol2020} presented a framework to analytically compute the light curves of spherical bodies with non-uniform surface intensities, such as stars and exoplanets, and the flux observed from their mutual occultations.
+This framework, known as \textit{starry}, has been successfully applied to a wide variety of science cases, like the study of transiting exoplanets \citep{Kostov2019}, precise modeling of radial velocities \citep{Dawson2021}, the study of exoplanets atmospheres \citep{Bell2023} or the study of stellar photospheres \citep{Almenara2022}.
+However, as datasets and models complexity grow, the tractable inference of the larger parameter spaces we have to deal with becomes increasingly challenging, and often motivates biased approximations.
+This is the case, for example, in transmission spectroscopy measurements conducted with the \textit{James Webb Space Telescope} NIRSpec instrument, that should ideally consist in the simultaneous modeling of more than 3000 transit light curves, one per pixel element (e.g.
+\citealt{Alderson2023}).\\\\
+In this paper we present \textsf{jaxoplanet}, a new implementation of the \textit{starry} framework in \texttt{JAX}, a high-performance machine-learning library designed for large-scale computations, including autodifferentiation capabilities, deployed on various hardware.
+In \autoref{starry}, we summarize the formalism introduced by \cite{starry} to analytically compute stellar light curves.
+In \autoref{optimization}, we describe the \textsf{jaxoplanet} implementation and the changes made to the \textsf{starry} implementation.
 Finally, in \autoref{performance}, we evaluate the performance of our new implementation that we release in the open-source Python package \textsf{jaxoplanet}.
@@
 
-\newpage
 \section{Light curve models}\label{starry}
@@
 Inspired by the idea of \cite{pal2012}, \cite{starry} and \cite{Agol2020} present a general framework to compute the light curves of spherical bodies with non-uniform surface intensities.
-This framework, known as \textit{starry}, allows for precise and efficient computation of the light curves of stars and exoplanets, and has been successfully applied to a wide variety of science cases.
 In this section, we describe the mathematical formalism behind the \textit{starry} model.
@@
 The surface of a spherical body can be naturally described by spherical harmonics: a set of special functions defined on the sphere that can serve as orthonormal basis to represent any spherical surface map.
-If $\bvec{\tilde y}$ denotes the spherical harmonics basis of degree $l_{max}$ (defined in \citealt[section X]{starry}) and $\bvec{y}$ a vector that contains the coefficients describing the surface in this basis, the intensity $I$ of the map projected onto the sky-projected $(x, y)$ plane of the observer is
+If $\bvec{\tilde y}$ denotes the spherical harmonics basis of degree $l_{max}$ (defined in \citealt[section 2.2]{starry}) and $\bvec{y}$ a vector that contains the coefficients describing the surface in this basis, the intensity $I$ of the map projected onto the sky-projected $(x, y)$ plane of the observer is
 \begin{equation}I(x, y) = \bvec{\tilde y^\mathsf{T}}(x, y) \, \bvec{y}.\end{equation}
@@
 \begin{equation}F = \iint I(x,y)\,\text{d}S = \iint \bvec{\tilde y^\mathsf{T}}\bvec{R}\bvec{y}\,\text{d}S\end{equation}
-In order to simplify this integral, \citealt{starry} introduce a change of basis from the spherical harmonics basis $\bvec{\tilde y}$ to a simple polynomial basis $\bvec{\tilde p}$ characterized by the change of basis matrix $\bvec{A_1}$.
+In order to simplify this integral, \citealt{starry} introduce a change of basis from the spherical harmonics basis $\bvec{\tilde y}$ to a simple polynomial basis $\bvec{\tilde p}$ (defined in \citealt[section 2.3]{starry}) characterized by the change of basis matrix $\bvec{A_1}$.
 In this new basis, $\bvec{y}$ can be expressed as
@@
 \subsection{Occultation light curve}\label{occultation}
-The same formalism can be used to compute the integrated intensity of a surface map occulted by another spherical body.
+In order to model transits and occultations, the same formalism can be used to compute the integrated intensity of a surface map occulted by another spherical body.
 In this case, the intensity of the map must be integrated only over the unocculted surface $S$ of the disk, such that
 \begin{equation} F = \iint_S I(x,y)\,\text{d}S = \left(\iint_S \bvec{\tilde p^\mathsf{T}}\,\text{d}S\right)\bvec{A_1}\bvec{R}\bvec{y}.\end{equation}
-To simplify this expression, \citealt{pal2012} note that any two-dimensional integral of the unocculted disk can be reduced to a one-dimensional integral over the contour $C$ of the occulting body using Green's theorem\footnote{Also known as Stokes' theorem in three-dimensional vector calculus.}.
+To simplify this expression, \cite{pal2012} note that any two-dimensional integral of the unocculted disk can be reduced to a one-dimensional integral over the contour $C$ of the occulting body using Green's theorem\footnote{Also known as Stokes' theorem in three-dimensional vector calculus.}.
 This can be done if the integrand of the integral over S can be described by a function $f$ such that
@@
 where $G$ is a vector field defined and differentiable over $S$.\\\\
-By applying a last change of basis matrix $\bvec{A_2}$ to the polynomial basis $\bvec{\tilde p}$, \citealt{starry} transforms $\bvec{\tilde p}$ to the \textit{Green's basis} $\bm{\tilde g}$ that satisfies this property, meaning that the surface integral of each of its components $\tilde{g_n}$ can be transformed to the line integral
+By applying a last change of basis matrix $\bvec{A_2}$ to the polynomial basis $\bvec{\tilde p}$, \cite{starry} transforms $\bvec{\tilde p}$ to the \textit{Green's basis} $\bm{\tilde g}$ that satisfies this property, meaning that the surface integral of each of its components $\tilde{g_n}$ can be transformed to the line integral
 \begin{equation}
@@
     \item The surface is simply described by $\bvec{u}$, a vector of polynomial limb darkening coefficients.
-    \item The Green's basis takes a simpler form (see \citealt[Eq. X]{Agol2020}).
+    \item The Green's basis takes a simpler form (see \citealt[Equation 14]{Agol2020}).
     \item Matrices involved in the computation of the integrated flux have $l_{max}$ columns, versus $(l_{max}+1)^2$ in the general case, reducing drastically the computational cost of the model.
@@
 \section{Implementation details}\label{optimization}
+The formalism introduced by \cite{starry}, summarized in \autoref{starry}, provides an elegant analytical model of stellar light curves.
+Implemented in C++ and released in the \texttt{starry} Python package, it was built for the probabilistic programming library \textsf{PyMC3} and led to several advances in the field of exoplanetary science.
+However, since its initial development, the \textsf{PyMC} project adopted several breaking changes that made \textsf{starry} incompatible with its latest versions.
+Updating \textsf{starry} to the latest version of \textsf{PyMC} would require a complete rewriting of the code, that would again be highly dependent on the latter.
+Hence, our first motivation comes from the broken dependencies of \textsf{starry} and aims to provide the community with an implementation compatible with a wider ecosystem of tools.
+In this respect \textsf{JAX} represents an interesting alternative.
+This high-performance machine-learning library can be used to perform hardware-accelerated computations on various type of devices (CPU, GPU, TPU) and gave rise to probabilistic tools developped by an ever-growing community of scientists and engineers.
+However, \textsf{JAX} capabilities really shine when models are written with a functional-programming mindset, and computations performed in a vectorized way.\\\\
+In this section, we describe the changes made to the original \textsf{starry} implementation to allow its rewriting in \textsf{JAX}.
+We also highlight details about the latest version of \textsf{starry}'s original implementation that were omitted in \citealt{starry} but described in parts of its code and documentation.
 
@@
 The multiplication of the base map represented by $\bvec{y}$ with the limb darkening map represented by $\bvec{y}_u$ can be seen as the multiplication of two polynomials, only with special coefficients (called Clebsch–Gordan coefficients) inherent to the properties of spherical harmonics.
-To simplify this operation, we transform both maps into the polynomial basis, so that their product is the actual product of two polynomials which doesn't need the computation of special coefficients.
+To simplify this operation, we transform both maps into the polynomial basis, so that their product is the actual product of two polynomials which doesn't require the computation of special coefficients.
 
@@
     \begin{center}
-        \includegraphics[width=0.6\textwidth]{../workflows/rotations/figures/rotation_basis.pdf}
-        \caption{Rotation of a primary body's surface to place it in the sky-projected frame of the observer.}
+        \includegraphics[width=0.6\textwidth]{../workflows/geometry/figures/rotation_basis.pdf}
+        \caption{Rotation of a primary body's surface to place it in the sky-projected frame of the observer. \codelink{workflows/geometry/scripts/rotations.ipynb}}
         \label{fig:rotation_basis}
@@
     \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/rotations/figures/rotation_starry.pdf}
+        \includegraphics[width=\textwidth]{../workflows/geometry/figures/rotation_starry.pdf}
         \caption{Decomposition of the full rotation of the surface into elementary rotations.
-Each schematic shows a sphere rotated from a previous orientation around an axis displayed as a black vector and an angle shown on top.}
+Each schematic shows a sphere rotated from a previous orientation around an axis displayed as a black vector and an angle shown on top.
+\codelink{workflows/geometry/scripts/rotations.ipynb}}
         \label{fig:rotation_starry}
@@
     \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/rotations/figures/rotation_jaxoplanet_1.pdf}
+        \includegraphics[width=\textwidth]{../workflows/geometry/figures/rotation_jaxoplanet_1.pdf}
         \caption{Consecutive rotations of the spherical harmonics basis in the \textsf{jaxoplanet} implementation.
-Each schematic shows a sphere rotated from a previous orientation around an axis displayed as a black vector and an angle shown on top.}
+Each schematic shows a sphere rotated from a previous orientation around an axis displayed as a black vector and an angle shown on top.
+\codelink{workflows/geometry/scripts/rotations.ipynb}}
         \label{fig:rotation_jaxoplanet}
@@
 On the left figure, the larger light grey disk corresponds to the primary body being occulted by the darker secondary body.
-On the right figure, the area $S$ on which the intensity is integrated is shown in light blue.
-By choosing an appropriate transform, the integration over $S$ can be reduced to a line integral over the contour of the occulting body, which can be decomposed into the contour of the occulting body within the area of the primary body (in cyan) and the contour of the primary body everywhere else (in blue).}
+On the right figure, the area $S$ over which the intensity is integrated is shown in light blue.
+By choosing an appropriate transform, the integration over $S$ can be reduced to a line integral over the contour of the occulting body, which can be decomposed into the contour of the occulting body within the area of the primary body (in cyan) and the contour of the primary body everywhere else (in blue).
+\codelink{workflows/geometry/scripts/occultation_geometry.ipynb}}
         \label{fig:geometry}
@@
 %
-In the general case, the $\mathcal{Q}$ integral takes a simple analytical form and can be computed using the recurrence relations from \cite{starry} (Equation D29).
+In the general case, the $\mathcal{Q}$ integral (over the blue contour shown in \autoref{fig:geometry}) takes a simple analytical form and can be computed using the recurrence relations from \cite{starry} (Equation D29).
 The case of a limb-darkened surface is even simpler as $\mathcal{Q}(\bvec{G}_n) = 0$ for all $n > 0$.
-On the other hand, the $\mathcal{P}$ integral is harder to evaluate and requires the use of recurrence relations and truncated series expansions, with a scheme that depends on the geometric configuration of the system (see \citealt{starry}, Section D.2).\\\\
-In order to translate the \textsf{starry}  implementation to \textsf{JAX}, we evaluate the $\mathcal{P}$ integrals numerically using Gauss-Legendre integrations, requiring the evaluation of the integrated functions only over a fixed number of points.\\\\
+On the other hand, the $\mathcal{P}$ integral (over the cyan contour shown in \autoref{fig:geometry}) is harder to evaluate and requires the use of recurrence relations and truncated series expansions, with a scheme that depends on the geometric configuration of the system (see \citealt{starry}, Section D.2).\\\\
+In order to translate the \textsf{starry}  implementation to \textsf{JAX}, we evaluate the $\mathcal{P}$ integrals numerically using Gauss-Legendre integrations, requiring the evaluation of the integrated functions only over a fixed number of points $q$.\\\\
 
@@
         \pi - \kappa_1 - r^2\kappa_0 + \textit{A}_{kite} & \qquad n = 0
+        \\[1em]
+        \text{see below} & \qquad n = 1
         \\[1em]
@@
 \begin{equation}
-    k = \frac{1-r^2 - b^2 + 2br}{4br}.
+    k = \frac{1-r^2 - b^2 + 2br}{4br},
 \end{equation}\\
 %
-For $n=1$, we use Equation 31, 32 and 34 from \cite{starry} to write
+also used in the case $n=3$. For $n=1$, we use Equation 31, 32 and 34 from \cite{starry} to write
 %
@@
 
-\newpage
+\subsection{Kepler's equation solver}
+\todo{}
+
 \section{Performance}\label{performance}
 Although occultation light curves have analytical expressions, their practical implementation relies on numerical approximations chosen to balance precision and computation time.
-This is the case, for example, of the solution vector $\bvec{s}^\mathsf{T}$, computed using numerical series expansions in \textsf{starry} (\citealt[section D.2.3]{starry}) and numerical integration in \textsf{jaxoplanet} (see \autoref{solution_vector}).
+This is the case for the solution vector $\bvec{s}^\mathsf{T}$, computed using numerical series expansions in \textsf{starry} (\citealt[section D.2.3]{starry}) and numerical integration in \textsf{jaxoplanet} (see \autoref{solution_vector}).
 Other expressions, such as the Wigner-D matrices used in the rotation of the spherical harmonics basis, are computed using reccurence relations prone to the accumulation of numerical errors and instability.\\\\
@@
 Although we show in \autoref{fig:precision_SAR} that the precision of the \textsf{starry} and \textsf{jaxoplanet} implentations slightly differ for many of these terms, we focus this section on the solution vector $\bvec{s}^\mathsf{T}$ and the overall flux $f$.\\\\
-Assuming that we perform the numerical integration described in \autoref{solution_vector} at a relatively high order ($q=500$; see \autoref{solution_vector}), \autoref{fig:precision_s} shows that our reimplentation of \textsf{starry} reaches relative errors lower than 1 part per billion for $\bvec{s}^\mathsf{T}$ and the overall flux $f$, considering both a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$.
+Assuming that we perform the numerical integration described in \autoref{solution_vector} at a relatively high order $q=500$, \autoref{fig:precision_s} shows that our reimplentation of \textsf{starry} reaches relative errors lower than 1 part per billion for $\bvec{s}^\mathsf{T}$ and the overall flux $f$, considering both a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$.
 For reference, \autoref{fig:precision_s_starry} shows comparable errors on the same quantities computed with the C++ implentation of \textsf{starry}.
-In addition, \autoref{fig:precision_degree} shows the evolution of the error on $f$ for increasing degrees of the spherical harmonics, both for \textsf{starry} and \textsf{jaxoplanet}.
-We note that the coefficients associated to spherical harmonics of degrees $l$ usually take values of the order of magnitude $10^{-l}$.
-Given that the numerical errors in computing \autoref{eq:starry} scale lineraly with the values of the coefficients $\bvec{y}$, the results shown in \autoref{fig:precision_degree}, for spherical harmonics coefficients equal to 1, translate to machine precision in most practical applications.
-With these results, we validate the precision of \textsf{jaxoplanet} and show it is on per with the legacy C++ \textsf{starry} implentation\footnote{version 1.2.0 \citep{starry_120}} described in \cite{starry}.\\\\
 \begin{figure}[H]
@@
         \includegraphics[width=\textwidth]{../workflows/precision/figures/error_jax.pdf}
-        \caption{Error of the solution vector $\bvec{s}^\mathsf{T}$ and the overall flux $f$ for all terms of the spherical harmonics basis up to degree $l=20$.
+        \caption{Error of the solution vector $\bvec{s}^\mathsf{T}$ and the overall flux $f$ for all terms of the spherical harmonics basis up to degree $l_{max}=20$.
 Errors are computed against arbitrary-precision computations of the solution vector and the flux for a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$.
@@
 \end{figure}
+In addition, \autoref{fig:precision_degree} shows the evolution of the error on $f$ for increasing degrees of the spherical harmonics for \textsf{starry} and \textsf{jaxoplanet}.
+With these results, we validate the precision of \textsf{jaxoplanet} and show it is on per with the legacy C++ \textsf{starry} implentation\footnote{version 1.2.0 \citep{starry_120}} described in \cite{starry}.
 \begin{figure}[H]
@@
         \caption{Maximum errors in the overall flux $f$ for all degrees of the spherical harmonics components.
-Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s}, maxima taken over the range of values $b$.
+Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s} over the range of values $b$, but computed with $q=250$.
 \codelink{workflows/precision/scripts/plot_error_vs_degree}}
@@
 For reference, \autoref{fig:precision_limbdark} shows the relative error in the flux of a limb-darkened star occulted by an opaque body for increasing orders of a polynomial limb darkening law.
-In this figure, we also report the error of the flux computed using the \textsf{exoplanet} Python package (using the same formalism but limited to linear and quadratic limb darkening).
+In this figure, we also report the error of the flux computed using the \textsf{exoplanet} Python package (which uses the same formalism as \cite{Agol2020} but limited to linear and quadratic limb darkening).
 \begin{figure}[H]
@@
         \caption{Maximum errors in the overall flux $f$ of a star occulted by an opaque body for increasing orders of its polynomial limb darkening law.
-Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s}, taken over the range of values $b$.
+Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s} over the range of values $b$, but computed with $q=100$.
 For each order $n$ all limb darkening coefficients are set to zero except the $n$-th coefficient $u_n=1$.
@@
 This explains the larger errors observed in \autoref{fig:precision_limbdark} (error versus limb darkening order) compared to the ones shown in \autoref{fig:precision_degree} (error versus degree of the spherical harmonics).
-We note that, here again, the errors scale with the values of $\bvec{u}$.\\\\
+We note that the errors scale with the values of $\bvec{u}$.\\\\
 As described in \autoref{solution_vector}, we compute occultation light curves using the Gauss Legendre quadrature to approximate the $\mathcal{P}$ integral involved in the expression of the solution vector $\bvec{s}^\mathsf{T}$.
@@
         \caption{relative numerical error in the flux $f$ of a linearly limb-darkened star (degree $l$=1) transited by an opaque companion over the $(r, b)$ parameter space.
-The error is evaluated for different orders $q$ of the Gauss-Legendre quadrature and computed against \textsf{starry} for speed and convenience.
+The error is evaluated for different orders $q$ of the Gauss-Legendre quadrature and computed against values of $f$ obtained with $q=800$.
 Note that \autoref{fig:precision_s} was produced using $q=500$ (right plot) for which $\bvec{s}$ reaches machine precision.
@@
 
-One of the advantage of the \texttt{JAX} library is its ability to perform hardware-accelerated computations on CPUs and GPUs.
+One of the advantage of the \textsf{JAX} library is its ability to perform hardware-accelerated computations on CPUs and GPUs.
 In this section, we evaluate the speed of the \textsf{jaxoplanet} implementation and compare it with the C++ implementation of \textsf{starry}.
 For quadratically limb darkened light curves, we also compare our implementation with the one provided by the \textsf{exoplanet} Python package.\\\\
-In \autoref{fig:speed_order_gausslegendre}, we show the evolution of the computing time required to compute an occultation light-curve of a limb-darkened star and that of a non-uniform star described by spherical harmonics with a maximum degrees $l=20$, depending on the order $q$ of the numerical integration.\\\\
+In \autoref{fig:speed_order_gausslegendre}, we show the evolution of the time required to compute the occultation light curve of a limb darkened star and that of a non-uniform star described by spherical harmonics with a maximum degrees $l_{max}=20$, depending on the order $q$ of the numerical integration.
+On CPU, our benchmark is done on the single core of Apple M2 max chip, while on GPU we use the single core of an NVIDIA Tesla V100 processor.\\\\
 \begin{figure}[H]
@@
         \includegraphics[width=\textwidth]{../workflows/speed/figures/speed_vs_order.pdf}
-        \caption{Computation time of the occultation light curve of a limb-darkened star (left) and a non-uniform star described by spherical harmonics up to order $l=20$ (right) depending on the order $q$ of the Gauss-Legendre quadrature used to compute the P integral numerically (see \autoref{solution_vector}).
+        \caption{Computation time of the occultation light curve of a limb-darkened star (left) and a non-uniform star described by spherical harmonics up to order $l_{max}=20$ (right) depending on the order $q$ of the Gauss-Legendre quadrature used to compute the $\mathcal{P}$ integral numerically (see \autoref{solution_vector}).
 The processing time reported for \textsf{exoplanet} and \textsf{starry} is independent of $q$ as these two codes provide closed-form solutions for the integral $\mathcal{P}$.
@@
         \includegraphics[width=\textwidth]{../workflows/speed/figures/speed_vs_N.pdf}
-        \caption{Computation time of the occultation light curve of a limb-darkened star (left) and a non-uniform star described by spherical harmonics up to order $l=20$ (right) depending on the number of points in the light curve.
-Light curves are computed at order $n=200$ and an occultor radius $r=0.1$.}
+        \caption{Computation time of the occultation light curve of a limb-darkened star (left) and a non-uniform star described by spherical harmonics up to order $l_{max}=20$ (right) depending on the number of points in the light curve.
+Light curves are computed at $q=200$ and an occultor radius $r=0.1$.}
         \label{fig:speed_N}
@@
 \autoref{fig:speed_order_gausslegendre}, in combination with \autoref{fig:precision_order_gausslegendre_degree}, can be used to determine a value of $n$ balancing precision and computation time.\\\\
-Finally, \autoref{fig:speed_N} shows the evaluation time of \textsf{jaxoplanet} as a function of the number of points in the light curve, compared to \textsf{starry} for an $l=20$ surface map, and \textsf{exoplanet} for a quadratically limb-darkened star.
-As in the previous figure, we show that \textsf{jaxoplanet} is competitive against state-of-the-art codes on CPUs and offer a clear advantage on GPUs.
+Finally, \autoref{fig:speed_N} shows the evaluation time of \textsf{jaxoplanet} as a function of the number of points in the light curve, compared to \textsf{starry} for a surface map of degree $l_{max} = 20$, and \textsf{exoplanet} for a quadratically limb darkened star.
+As in the previous figure, we show that \textsf{jaxoplanet} is competitive against state-of-the-art codes on CPUs and offers a clear advantage on GPUs for large datasets.
+
+\section{Discussion}
+\todo{}
 
 \section{Conclusion}
+In this paper we presented \textsf{jaxoplanet}, a new implementation of the analytical light curve model \textit{starry} developed using the machine learning library \textsf{JAX}.
+We showed that our implementation is on par with the legacy C++ \textsf{starry} implementation, and that it offers a significant speed advantage on GPUs for large datasets, enabled by \textsf{JAX}'s just-in-time compilation and hardware acceleration.
+Our code can be used in association with the large ecosystem of probabilistic and sampling libraries developed in \textsf{JAX}, and will facilitate the development of tools and pipeline for the modeling of TESS and JWST observations, but also future missions such as PLATO \citep{Rauer2014} and ARIEL \citep{Tinetti2018}.\\\\
+As the core of our model is developed in Python, we expect \textsf{jaxoplanet} to serve as a basis for the implementation of more complex light curve models, such as light curves of Doppler maps \citep{Luger2021c}, reflected maps \citep{Luger2022}, or non-spherical bodies maps \citep{Dholakia2024}.
+Eventually, \textsf{jaxoplanet} will take over \textsf{starry} in acting as a foundation to explore and develop analytical methods to numerically reconstruct the surface of exoplanets and stars, venturing a bit further into the understanding of exoplanetary systems and their worlds.\\\\
+The \textsf{jaxoplanet} code is open source under the MIT License and is hosted at \href{https://github.com/exoplanet-dev/jaxoplanet/}{https://github.com/exoplanet-dev/jaxoplanet}, with documentation and tutorials hosted at \href{https://jax.exoplanet.codes}{https://jax.exoplanet.codes}.
+A permanent version of the code used to generate the figures and results in this paper is archived at https://doi.org/10.5281/zenodo.1312286.
 
@@
         \includegraphics[width=\textwidth]{../workflows/precision/figures/error_starry.pdf}
-        \caption{Same as \autoref{fig:precision_s} but with $bvec{s}$ and $f$ computed with the C++ implementation of \textsf{starry}. \codelink{workflows/precision/scripts/plot_error}}
+        \caption{Same as \autoref{fig:precision_s} but with $\bvec{s}$ and $f$ computed with the C++ implementation of \textsf{starry}. \codelink{workflows/precision/scripts/plot_error}}
         \label{fig:precision_s_starry}

### 2024-10-11 workflows/precision/scripts/concatenate_errors_GL.py

@@ -1,6 +1,10 @@
+import jax
+
+jax.config.update("jax_enable_x64", True)
+
 import numpy as np
 import matplotlib.pyplot as plt
-from jaxoplanet.experimental.starry.mpcore import utils, mp
+from jaxoplanet.experimental.starry.multiprecision import utils, mp
 from tqdm import tqdm
 
 rs = snakemake.params.rs

### 2024-10-11 workflows/precision/scripts/jax_f_nonuniform_limbdark.py

@@ -26,7 +26,7 @@ def u(deg):
 
 function = jax.vmap(
     lambda deg, b: surface_light_curve(
-        Surface(u=u(deg)), y=b, z=10.0, r=r, order=order
+        Surface(u=u(deg)), y=b, z=10.0, r=r, order=order, higher_precision=True
     ),
     (None, 0),
 )

### 2024-10-11 workflows/precision/scripts/jax_fs.py

@@ -21,7 +21,9 @@ order = int(snakemake.wildcards.order)
 @jax.jit
 def jax_flux(y, b):
     surface = Surface(y=Ylm.from_dense(y, normalize=False), normalize=False)
-    return surface_light_curve(surface, r=r, y=b, z=10.0, order=order)
+    return surface_light_curve(
+        surface, r=r, y=b, z=10.0, order=order, higher_precision=True
+    )
 
 
 jax_s_function = jax.jit(

### 2024-10-11 workflows/precision/scripts/plot_br_error.py

@@ -7,7 +7,7 @@ br = np.load(snakemake.input["br"])
 r = br["r"]
 b = br["b"]
 data = {
-    "starry": np.load(snakemake.input["starry"]),
+    "ref": np.load(snakemake.input["reference"]),
     "jaxoplanet": {
         orders[i]: np.load(snakemake.input["jaxoplanet"][i]) for i in range(len(orders))
     },
@@ -36,7 +36,7 @@ fig, axes = plt.subplots(1, len(orders), figsize=(8.5, 2.5))
 for i, order in enumerate(orders):
     ax = axes[i]
     ax.set_aspect("equal", "box")
-    errors = np.abs(data["starry"]["flux"] - data["jaxoplanet"][order]["flux"])
+    errors = np.abs(data["ref"]["flux"] - data["jaxoplanet"][order]["flux"])
     errors[errors == 0] = 1e-20
     errors[np.logical_not(np.isfinite(errors))] = 1e-20
 

### 2024-10-11 workflows/precision/scripts/plot_error_limbdark.py

@@ -12,14 +12,14 @@ q_order = int(snakemake.params.order)
 data = {
     "small": {
         "jax": np.load(snakemake.input.jax[0], allow_pickle=True)["f"],
-        "jax_nu": np.load(snakemake.input.jax_nu[0], allow_pickle=True)["f"],
+        "jax_nu": np.load(snakemake.input.jax_ld[0], allow_pickle=True)["f"],
         "num": np.load(snakemake.input.num[0], allow_pickle=True)["f"],
         "starry": np.load(snakemake.input.starry[0], allow_pickle=True)["f"],
         "exo": np.load(snakemake.input.exo[0], allow_pickle=True)["f"],
     },
     "large": {
         "jax": np.load(snakemake.input.jax[1], allow_pickle=True)["f"],
-        "jax_nu": np.load(snakemake.input.jax_nu[0], allow_pickle=True)["f"],
+        "jax_nu": np.load(snakemake.input.jax_ld[1], allow_pickle=True)["f"],
         "num": np.load(snakemake.input.num[1], allow_pickle=True)["f"],
         "starry": np.load(snakemake.input.starry[1], allow_pickle=True)["f"],
         "exo": np.load(snakemake.input.exo[1], allow_pickle=True)["f"],
@@ -40,7 +40,7 @@ err = {
     },
     "large": {
         "jax": errors(data["large"]["jax"], data["large"]["num"]).max(1)[1:],
-        "jax_nu": errors(data["small"]["jax_nu"], data["small"]["num"]).max(1)[1:],
+        "jax_nu": errors(data["large"]["jax_nu"], data["large"]["num"]).max(1)[1:],
         "starry": errors(data["large"]["starry"], data["large"]["num"]).max(1)[1:],
         "exo": errors(data["large"]["exo"], data["large"]["num"][0:3, :]).max(1)[1:],
     },

### 2024-10-15 latex/paper.tex
feat: realistic speed

@@
 \todo{}
+\subsection{Light curve functions and transformations}
+\todo{}
 
@@
 
-\subsection{Precision}
+\subsection{Precision}\label{precision}
 We evaluate the precision of the \textsf{starry} and \textsf{jaxoplanet} terms against quantities computed at arbitrary precision, using a ground truth version of the code implemented with the \texttt{mpmath} Python library\footlink{https://mpmath.org/}.
@@
 For quadratically limb darkened light curves, we also compare our implementation with the one provided by the \textsf{exoplanet} Python package.\\\\
-In \autoref{fig:speed_order_gausslegendre}, we show the evolution of the time required to compute the occultation light curve of a limb darkened star and that of a non-uniform star described by spherical harmonics with a maximum degrees $l_{max}=20$, depending on the order $q$ of the numerical integration.
-On CPU, our benchmark is done on the single core of Apple M2 max chip, while on GPU we use the single core of an NVIDIA Tesla V100 processor.\\\\
+In \autoref{fig:speed_order_gausslegendre}, we show the evolution of the time required to compute the occultation light curve of a limb darkened star and that of a non-uniform star described by spherical harmonics of degree $l_{max}=20$, depending on the order $q$ of the numerical integration.
+On CPU, our benchmark is done on the single core of an Apple M2 max chip, while on GPU we use the single core of an NVIDIA Tesla V100 processor.\\\\
 \begin{figure}[H]
@@
 \end{figure}
+Finally, \autoref{fig:speed_N} shows the evaluation time of \textsf{jaxoplanet} as a function of the number of points in the light curve compared to \textsf{exoplanet}, for a quadratically limb darkened star, and \textsf{starry}, for a surface map of degree $l_{max} = 20$.
 \begin{figure}[H]
@@
 \end{figure}
-\autoref{fig:speed_order_gausslegendre}, in combination with \autoref{fig:precision_order_gausslegendre_degree}, can be used to determine a value of $n$ balancing precision and computation time.\\\\
-Finally, \autoref{fig:speed_N} shows the evaluation time of \textsf{jaxoplanet} as a function of the number of points in the light curve, compared to \textsf{starry} for a surface map of degree $l_{max} = 20$, and \textsf{exoplanet} for a quadratically limb darkened star.
-As in the previous figure, we show that \textsf{jaxoplanet} is competitive against state-of-the-art codes on CPUs and offers a clear advantage on GPUs for large datasets.
+These benchmarks are produced for values of $q$ which guarantee that our models reach the same precision as state-of-the-art codes.
+As discussed in \autoref{discussion}, \textsf{jaxoplanet} can be further optimized by reducing the order $q$ of numerical integration while keeping the precision required for specific applications.
 
-\section{Discussion}
-\todo{}
+\section{Discussion}\label{discussion}
+
+In \autoref{precision}, we showed that \textsf{jaxoplanet} is on par with the C++ implementation of \textsf{starry} in terms of precision.
+As relatively high values of $q$ are required to achieve this performance, results from \autoref{speed} are deceiving and suggest that \textsf{jaxoplanet} is about an order of magnitude slower than \textsf{exoplanet} or \textsf{starry} on CPU, and only advantageous on GPU for very large datasets.
+Although we expect to further optimize our \textsf{JAX} implementation and bridge this gap in the future, \textsf{jaxoplanet} can currently outperform state-of-the-art codes for most realistic datasets.
+Indeed, models computed at low values of $q$, which are very fast, match the precision of measurements from modern instrumentation.\\\\
+To demonstrate our point, we first consider the transmission spectroscopy of TRAPPIST-1 b performed with the JWST/NIRISS instrument (GO-2589, \citealt{Lim2023}).
+These observations reached an estimated precision of 89 ppm for a planet with a relative radius of the order $r\sim0.1$.
+In this context, numerical integration with an order as low as $q=10$ guarantees that our model reaches a precision of $10^{-13}$, an execution time on par with \textsf{exoplanet} on CPU and up to two orders of magnitude faster on GPU for large datasets (see left plot of \autoref{real_benchmark}).
+The second example we consider is the secondary ecclipse mapping of HD 189733b ($r\sim0.3$) using Spitzer's 8 micron channel \citep{majeau2012} obtained at a precision of about 1 ppt.
+For a spherical harmonics map of degree $l_{max}=3$ (although $l_{max}=1$ is enough in this case), $q=10$ guarantees a model with a precision of $10^{-10}$, an execution time one order of magnitude faster than \textsf{starry} on CPU and up to two orders of magnitude faster on GPU for large datasets (see right plot of \autoref{real_benchmark}).\\\\
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[width=\textwidth]{../workflows/speed/figures/speed_vs_N_realistic.pdf}
+        \caption{}
+        \label{fig:real_benchmark}
+    \end{center}
+\end{figure}
 

### 2024-10-15 workflows/speed/scripts/flux_exoplanet_quadratic.py

@@ -2,7 +2,7 @@ from exoplanet.light_curves import LimbDarkLightCurve
 import numpy as np
 from time import time
 
-radius = snakemake.params.radius
+radius = float(snakemake.wildcards.radius)
 u = snakemake.params.u
 N = int(snakemake.wildcards.N)
 assert radius < 1

### 2024-10-15 workflows/speed/scripts/flux_jaxoplanet.py

@@ -8,23 +8,29 @@ from jaxoplanet.experimental.starry import Surface, Ylm
 from jaxoplanet.experimental.starry.light_curves import surface_light_curve
 from time import time
 
-radius = float(snakemake.params.radius)
+radius = float(snakemake.wildcards.radius)
 u = snakemake.params.u
 N = int(snakemake.wildcards.N)
 assert radius < 1
 b = np.linspace(0, 1 + radius, N)
 order = int(snakemake.wildcards.order)
+lmax = snakemake.wildcards.lmax
+lmax = None if lmax == "None" else int(lmax)
 
-if u is not None:
-    surface = Surface(period=None, u=u)
+if lmax is None:
+    from jaxoplanet.core.limb_dark import light_curve
+
+    flux_function = jax.jit(jax.vmap(lambda b: light_curve(u, b, radius, order=order)))
 else:
-    l_max = 20
+    l_max = lmax
     y = Ylm.from_dense(np.ones((l_max + 1) ** 2))
-    surface = Surface(period=None, y=y)
+    surface = Surface(period=None, y=y, u=u)
 
-flux_function = jax.jit(
-    jax.vmap(lambda b: surface_light_curve(surface, r=radius, y=b, z=10.0, order=order))
-)
+    flux_function = jax.jit(
+        jax.vmap(
+            lambda b: surface_light_curve(surface, r=radius, y=b, z=10.0, order=order)
+        )
+    )
 
 times = []
 

### 2024-10-15 workflows/speed/scripts/flux_jaxoplanet_quadratic.py

@@ -1,29 +1,0 @@
-import jax
-
-jax.config.update("jax_platform_name", "cpu")
-jax.config.update("jax_enable_x64", True)
-
-import numpy as np
-from jaxoplanet.core.limb_dark import light_curve
-from time import time
-
-radius = float(snakemake.params.radius)
-u = snakemake.params.u
-N = int(snakemake.wildcards.N)
-assert radius < 1
-b = np.linspace(0, 1 + radius, N)
-order = int(snakemake.wildcards.order)
-
-
-flux_function = jax.jit(jax.vmap(lambda b: light_curve(u, b, radius, order=order)))
-flux_function(b).block_until_ready()
-
-times = []
-
-for _ in range(20):
-    t0 = time()
-    flux_function(b).block_until_ready()
-    times.append(time() - t0)
-
-
-np.savez(snakemake.output[0], time=times, N=N, b=b)

### 2024-10-15 workflows/speed/scripts/flux_starry.py

@@ -7,14 +7,15 @@ import theano.tensor as tt
 starry.config.lazy = True
 
 
-radius = snakemake.params.radius
+radius = float(snakemake.wildcards.radius)
 u = snakemake.params.u
 N = int(snakemake.wildcards.N)
 assert radius < 1
 b = np.linspace(0, 1 + radius, N)
+lmax = snakemake.wildcards.lmax
+lmax = None if lmax == "None" else int(lmax)
 
-
-if u is not None:
+if lmax is None:
     ms = starry.Map(ydeg=len(u), udeg=len(u))
     ms[1:] = u
 else:

### 2024-10-16 latex/paper.tex
feat: update

@@
 \subsection{Precision}\label{precision}
-We evaluate the precision of the \textsf{starry} and \textsf{jaxoplanet} terms against quantities computed at arbitrary precision, using a ground truth version of the code implemented with the \texttt{mpmath} Python library\footlink{https://mpmath.org/}.
+We evaluate the precision of the \textsf{starry} and \textsf{jaxoplanet} terms against quantities computed at arbitrary precision, using a ground truth version of the code implemented with the \texttt{mpmath} Python library\footnote{\href{https://mpmath.org/}{https://mpmath.org/}}.
 The precision of the overall flux $f$ can only be understood by considering the precision of the terms involved in \autoref{eq:starry}: the solution vector $\bvec{s}^\mathsf{T}$ ($\bvec{r}^\mathsf{T}$ if no occultation), the change of basis matrices $\bvec{A_1}$ and $\bvec{A_2}$, and the Wigner-D rotation matrix $\bvec{R}$.
 Although we show in \autoref{fig:precision_SAR} that the precision of the \textsf{starry} and \textsf{jaxoplanet} implentations slightly differ for many of these terms, we focus this section on the solution vector $\bvec{s}^\mathsf{T}$ and the overall flux $f$.\\\\
-Assuming that we perform the numerical integration described in \autoref{solution_vector} at a relatively high order $q=500$, \autoref{fig:precision_s} shows that our reimplentation of \textsf{starry} reaches relative errors lower than 1 part per billion for $\bvec{s}^\mathsf{T}$ and the overall flux $f$, considering both a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$.
+Assuming that we perform the numerical integration described in \autoref{solution_vector} at a relatively high order $q=500$, \autoref{fig:precision_s} shows that our reimplentation of \textsf{starry} reaches relative errors lower than 1 part per billion for $\bvec{s}^\mathsf{T}$ and the overall flux $f$, for almost all degrees up to $l_{max}=20$ and considering both a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$.
 For reference, \autoref{fig:precision_s_starry} shows comparable errors on the same quantities computed with the C++ implentation of \textsf{starry}.
@@
 In addition, \autoref{fig:precision_degree} shows the evolution of the error on $f$ for increasing degrees of the spherical harmonics for \textsf{starry} and \textsf{jaxoplanet}.
-With these results, we validate the precision of \textsf{jaxoplanet} and show it is on per with the legacy C++ \textsf{starry} implentation\footnote{version 1.2.0 \citep{starry_120}} described in \cite{starry}.
 \begin{figure}[H]
@@
 \end{figure}
-For reference, \autoref{fig:precision_limbdark} shows the relative error in the flux of a limb-darkened star occulted by an opaque body for increasing orders of a polynomial limb darkening law.
+In \autoref{fig:precision_limbdark}, we compare the precision of our limb darkened light curve model with \textsf{exoplanet} Python package (which uses the same formalism as \cite{Agol2020} but limited to linear and quadratic limb darkening).from shows the relative error in the flux of a limb-darkened star occulted by an opaque body for increasing orders of a polynomial limb darkening law.
 In this figure, we also report the error of the flux computed using the \textsf{exoplanet} Python package (which uses the same formalism as \cite{Agol2020} but limited to linear and quadratic limb darkening).
@@
 We note that the errors scale with the values of $\bvec{u}$.\\\\
+With these results, we validate the precision of \textsf{jaxoplanet} and show it is on per with the legacy C++ \textsf{starry} implentation\footnote{version 1.2.0 \citep{starry_120}} described in \cite{starry}.\\\\
 As described in \autoref{solution_vector}, we compute occultation light curves using the Gauss Legendre quadrature to approximate the $\mathcal{P}$ integral involved in the expression of the solution vector $\bvec{s}^\mathsf{T}$.
@@
 As discussed in \autoref{discussion}, \textsf{jaxoplanet} can be further optimized by reducing the order $q$ of numerical integration while keeping the precision required for specific applications.
+In this context, the \textsf{jaxoplanet} Python package provide helper functions to estimate the precision of the light curve model given the relative radius $r$, the spherical harmonics degree $l_{max}$ (if general non-uniform maps are requires), the order of the polynomial limb darkening, and the order $q$ of the numerical integration.\\\\
 
@@
 Although we expect to further optimize our \textsf{JAX} implementation and bridge this gap in the future, \textsf{jaxoplanet} can currently outperform state-of-the-art codes for most realistic datasets.
-Indeed, models computed at low values of $q$, which are very fast, match the precision of measurements from modern instrumentation.\\\\
+Indeed, models computed at low values of $q$, which are very fast, match the precision of measurements from modern instruments.\\\\
 To demonstrate our point, we first consider the transmission spectroscopy of TRAPPIST-1 b performed with the JWST/NIRISS instrument (GO-2589, \citealt{Lim2023}).
 These observations reached an estimated precision of 89 ppm for a planet with a relative radius of the order $r\sim0.1$.
-In this context, numerical integration with an order as low as $q=10$ guarantees that our model reaches a precision of $10^{-13}$, an execution time on par with \textsf{exoplanet} on CPU and up to two orders of magnitude faster on GPU for large datasets (see left plot of \autoref{real_benchmark}).
+In this context, numerical integration with an order as low as $q=10$ guarantees that our model reaches a precision of $10^{-16}$, an execution time on par with \textsf{exoplanet} on CPU and up to an order of magnitude faster on GPU for large datasets (see left plot of \autoref{fig:real_benchmark}).
 The second example we consider is the secondary ecclipse mapping of HD 189733b ($r\sim0.3$) using Spitzer's 8 micron channel \citep{majeau2012} obtained at a precision of about 1 ppt.
-For a spherical harmonics map of degree $l_{max}=3$ (although $l_{max}=1$ is enough in this case), $q=10$ guarantees a model with a precision of $10^{-10}$, an execution time one order of magnitude faster than \textsf{starry} on CPU and up to two orders of magnitude faster on GPU for large datasets (see right plot of \autoref{real_benchmark}).\\\\
+For a spherical harmonics map of degree $l_{max}=3$ (although $l_{max}=1$ is enough in this case), $q=10$ guarantees a model with a precision of $10^{-10}$, an execution time one order of magnitude faster than \textsf{starry} on CPU and up to two orders of magnitude faster on GPU for large datasets (see right plot of \autoref{fig:real_benchmark}).\\\\
 \begin{figure}[H]
@@
 As the core of our model is developed in Python, we expect \textsf{jaxoplanet} to serve as a basis for the implementation of more complex light curve models, such as light curves of Doppler maps \citep{Luger2021c}, reflected maps \citep{Luger2022}, or non-spherical bodies maps \citep{Dholakia2024}.
-Eventually, \textsf{jaxoplanet} will take over \textsf{starry} in acting as a foundation to explore and develop analytical methods to numerically reconstruct the surface of exoplanets and stars, venturing a bit further into the understanding of exoplanetary systems and their worlds.\\\\
+Eventually, \textsf{jaxoplanet} will take over \textsf{starry} in acting as a foundation to explore and develop analytical methods to numerically reconstruct the surface of exoplanets and stars, deepening our understanding of exoplanetary systems and their worlds.\\\\
 The \textsf{jaxoplanet} code is open source under the MIT License and is hosted at \href{https://github.com/exoplanet-dev/jaxoplanet/}{https://github.com/exoplanet-dev/jaxoplanet}, with documentation and tutorials hosted at \href{https://jax.exoplanet.codes}{https://jax.exoplanet.codes}.
@@
         \includegraphics[width=\textwidth]{../workflows/precision/figures/error_starry.pdf}
-        \caption{Same as \autoref{fig:precision_s} but with $\bvec{s}$ and $f$ computed with the C++ implementation of \textsf{starry}. \codelink{workflows/precision/scripts/plot_error}}
+        \caption{Same as \autoref{fig:precision_s} but with $\bvec{s}^\mathsf{T}$ and $f$ computed with the C++ implementation of \textsf{starry}. \codelink{workflows/precision/scripts/plot_error}}
         \label{fig:precision_s_starry}

### 2024-10-16 latex/paper.tex
feat: speed benchmark done on realistic examples

@@
 \end{figure}
-In \autoref{fig:precision_limbdark}, we compare the precision of our limb darkened light curve model with \textsf{exoplanet} Python package (which uses the same formalism as \cite{Agol2020} but limited to linear and quadratic limb darkening).from shows the relative error in the flux of a limb-darkened star occulted by an opaque body for increasing orders of a polynomial limb darkening law.
-In this figure, we also report the error of the flux computed using the \textsf{exoplanet} Python package (which uses the same formalism as \cite{Agol2020} but limited to linear and quadratic limb darkening).
+In \autoref{fig:precision_limbdark}, we compare the precision of our limb darkened light curve model with that of the \textsf{exoplanet} Python package (which uses the same formalism as \cite{Agol2020} but limited to linear and quadratic limb darkening).
+The larger errors observed in this figure are due to the transformation of the polynomial limb darkening coefficients to the spherical harmonics coefficients.
+For example, a set of unitary polynomial limb darkening coefficients of order 20 leads to spherical harmonics coefficients as large as $10^6$.
+We note that these errors scale with the values of $\bvec{u}$.\\\\
 \begin{figure}[H]
@@
 \end{figure}
-A set of unitary polynomial limb darkening coefficients of order 20 leads to spherical harmonics coefficients as large as $10^6$.
-This explains the larger errors observed in \autoref{fig:precision_limbdark} (error versus limb darkening order) compared to the ones shown in \autoref{fig:precision_degree} (error versus degree of the spherical harmonics).
-We note that the errors scale with the values of $\bvec{u}$.\\\\
-With these results, we validate the precision of \textsf{jaxoplanet} and show it is on per with the legacy C++ \textsf{starry} implentation\footnote{version 1.2.0 \citep{starry_120}} described in \cite{starry}.\\\\
 As described in \autoref{solution_vector}, we compute occultation light curves using the Gauss Legendre quadrature to approximate the $\mathcal{P}$ integral involved in the expression of the solution vector $\bvec{s}^\mathsf{T}$.
-Hence, the precision of the computed flux $f$ depends on the order of the Legendre polynomial which defines the number of points used to approximate $\mathcal{P}$.
-For this reason, users must carefully set the order parameter $q$ to reach the precision required for their application, as shown in \autoref{fig:relative_error_1}.\\\\
+Hence, the precision of the computed flux $f$ depends on the order of the Legendre polynomial which defines the number of points used to approximate $\mathcal{P}$ (see \autoref{fig:relative_error_1}).Hence, users must carefully set the order parameter to balance precision and computational speed (see \autoref{speed}).
 \begin{figure}[H]
@@
 % \end{figure}
-The minimum error is always achieved for higher values of $q$, which comes with increased computation time (see \autoref{speed}).
-Hence users must carefully set the order parameter to balance precision and computational speed.
-As the minimum error reached also depends on the degree of the spherical harmonics basis used, \autoref{fig:precision_order_gausslegendre_degree} shows the errors on the flux for each degree of the spherical harmonics basis, for an occultor with relative radii $r=0.01$, $r=1$ and $r=100$.
+
+With the results of this section, we validate the precision of \textsf{jaxoplanet} and show it is on per with the legacy C++ \textsf{starry} implentation\footnote{version 1.2.0 \citep{starry_120}} described in \cite{starry}.\\\\
+
+\subsection{Speed}\label{speed}
+
+One of the advantage of the \textsf{JAX} library is its ability to perform hardware-accelerated computations on CPUs and GPUs.
+In this section, we evaluate the speed of the \textsf{jaxoplanet} implementation and compare it with the C++ implementation of \textsf{starry}.
+For quadratically limb darkened light curves, we also compare our implementation with the one provided by the \textsf{exoplanet} Python package.\\\\
+While performing numerical integration may seem as the downside of our implementation, adapting the value of $q$ to the precision required for specific applications allows users to optimize the computation time of the model to their advantage.
+To demonstrate this point, we first consider the transmission spectroscopy of TRAPPIST-1 b performed with the JWST/NIRISS instrument (GO-2589, \citealt{Lim2023}).
+These observations reached an estimated precision of 89 ppm for a planet with a relative radius of the order $r\sim0.1$.
+In this context, numerical integration with an order as low as $q=10$ guarantees that our model reaches a precision of $10^{-16}$, an execution time on par with \textsf{exoplanet} on CPU and up to an order of magnitude faster on GPU for large datasets (see left plot of \autoref{fig:real_benchmark}).
+The second example we consider is the secondary ecclipse mapping of HD 189733b ($r\sim0.3$) using Spitzer's 8 micron channel \citep{majeau2012} obtained at a precision of about 1 ppt.
+For a spherical harmonics map of degree $l_{max}=3$ (although $l_{max}=1$ is enough in this case), $q=10$ guarantees a model with a precision of $10^{-10}$, an execution time one order of magnitude faster than \textsf{starry} on CPU and up to two orders of magnitude faster on GPU for large datasets (see right plot of \autoref{fig:real_benchmark}).\\\\
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[width=\textwidth]{../workflows/speed/figures/speed_vs_N_realistic.pdf}
+        \caption{}
+        \label{fig:real_benchmark}
+    \end{center}
+\end{figure}
+
+In general, we show that models computed at low values of $q$, which are very fast, match the precision of measurements from modern instruments.
+Although this might not be required, \autoref{precise_speed} shows reference speed benchmarks for models evaluated at state-of-the-art precision.\\\\
+In order to help users set the value of $q$ according to their required precision, the \textsf{jaxoplanet} Python package provide helper functions to estimate the precision of the light curve model given the relative radius $r$, the spherical harmonics degree $l_{max}$ (if general non-uniform maps are required), and the order of the polynomial limb darkening law\footnote{See tutorial at \href{}{}}.\\\\
+
+\section{Conclusion}
+In this paper we presented \textsf{jaxoplanet}, a new implementation of the analytical light curve model \textit{starry} developed using the machine learning library \textsf{JAX}.
+We showed that our implementation is on par with the legacy C++ \textsf{starry} implementation, and that it offers a significant speed advantage on GPUs for large datasets, enabled by \textsf{JAX}'s just-in-time compilation and hardware acceleration.
+Our code can be used in association with the large ecosystem of probabilistic and sampling libraries developed in \textsf{JAX}, and will facilitate the development of tools and pipeline for the modeling of TESS and JWST observations, but also future missions such as PLATO \citep{Rauer2014} and ARIEL \citep{Tinetti2018}.\\\\
+As the core of our model is developed in Python, we expect \textsf{jaxoplanet} to serve as a basis for the implementation of more complex light curve models, such as light curves of Doppler maps \citep{Luger2021c}, reflected maps \citep{Luger2022}, or non-spherical bodies maps \citep{Dholakia2024}.
+Eventually, \textsf{jaxoplanet} will take over \textsf{starry} in acting as a foundation to explore and develop analytical methods to numerically reconstruct the surface of exoplanets and stars, deepening our understanding of exoplanetary systems and their worlds.\\\\
+The \textsf{jaxoplanet} code is open source under the MIT License and is hosted at \href{https://github.com/exoplanet-dev/jaxoplanet/}{https://github.com/exoplanet-dev/jaxoplanet}, with documentation and tutorials hosted at \href{https://jax.exoplanet.codes}{https://jax.exoplanet.codes}.
+A permanent version of the code used to generate the figures and results in this paper is archived at https://doi.org/10.5281/zenodo.1312286.
+
+\bibliography{ref}
+
+\appendix
+
+\section{Precision}\label{precise_precision}
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[width=\textwidth]{../workflows/precision/figures/error_SAR.pdf}
+        \caption{Typical errors in the matrices used to compute the flux (\autoref{eq:starry}) with \textsf{starry} and \textsf{jaxoplanet}. \codelink{workflows/precision/scripts/plot_error_SAR}}
+        \label{fig:precision_SAR}
+    \end{center}
+\end{figure}
+
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[width=\textwidth]{../workflows/precision/figures/error_starry.pdf}
+        \caption{Same as \autoref{fig:precision_s} but with $\bvec{s}^\mathsf{T}$ and $f$ computed with the C++ implementation of \textsf{starry}. \codelink{workflows/precision/scripts/plot_error}}
+        \label{fig:precision_s_starry}
+    \end{center}
+\end{figure}
+
 \begin{figure}[H]
@@
 \end{figure}
-\subsection{Speed}\label{speed}
 
-One of the advantage of the \textsf{JAX} library is its ability to perform hardware-accelerated computations on CPUs and GPUs.
-In this section, we evaluate the speed of the \textsf{jaxoplanet} implementation and compare it with the C++ implementation of \textsf{starry}.
-For quadratically limb darkened light curves, we also compare our implementation with the one provided by the \textsf{exoplanet} Python package.\\\\
+\section{Speed}\label{precise_speed}
 In \autoref{fig:speed_order_gausslegendre}, we show the evolution of the time required to compute the occultation light curve of a limb darkened star and that of a non-uniform star described by spherical harmonics of degree $l_{max}=20$, depending on the order $q$ of the numerical integration.
@@
 As discussed in \autoref{discussion}, \textsf{jaxoplanet} can be further optimized by reducing the order $q$ of numerical integration while keeping the precision required for specific applications.
-In this context, the \textsf{jaxoplanet} Python package provide helper functions to estimate the precision of the light curve model given the relative radius $r$, the spherical harmonics degree $l_{max}$ (if general non-uniform maps are requires), the order of the polynomial limb darkening, and the order $q$ of the numerical integration.\\\\
-
-\section{Discussion}\label{discussion}
-
-In \autoref{precision}, we showed that \textsf{jaxoplanet} is on par with the C++ implementation of \textsf{starry} in terms of precision.
-As relatively high values of $q$ are required to achieve this performance, results from \autoref{speed} are deceiving and suggest that \textsf{jaxoplanet} is about an order of magnitude slower than \textsf{exoplanet} or \textsf{starry} on CPU, and only advantageous on GPU for very large datasets.
-Although we expect to further optimize our \textsf{JAX} implementation and bridge this gap in the future, \textsf{jaxoplanet} can currently outperform state-of-the-art codes for most realistic datasets.
-Indeed, models computed at low values of $q$, which are very fast, match the precision of measurements from modern instruments.\\\\
-To demonstrate our point, we first consider the transmission spectroscopy of TRAPPIST-1 b performed with the JWST/NIRISS instrument (GO-2589, \citealt{Lim2023}).
-These observations reached an estimated precision of 89 ppm for a planet with a relative radius of the order $r\sim0.1$.
-In this context, numerical integration with an order as low as $q=10$ guarantees that our model reaches a precision of $10^{-16}$, an execution time on par with \textsf{exoplanet} on CPU and up to an order of magnitude faster on GPU for large datasets (see left plot of \autoref{fig:real_benchmark}).
-The second example we consider is the secondary ecclipse mapping of HD 189733b ($r\sim0.3$) using Spitzer's 8 micron channel \citep{majeau2012} obtained at a precision of about 1 ppt.
-For a spherical harmonics map of degree $l_{max}=3$ (although $l_{max}=1$ is enough in this case), $q=10$ guarantees a model with a precision of $10^{-10}$, an execution time one order of magnitude faster than \textsf{starry} on CPU and up to two orders of magnitude faster on GPU for large datasets (see right plot of \autoref{fig:real_benchmark}).\\\\
-\begin{figure}[H]
-    \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/speed/figures/speed_vs_N_realistic.pdf}
-        \caption{}
-        \label{fig:real_benchmark}
-    \end{center}
-\end{figure}
-
-\section{Conclusion}
-In this paper we presented \textsf{jaxoplanet}, a new implementation of the analytical light curve model \textit{starry} developed using the machine learning library \textsf{JAX}.
-We showed that our implementation is on par with the legacy C++ \textsf{starry} implementation, and that it offers a significant speed advantage on GPUs for large datasets, enabled by \textsf{JAX}'s just-in-time compilation and hardware acceleration.
-Our code can be used in association with the large ecosystem of probabilistic and sampling libraries developed in \textsf{JAX}, and will facilitate the development of tools and pipeline for the modeling of TESS and JWST observations, but also future missions such as PLATO \citep{Rauer2014} and ARIEL \citep{Tinetti2018}.\\\\
-As the core of our model is developed in Python, we expect \textsf{jaxoplanet} to serve as a basis for the implementation of more complex light curve models, such as light curves of Doppler maps \citep{Luger2021c}, reflected maps \citep{Luger2022}, or non-spherical bodies maps \citep{Dholakia2024}.
-Eventually, \textsf{jaxoplanet} will take over \textsf{starry} in acting as a foundation to explore and develop analytical methods to numerically reconstruct the surface of exoplanets and stars, deepening our understanding of exoplanetary systems and their worlds.\\\\
-The \textsf{jaxoplanet} code is open source under the MIT License and is hosted at \href{https://github.com/exoplanet-dev/jaxoplanet/}{https://github.com/exoplanet-dev/jaxoplanet}, with documentation and tutorials hosted at \href{https://jax.exoplanet.codes}{https://jax.exoplanet.codes}.
-A permanent version of the code used to generate the figures and results in this paper is archived at https://doi.org/10.5281/zenodo.1312286.
-
-\bibliography{ref}
-
-\appendix
-
-\section{Error budget}
-\begin{figure}[H]
-    \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/precision/figures/error_SAR.pdf}
-        \caption{\codelink{workflows/precision/scripts/plot_error_SAR}}
-        \label{fig:precision_SAR}
-    \end{center}
-\end{figure}
-
-\begin{figure}[H]
-    \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/precision/figures/error_starry.pdf}
-        \caption{Same as \autoref{fig:precision_s} but with $\bvec{s}^\mathsf{T}$ and $f$ computed with the C++ implementation of \textsf{starry}. \codelink{workflows/precision/scripts/plot_error}}
-        \label{fig:precision_s_starry}
-    \end{center}
-\end{figure}
 

### 2024-10-16 latex/paper.tex
feat: add kepler's equation solver

@@
 \subsection{Kepler's equation solver}
-\todo{}
+A core difference between our implementation and that of \cite{starry} and \cite{exoplanet} lies in the method used to solve Kepler's equation
+\begin{equation}
+    M = E - e \sin(E)
+\end{equation}
+Indeed, iterative methods used in previous implementations (such as \citealt{Raposo2017} and \citealt{Brandt2021}) cannot be easily implemented in \textsf{JAX} and are not extensible to GPU or TPU acceleration.
+Instead, our implementation makes use of \cite{Markley1995} which leads to state-of-the-art precision and runtime performance.\\\\
+Although JAX features automatic differentiation, we find that explicitly implementing the derivatives of the Kepler equation solver is more efficient.
+To do so, we use the implicit function theorem to derive the “Jacobian-vector product” required by \textsf{JAX} to specify custom derivatives.
+This is done by differentiating Kepler's equation as
+\begin{equation}
+    \mathrm{d}M = \mathrm{d}E\,(1 - e \cos E) - \mathrm{d}e\;\sin E,
+\end{equation}
+which leads to
+\begin{equation}
+\mathrm{d}E = \frac{1}{1 - e \cos E}\;\mathrm{d}M + \frac{\sin E}{1 - e \cos E}\;\mathrm{d}e.
+\end{equation}
+Then, for a given value of $E$, its derivatives can be easily evaluated.
+We find that this method is 6 times faster than using \textsf{JAX}'s automatic differentiation\footnote{Test performed on an M2 Max Macbook.}.\\\\
 \subsection{Light curve functions and transformations}

### 2024-10-17 workflows/precision/scripts/concatenate_errors_GL.py
refactor: with starry out of experimental

@@ -4,7 +4,7 @@ jax.config.update("jax_enable_x64", True)
 
 import numpy as np
 import matplotlib.pyplot as plt
-from jaxoplanet.experimental.starry.multiprecision import utils, mp
+from jaxoplanet.starry.multiprecision import utils, mp
 from tqdm import tqdm
 
 rs = snakemake.params.rs

### 2024-10-17 workflows/precision/scripts/concatenate_fs.py

@@ -1,7 +1,7 @@
 import pickle
 import numpy as np
 from tqdm import tqdm
-from jaxoplanet.experimental.starry.multiprecision import mp
+from jaxoplanet.starry.multiprecision import mp
 
 flux = []
 s = []

### 2024-10-17 workflows/precision/scripts/jax_f_nonuniform_limbdark.py

@@ -2,8 +2,8 @@ import jax
 
 jax.config.update("jax_enable_x64", True)
 
-from jaxoplanet.experimental.starry import Surface
-from jaxoplanet.experimental.starry.light_curves import surface_light_curve
+from jaxoplanet.starry import Surface
+from jaxoplanet.starry.light_curves import surface_light_curve
 import numpy as np
 from tqdm import tqdm
 

### 2024-10-17 workflows/precision/scripts/jax_fs.py

@@ -2,10 +2,10 @@ import jax
 
 jax.config.update("jax_enable_x64", True)
 
-from jaxoplanet.experimental.starry.solution import solution_vector
-from jaxoplanet.experimental.starry.light_curves import surface_light_curve
-from jaxoplanet.experimental.starry.surface import Surface
-from jaxoplanet.experimental.starry.ylm import Ylm
+from jaxoplanet.starry.solution import solution_vector
+from jaxoplanet.starry.light_curves import surface_light_curve
+from jaxoplanet.starry.surface import Surface
+from jaxoplanet.starry.ylm import Ylm
 import numpy as np
 import jax.numpy as jnp
 from functools import partial

### 2024-10-17 workflows/precision/scripts/jax_fs_GL.py

@@ -3,14 +3,14 @@ import jax
 
 jax.config.update("jax_enable_x64", True)
 
-from jaxoplanet.experimental.starry.solution import solution_vector
-from jaxoplanet.experimental.starry.light_curves import *
+from jaxoplanet.starry.solution import solution_vector
+from jaxoplanet.starry.light_curves import *
 import numpy as np
 import jax.numpy as jnp
-from jaxoplanet.experimental.starry.solution import rT
-from jaxoplanet.experimental.starry.mpcore.basis import A1
-from jaxoplanet.experimental.starry.basis import A2_inv
-from jaxoplanet.experimental.starry.mpcore.utils import to_numpy
+from jaxoplanet.starry.solution import rT
+from jaxoplanet.starry.mpcore.basis import A1
+from jaxoplanet.starry.core.basis import A2_inv
+from jaxoplanet.starry.mpcore.utils import to_numpy
 
 
 def jax_flux(deg, order=20, inc=np.pi / 2, obl=0.0):

### 2024-10-17 workflows/precision/scripts/num_f_limbdark.py

@@ -1,8 +1,8 @@
 import numpy as np
 import pickle
-from jaxoplanet.experimental.starry.multiprecision.utils import mp
-from jaxoplanet.experimental.starry import basis
-from jaxoplanet.experimental.starry.multiprecision import solution
+from jaxoplanet.starry.multiprecision.utils import mp
+from jaxoplanet.starry import basis
+from jaxoplanet.starry.multiprecision import solution
 from tqdm import tqdm
 
 # some params

### 2024-10-17 workflows/precision/scripts/num_fs.py

@@ -1,6 +1,6 @@
 import numpy as np
-from jaxoplanet.experimental.starry.multiprecision import mp
-from jaxoplanet.experimental.starry.multiprecision.flux import flux_function
+from jaxoplanet.starry.multiprecision import mp
+from jaxoplanet.starry.multiprecision.flux import flux_function
 
 r = float(snakemake.wildcards.r)
 l_max = snakemake.params.l_max

### 2024-10-17 workflows/precision/scripts/num_fs_matrices.py

@@ -1,8 +1,8 @@
 l_max = snakemake.params.l_max
 
-from jaxoplanet.experimental.starry.multiprecision import mp
-from jaxoplanet.experimental.starry.multiprecision.basis import A1, A2
-from jaxoplanet.experimental.starry.multiprecision.rotation import get_R
+from jaxoplanet.starry.multiprecision import mp
+from jaxoplanet.starry.multiprecision.basis import A1, A2
+from jaxoplanet.starry.multiprecision.rotation import get_R
 from collections import defaultdict
 
 inc = mp.pi / 2

### 2024-10-17 workflows/precision/scripts/num_sT.py

@@ -1,4 +1,4 @@
-from jaxoplanet.experimental.starry.multiprecision.solution import sT
+from jaxoplanet.starry.multiprecision.solution import sT
 import numpy as np
 from tqdm import tqdm
 

### 2024-10-17 workflows/precision/scripts/plot_error.py

@@ -9,8 +9,8 @@ jax.config.update("jax_enable_x64", True)
 import matplotlib.pyplot as plt
 from matplotlib import gridspec
 import numpy as np
-from jaxoplanet.experimental.starry.ylm import Ylm
-from jaxoplanet.experimental.starry.multiprecision import mp, utils
+from jaxoplanet.starry.ylm import Ylm
+from jaxoplanet.starry.multiprecision import mp, utils
 
 ref = "num"
 comp = "comp"
@@ -106,7 +106,7 @@ l_max = snakemake.params.l_max
 
 cmap = plt.get_cmap("plasma")
 
-fig = plt.figure(constrained_layout=True, figsize=(8.0, 7))
+fig = plt.figure(constrained_layout=True, figsize=(8.0, 6))
 
 N = 8
 gs = gridspec.GridSpec(3, N, figure=fig, height_ratios=[1, 2, 2])

### 2024-10-17 workflows/precision/scripts/plot_error_SAR.py

@@ -3,8 +3,8 @@ import jax
 jax.config.update("jax_enable_x64", True)
 
 import numpy as np
-from jaxoplanet.experimental.starry.multiprecision import mp
-from jaxoplanet.experimental.starry.multiprecision.utils import diff_mp, to_numpy
+from jaxoplanet.starry.multiprecision import mp
+from jaxoplanet.starry.multiprecision.utils import diff_mp, to_numpy
 
 l_max = snakemake.params.l_max
 

### 2024-10-17 workflows/precision/scripts/plot_error_limbdark.py

@@ -4,7 +4,7 @@ jax.config.update("jax_enable_x64", True)
 
 import matplotlib.pyplot as plt
 import numpy as np
-from jaxoplanet.experimental.starry.multiprecision import mp, utils
+from jaxoplanet.starry.multiprecision import mp, utils
 
 degree = snakemake.params.degree
 q_order = int(snakemake.params.order)

### 2024-10-17 workflows/precision/scripts/plot_error_several_r_GL_degree_SH.py

@@ -2,7 +2,7 @@
 Gauss-Legendre quadrature, and spherical harmonic degrees.
 """
 
-from jaxoplanet.experimental.starry import Ylm
+from jaxoplanet.starry import Ylm
 import matplotlib.pyplot as plt
 import numpy as np
 

### 2024-10-17 workflows/precision/scripts/plot_error_vs_degree.py

@@ -3,7 +3,7 @@ import jax
 jax.config.update("jax_enable_x64", True)
 import matplotlib.pyplot as plt
 import numpy as np
-from jaxoplanet.experimental.starry.multiprecision import mp, utils
+from jaxoplanet.starry.multiprecision import mp, utils
 
 l_max = snakemake.params.l_max
 
@@ -41,7 +41,7 @@ def errors(M_mp, M_np):
     return frac
 
 
-from jaxoplanet.experimental.starry.ylm import Ylm
+from jaxoplanet.starry.ylm import Ylm
 
 indices = {
     l: np.array([Ylm.index(l, m) for m in range(-l, l + 1)]) for l in range(l_max + 1)

### 2024-10-17 workflows/speed/scripts/flux_all_single_b.py

@@ -5,8 +5,8 @@ jax.config.update("jax_enable_x64", True)
 
 from exoplanet.light_curves import LimbDarkLightCurve
 from jaxoplanet.core.limb_dark import light_curve
-from jaxoplanet.experimental.starry.light_curves import surface_light_curve
-from jaxoplanet.experimental.starry import Surface, Ylm
+from jaxoplanet.starry.light_curves import surface_light_curve
+from jaxoplanet.starry import Surface, Ylm
 import numpy as np
 import starry
 from tqdm import tqdm
@@ -20,9 +20,9 @@ b = 1 - radius
 orders = snakemake.params.orders
 y_20 = np.ones(21**2)
 
-from jaxoplanet.experimental.starry.multiprecision.flux import flux_function
-from jaxoplanet.experimental.starry.multiprecision import mp, utils
-from jaxoplanet.experimental.starry.basis import U
+from jaxoplanet.starry.multiprecision.flux import flux_function
+from jaxoplanet.starry.multiprecision import mp, utils
+from jaxoplanet.starry.core.basis import U
 
 f_q = flux_function(3**2, mp.pi / 2, 0.0)
 y_q = mp.matrix([[1.0, 0.1, 0.1]]) @ utils.to_mp(U(2))

### 2024-10-17 workflows/speed/scripts/flux_jaxoplanet.py

@@ -4,8 +4,8 @@ jax.config.update("jax_platform_name", "cpu")
 jax.config.update("jax_enable_x64", True)
 
 import numpy as np
-from jaxoplanet.experimental.starry import Surface, Ylm
-from jaxoplanet.experimental.starry.light_curves import surface_light_curve
+from jaxoplanet.starry import Surface, Ylm
+from jaxoplanet.starry.light_curves import surface_light_curve
 from time import time
 
 radius = float(snakemake.wildcards.radius)

### 2024-10-17 workflows/precision/scripts/SAR_jax_starry.py
update

@@ -11,7 +11,7 @@ import jax.numpy as jnp
 from jaxoplanet.starry.s2fft_rotation import compute_rotation_matrices
 from jaxoplanet.starry.multiprecision.rotation import R
 
-from jaxoplanet.starry.solution import solution_vector
+from jaxoplanet.starry.core.solution import solution_vector
 from jaxoplanet.starry.multiprecision.solution import sT
 from jaxoplanet.starry.multiprecision.utils import to_numpy
 from jaxoplanet.starry.multiprecision import basis as mbasis

### 2024-10-17 workflows/precision/scripts/jax_fs_GL.py

@@ -3,11 +3,11 @@ import jax
 
 jax.config.update("jax_enable_x64", True)
 
-from jaxoplanet.starry.solution import solution_vector
+from jaxoplanet.starry.core.solution import solution_vector
 from jaxoplanet.starry.light_curves import *
 import numpy as np
 import jax.numpy as jnp
-from jaxoplanet.starry.solution import rT
+from jaxoplanet.starry.core.solution import rT
 from jaxoplanet.starry.mpcore.basis import A1
 from jaxoplanet.starry.core.basis import A2_inv
 from jaxoplanet.starry.mpcore.utils import to_numpy

### 2024-10-17 workflows/precision/scripts/num_fs.py

@@ -1,6 +1,6 @@
 import numpy as np
 from jaxoplanet.starry.multiprecision import mp
-from jaxoplanet.starry.multiprecision.flux import flux_function
+from jaxoplanet.starry.multiprecision.flux import flux
 
 r = float(snakemake.wildcards.r)
 l_max = snakemake.params.l_max
@@ -25,7 +25,7 @@ def is_occ(b, r):
 if not np.all(is_occ(bs, r)):
     raise ValueError("Impact parameter out of bounds")
 
-func = flux_function(l_max, mp.pi / 2, 0.0, cache=num_matrices)
+func = flux(ydeg=l_max, cache=num_matrices)
 
 from tqdm import tqdm
 
@@ -33,7 +33,7 @@ results = []
 sT = []
 
 for b in tqdm(bs):
-    results.append(func(y, b, r, 0.0))
+    results.append(func(b, r, y=y))
     sT.append(num_matrices[l_max]["sT"][(b, r)])
 
 import pickle

### 2024-10-17 latex/paper.tex

@@
 \begin{abstract}
-    Modeling light curves of extrasolar systems serves a wide variety of science cases.
-Transit and occultation light curves, for example, are used to characterize their orbital configuration, and infer the atmospheric properties of their transiting bodies.
-Phase curves, on the other hand, can be used to study the thermal emission of exoplanets, and the atmospheric circulation of stars.
-In order to infer these properties, a fast, accurate and robust model of stellar light curves is needed, one that accounts for the non-uniform surface intensity of spherical bodies and their mutual occultations.
-In this context, the analytical framework known as \textsf{starry} provided the community with a powerful tool, instrumental in ....
-However, as datasets and models complexity grow, the tractable inference of these parameters becomes increasingly challenging, and often motivates biased approximations.
-In this paper, we present a new implementation of the \textsf{starry} framework in \texttt{jax}, a high-performance machine-learning library designed for large-scale computations.
+    Modeling light curves of celestial bodies serves a wide variety of science cases.
+From the orbital characterization of stellar systems to the study of exoplanet and stellar atmospheres.
+In this context, the analytical framework known as \textsf{starry} provided the community with a powerful tool to analytically compute light curves of bodies with non-uniform intensities.
+However, as datasets and models complexity grow, the tractable inference of their physical parameters becomes increasingly challenging, and often motivates biased approximations.
+In this paper, we present a new implementation of the \textsf{starry} framework in \textsf{JAX}, a high-performance machine-learning library designed for large-scale computations.
 We describe the changes made to the original version of \textsf{starry} and evaluate the performance of its new implementation that we release in the open-source Python package \textsf{jaxoplanet}.
@@
 
-\newpage
 \section{Implementation details}\label{optimization}
@@
 Although JAX features automatic differentiation, we find that explicitly implementing the derivatives of the Kepler equation solver is more efficient.
-To do so, we use the implicit function theorem to derive the “Jacobian-vector product” required by \textsf{JAX} to specify custom derivatives.
+To do so, we use the implicit function theorem to derive the \textit{Jacobian-vector product} required by \textsf{JAX} to specify custom derivatives.
 This is done by differentiating Kepler's equation as
@@
 \begin{equation}
-\mathrm{d}E = \frac{1}{1 - e \cos E}\;\mathrm{d}M + \frac{\sin E}{1 - e \cos E}\;\mathrm{d}e.
+\mathrm{d}E = \frac{1}{1 - e \cos E}\;\mathrm{d}M + \frac{\sin E}{1 - e \cos E}\;\mathrm{d}e,
 \end{equation}
-Then, for a given value of $E$, its derivatives can be easily evaluated.
+so that point estimates for the derivatives of $E$ can be easily evaluated.
 We find that this method is 6 times faster than using \textsf{JAX}'s automatic differentiation\footnote{Test performed on an M2 Max Macbook.}.\\\\
 \subsection{Light curve functions and transformations}
-\todo{}
+Using all the optimizations described in the previous sections, the \textsf{jaxoplanet} Python package provides code to compute light curves of mutually occulting celestial bodies.
+Our implementation is written in \textsf{JAX} and contains fully vectorized light curve functions that can be compiled and run on several hardware accelerator.
+\begin{itemize}
+    \item The \texttt{jaxoplanet.light\_curves.limbdark} module contains functions to compute  occultation light curves of limb darkened stars described by polynomial limb darkening laws \citep{Agol2020}.
+    \item The \texttt{jaxoplanet.light\_curves.emission} module contains functions to compute light curves of emissive limb darkened bodies.
+This module usess the same formalism as the \texttt{limbdark} module, but allows for secondary ecclipses to be computed.
+    \item The \texttt{jaxoplanet.starry.light\_curves} module contains functions to compute occultation light curve of bodies with non-uniform surfaces \citep{starry}.
+\end{itemize}
+These functions can optionally be transformed using special wrappers provided in the \texttt{jaxoplanet.light\_curves.transforms}  module.
+For example, the \texttt{integrate} transform enables the computation of light curves observed at finite exposure time, using the methods described in \cite{Kipping2010}\footnote{See tutorial at}.
+Additionlly, the \texttt{interpolate} transform can be used to precompute light curves on a grid near a transit, and then interpolate those computations to the required phases when computing the full model, leading to substantial speed ups when several transits are present in the data.\\\\
+By providing the community with an alternative to \textsf{starry} entirely written in Python, we hope to foster the development of new tools and methods to model light curves from stellar and planetary systems.
+In the next section, we evaluate the performance of our implementation and compare it with the original \textsf{starry} implementation.\\\\
 
@@
 Although we show in \autoref{fig:precision_SAR} that the precision of the \textsf{starry} and \textsf{jaxoplanet} implentations slightly differ for many of these terms, we focus this section on the solution vector $\bvec{s}^\mathsf{T}$ and the overall flux $f$.\\\\
-Assuming that we perform the numerical integration described in \autoref{solution_vector} at a relatively high order $q=500$, \autoref{fig:precision_s} shows that our reimplentation of \textsf{starry} reaches relative errors lower than 1 part per billion for $\bvec{s}^\mathsf{T}$ and the overall flux $f$, for almost all degrees up to $l_{max}=20$ and considering both a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$.
-For reference, \autoref{fig:precision_s_starry} shows comparable errors on the same quantities computed with the C++ implentation of \textsf{starry}.
 \begin{figure}[H]
@@
 \end{figure}
+Assuming that we perform the numerical integration described in \autoref{solution_vector} at a relatively high order $q=500$, \autoref{fig:precision_s} shows that our reimplentation of \textsf{starry} reaches relative errors lower than 1 part per billion for $\bvec{s}^\mathsf{T}$ and the overall flux $f$, for almost all degrees up to $l_{max}=20$ and considering both a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$.
+For reference, \autoref{fig:precision_s_starry} shows comparable errors on the same quantities computed with the C++ implentation of \textsf{starry}.
+\\\\
 In addition, \autoref{fig:precision_degree} shows the evolution of the error on $f$ for increasing degrees of the spherical harmonics for \textsf{starry} and \textsf{jaxoplanet}.
@@
 \end{figure}
-In \autoref{fig:precision_limbdark}, we compare the precision of our limb darkened light curve model with that of the \textsf{exoplanet} Python package (which uses the same formalism as \cite{Agol2020} but limited to linear and quadratic limb darkening).
-The larger errors observed in this figure are due to the transformation of the polynomial limb darkening coefficients to the spherical harmonics coefficients.
-For example, a set of unitary polynomial limb darkening coefficients of order 20 leads to spherical harmonics coefficients as large as $10^6$.
-We note that these errors scale with the values of $\bvec{u}$.\\\\
 \begin{figure}[H]
@@
 \end{figure}
+In \autoref{fig:precision_limbdark}, we compare the precision of our limb darkened light curve model with that of the \textsf{exoplanet} Python package (which uses the same formalism as \cite{Agol2020} but limited to linear and quadratic limb darkening).
+The larger errors observed in this figure are due to the transformation of the polynomial limb darkening coefficients to the spherical harmonics coefficients.
+For example, a set of unitary polynomial limb darkening coefficients of order 20 leads to spherical harmonics coefficients as large as $10^6$.
+We note that these errors scale with the values of $\bvec{u}$.\\\\
 As described in \autoref{solution_vector}, we compute occultation light curves using the Gauss Legendre quadrature to approximate the $\mathcal{P}$ integral involved in the expression of the solution vector $\bvec{s}^\mathsf{T}$.
-Hence, the precision of the computed flux $f$ depends on the order of the Legendre polynomial which defines the number of points used to approximate $\mathcal{P}$ (see \autoref{fig:relative_error_1}).Hence, users must carefully set the order parameter to balance precision and computational speed (see \autoref{speed}).
+Hence, the precision of the computed flux $f$ depends on the order of the Legendre polynomial which defines the number of points used to approximate $\mathcal{P}$ (see \autoref{fig:relative_error_1}, or  \autoref{fig:precision_order_gausslegendre_degree} in the Appendix to see the effect for different spherical harmonics degrees $l$).
+Hence, users must carefully set the order parameter $q$ to balance precision and computational speed (see \autoref{speed}).
 \begin{figure}[H]
@@
 For quadratically limb darkened light curves, we also compare our implementation with the one provided by the \textsf{exoplanet} Python package.\\\\
-While performing numerical integration may seem as the downside of our implementation, adapting the value of $q$ to the precision required for specific applications allows users to optimize the computation time of the model to their advantage.
+While performing numerical integration may be seen as a downside in our implementation, adapting the value of $q$ to the precision required for specific applications allows users to optimize the computation time of the model to their advantage.
 To demonstrate this point, we first consider the transmission spectroscopy of TRAPPIST-1 b performed with the JWST/NIRISS instrument (GO-2589, \citealt{Lim2023}).
@@
 In this context, numerical integration with an order as low as $q=10$ guarantees that our model reaches a precision of $10^{-16}$, an execution time on par with \textsf{exoplanet} on CPU and up to an order of magnitude faster on GPU for large datasets (see left plot of \autoref{fig:real_benchmark}).
-The second example we consider is the secondary ecclipse mapping of HD 189733b ($r\sim0.3$) using Spitzer's 8 micron channel \citep{majeau2012} obtained at a precision of about 1 ppt.
-For a spherical harmonics map of degree $l_{max}=3$ (although $l_{max}=1$ is enough in this case), $q=10$ guarantees a model with a precision of $10^{-10}$, an execution time one order of magnitude faster than \textsf{starry} on CPU and up to two orders of magnitude faster on GPU for large datasets (see right plot of \autoref{fig:real_benchmark}).\\\\
+The second example we consider is the secondary ecclipse mapping of HD189733 b ($r\sim0.3$) using Spitzer's 8 micron channel \citep{majeau2012} obtained at a precision of about 1 ppt.
+For a spherical harmonics map of degree $l_{max}=3$, $q=10$ guarantees a model with a precision of $10^{-10}$, an execution time one order of magnitude faster than \textsf{starry} on CPU and up to two orders of magnitude faster on GPU for large datasets (see right plot of \autoref{fig:real_benchmark}).\\\\
 \begin{figure}[H]
@@
         \includegraphics[width=\textwidth]{../workflows/speed/figures/speed_vs_N_realistic.pdf}
-        \caption{}
+        \caption{Evaluation time versus number of points for different light curves models used in realistic applications.
+On the left, we evaluate the time required by \textsf{jaxoplanet} and \textsf{exoplanet} to compute the quadratically limb darkened transit light curve of TRAPPIST-1 b observed with the JWST/NIRISS instrument (for which $q=10$ yields machine precision).
+On the right we evaluate the time required by \textsf{jaxoplanet} and \textsf{starry} to compute the eclipse light curve of HD189733 b whose surface is modeled using spherical harmonics of degree $l_{max}=3$ (with $q=10$ leading to a precision of $10^{-10}$).
+On CPU, our benchmark is done on the single core of an Apple M2 max chip, while on GPU we use the single core of an NVIDIA Tesla V100 processor.}
         \label{fig:real_benchmark}
@@
 In general, we show that models computed at low values of $q$, which are very fast, match the precision of measurements from modern instruments.
-Although this might not be required, \autoref{precise_speed} shows reference speed benchmarks for models evaluated at state-of-the-art precision.\\\\
+Although this might not be required, \autoref{precise_speed} shows reference speed benchmarks for models evaluated at state-of-the-art precisions.\\\\
 In order to help users set the value of $q$ according to their required precision, the \textsf{jaxoplanet} Python package provide helper functions to estimate the precision of the light curve model given the relative radius $r$, the spherical harmonics degree $l_{max}$ (if general non-uniform maps are required), and the order of the polynomial limb darkening law\footnote{See tutorial at \href{}{}}.\\\\
@@
 
-\section{Precision}\label{precise_precision}
+\section{Additional precision benchmarks}\label{precise_precision}
+This section contains additional figures showing the precision of the \textsf{jaxoplanet} implementation and its comparison to the C++ implementation of \textsf{starry}.
 \begin{figure}[H]
@@
         \includegraphics[width=\textwidth]{../workflows/precision/figures/error_SAR.pdf}
-        \caption{Typical errors in the matrices used to compute the flux (\autoref{eq:starry}) with \textsf{starry} and \textsf{jaxoplanet}. \codelink{workflows/precision/scripts/plot_error_SAR}}
+        \caption{Typical errors in the matrices involved in the computation of the flux from \autoref{eq:starry}. \codelink{workflows/precision/scripts/plot_error_SAR}}
         \label{fig:precision_SAR}
@@
 \end{figure}
-
 \begin{figure}[H]
@@
         \includegraphics[width=\textwidth]{../workflows/precision/figures/error_order_degree.pdf}
-        \caption{Maximum errors in the overall flux $f$ for all degrees of the spherical harmonics components for different occultor radii and orders $q$ of the Gauss-Legendre quadrature used to compute the P integral (see \autoref{solution_vector}).
+        \caption{Maximum errors in the overall flux $f$ for all degrees of the spherical harmonics components for different occultor radii and orders $q$ of the Gauss-Legendre quadrature used to compute the $\mathcal{P}$ integral (see \autoref{solution_vector}).
 Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s}, maxima taken over the range of values $b$.}
@@
 These benchmarks are produced for values of $q$ which guarantee that our models reach the same precision as state-of-the-art codes.
-As discussed in \autoref{discussion}, \textsf{jaxoplanet} can be further optimized by reducing the order $q$ of numerical integration while keeping the precision required for specific applications.
+As discussed in \autoref{speed}, \textsf{jaxoplanet} can be further optimized by reducing the order $q$ of numerical integration while keeping the precision required for specific applications.
 

### 2024-10-17 latex/paper.tex

@@
 
-\title{\textsf{jaxoplanet}: Hardware-Accelerated Computation of Stellar Light Curves}
+\title{\textsf{jaxoplanet}: Hardware\,-Accelerated Orbits and Light Curves}
 
 % author info
-\author{Lionel J. Garcia}
+\author[0000-0002-4296-2246]{Lionel J. Garcia}
 \author{Soichiro Hattori}
-\author{Daniel Foreman-Mackey}
+\author[0000-0002-9328-5652]{Daniel Foreman-Mackey}
 \affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY, USA}
@@
 From the orbital characterization of stellar systems to the study of exoplanet and stellar atmospheres.
-In this context, the analytical framework known as \textsf{starry} provided the community with a powerful tool to analytically compute light curves of bodies with non-uniform intensities.
+In this context, the analytical framework known as \textit{starry} provided the community with a powerful tool to analytically compute light curves of bodies with non-uniform surface intensities.
 However, as datasets and models complexity grow, the tractable inference of their physical parameters becomes increasingly challenging, and often motivates biased approximations.
-In this paper, we present a new implementation of the \textsf{starry} framework in \textsf{JAX}, a high-performance machine-learning library designed for large-scale computations.
-We describe the changes made to the original version of \textsf{starry} and evaluate the performance of its new implementation that we release in the open-source Python package \textsf{jaxoplanet}.
+In this paper, we present a new implementation of the \textsf{starry} C++ implementation in \textsf{JAX}, a high-performance machine-learning library designed for large-scale computations.
+We describe the changes made to the original version of \textsf{starry} and evaluate the performance of our new implementation, that we release in the open-source Python package \textsf{jaxoplanet}.
 Through this implementation, we provide the community with a differentiable model of stellar light curves, and enable the use of powerful machine-learning tools available in the \textsf{JAX} ecosystem.
 \gitlink{}
-
 \end{abstract}
@@
 In this context, the development of accurate and robust models of stellar light curves is crucial.\\\\
-In the age of wide-field surveys (such as) and high-precision spectroscopic instruments, these models are being compared to large datasets in probabilistic ways, requiring fast and robust evaluations of their posterior likelihood over larger and larger parameter spaces.
-These inferences can be highly facilitated by gradient-based optimization and sampling algorithms, so that being able to differentiate these models with respect to their parameters offers clear advantages.\\\\
+In the age of wide-field surveys (such as the \textit{Transiting Exoplanet Survey Satellite}; \citealt{tess}) and high-precision spectroscopic instruments, these models are being compared to large datasets in probabilistic ways, requiring fast and robust evaluations of their posterior likelihood over larger and larger parameter spaces.
+These inferences can be highly facilitated by gradient-based optimization and sampling algorithms (see for example \citealt{Betancourt2017}), so that being able to differentiate these models with respect to their parameters offers clear advantages.\\\\
 In this context, \cite{starry} and \cite{Agol2020} presented a framework to analytically compute the light curves of spherical bodies with non-uniform surface intensities, such as stars and exoplanets, and the flux observed from their mutual occultations.
-This framework, known as \textit{starry}, has been successfully applied to a wide variety of science cases, like the study of transiting exoplanets \citep{Kostov2019}, precise modeling of radial velocities \citep{Dawson2021}, the study of exoplanets atmospheres \citep{Bell2023} or the study of stellar photospheres \citep{Almenara2022}.
+This framework, known as \textit{starry}, has been successfully applied to a wide variety of science cases, like the study of transiting exoplanets \citep{Kostov2019}, precise modeling of radial velocities \citep{Dawson2021}, the study of exoplanets atmospheres \citep{Bell2023}, the study of stellar photospheres \citep{Almenara2022}, or even the surface mapping of stellar system objects \citep{Bartolic2022}.
 However, as datasets and models complexity grow, the tractable inference of the larger parameter spaces we have to deal with becomes increasingly challenging, and often motivates biased approximations.
@@
 \citealt{Alderson2023}).\\\\
-In this paper we present \textsf{jaxoplanet}, a new implementation of the \textit{starry} framework in \texttt{JAX}, a high-performance machine-learning library designed for large-scale computations, including autodifferentiation capabilities, deployed on various hardware.
-In \autoref{starry}, we summarize the formalism introduced by \cite{starry} to analytically compute stellar light curves.
+In this paper we present \textsf{jaxoplanet}, a new implementation of the \textit{starry} framework in \textsf{JAX}, a high-performance machine-learning library designed for large-scale computations, including autodifferentiation capabilities, with code compiled and deployable on various hardware accelerators.
+In \autoref{starry}, we summarize the formalism of \textit{starry}, introduced by \cite{starry} to analytically compute stellar light curves.
 In \autoref{optimization}, we describe the \textsf{jaxoplanet} implementation and the changes made to the \textsf{starry} implementation.
 Finally, in \autoref{performance}, we evaluate the performance of our new implementation that we release in the open-source Python package \textsf{jaxoplanet}.
-Through this implementation, we provide the community with a differentiable model of stellar light curves, and enable the use of powerful machine-learning tools available in the \texttt{JAX} ecosystem.\\\\
+Through this implementation, we provide the community with a differentiable model of stellar light curves, and enable the use of powerful machine-learning tools available in the \textsf{JAX} ecosystem.\\\\
+For convenience, we provide links to Python code and Jupyter notebooks ( \codelink{}) to reproduce all of the figures throughout the paper.
+Code to generate this paper is hosted at \href{https://github.com/lgrcia/paper-jaxoplanet}{https://github.com/lgrcia/paper-jaxoplanet} and makes used of \textsf{jaxoplanet} version 3.0.0\footnote{TODO make release}.
 
@@
 
-\subsection{Spherical harmonics}
+\subsection{Spherical harmonics}\label{SH}
 
-The surface of a spherical body can be naturally described by spherical harmonics: a set of special functions defined on the sphere that can serve as orthonormal basis to represent any spherical surface map.
+The surface of a spherical body can be naturally described by real two-dimensional spherical harmonics: a set of special functions defined on the sphere that can serve as orthonormal basis to represent any spherical surface map.
 If $\bvec{\tilde y}$ denotes the spherical harmonics basis of degree $l_{max}$ (defined in \citealt[section 2.2]{starry}) and $\bvec{y}$ a vector that contains the coefficients describing the surface in this basis, the intensity $I$ of the map projected onto the sky-projected $(x, y)$ plane of the observer is
-\begin{equation}I(x, y) = \bvec{\tilde y^\mathsf{T}}(x, y) \, \bvec{y}.\end{equation}
+\begin{equation}\label{eq:I}I(x, y) = \bvec{\tilde y^\mathsf{T}}(x, y) \, \bvec{y}.\end{equation}
 In order to orient this surface, for example with a given inclination, obliquity and phase around its rotation axis, we define a rotation matrix $\bvec{R}$ that acts on the spherical harmonics coefficients $\bvec{y}$, such that the intensity of the map projected onto the sky of the observer is
-\begin{equation}I(x, y) = \bvec{\tilde y}^\mathsf{T}(x, y) \, \bvec{R} \, \bvec{y}.\end{equation}\\
+\begin{equation}\label{eq:IR}I(x, y) = \bvec{\tilde y}^\mathsf{T}(x, y) \, \bvec{R} \, \bvec{y}.\end{equation}\\
 \subsection{Rotation light curve}
@@
 In this new basis, $\bvec{y}$ can be expressed as
-\begin{equation}\bvec{p} = \bvec{A_1}\bvec{y},\end{equation}
+\begin{equation}\label{eq:p}\bvec{p} = \bvec{A_1}\bvec{y},\end{equation}
 such that 
@@
 
-\subsection{Limb-darkened surfaces}
-We notice that a simpler situation is encountered when the surface of the spherical body can be described by a radial intensity profile $I(r)$, such as a limb-darkened star.
-In this case, the vector of spherical harmonics coefficients $\bvec{y}$ is very sparse (because all $m\neq0$ components are zero) and the surface map doesn't need to be rotated.
-Contemporary to \cite{starry}, \cite{Agol2020} explore this case and provide simple expressions for the integrated intensity of a star whose surface can be described by a polynomial limb darkening law occulted by an opaque body.
+\subsection{Limb darkened surfaces}
+A simpler situation is encountered when the surface of the spherical body can be described by a radially-symmetric intensity profile $I(r)$, such as a limb darkened star.
+In this case, the vector of spherical harmonics coefficients $\bvec{y}$ is very sparse (because all $m\neq0$ components are zero) and the surface map doesn't need to be rotated.\\\\
+Contemporary to \cite{starry}, \cite{Agol2020} explore this case and provide simple expressions for the integrated intensity of a star whose surface can be described by a polynomial limb darkening law and occulted by an opaque body.
 In this case:
@@
 \end{enumerate}
-For these reasons, and because of its wide application, our implementation treats radially-symmetric maps separately, leading to highly optimize computations of limb-darkened light curves.\\\\
+For these reasons, and because of its wide application, our implementation treats radially-symmetric maps separately, leading to highly optimize computations of limb darkened light curves.\\\\
 
+\newpage
 \section{Implementation details}\label{optimization}
@@
 In this respect \textsf{JAX} represents an interesting alternative.
-This high-performance machine-learning library can be used to perform hardware-accelerated computations on various type of devices (CPU, GPU, TPU) and gave rise to probabilistic tools developped by an ever-growing community of scientists and engineers.
-However, \textsf{JAX} capabilities really shine when models are written with a functional-programming mindset, and computations performed in a vectorized way.\\\\
+This high-performance machine-learning library can be used to perform hardware-accelerated computations on various types of devices (CPU, GPU, TPU) and gave rise to probabilistic tools developped by an ever-growing community of scientists and engineers.
+However, \textsf{JAX} capabilities really shine when models are written with a functional-programming mindset, and computations performed in a vectorized fashion that allows parallelization.\\\\
 In this section, we describe the changes made to the original \textsf{starry} implementation to allow its rewriting in \textsf{JAX}.
@@
 The multiplication of the base map represented by $\bvec{y}$ with the limb darkening map represented by $\bvec{y}_u$ can be seen as the multiplication of two polynomials, only with special coefficients (called Clebsch–Gordan coefficients) inherent to the properties of spherical harmonics.
-To simplify this operation, we transform both maps into the polynomial basis, so that their product is the actual product of two polynomials which doesn't require the computation of special coefficients.
+To simplify this operation, we first transform both maps into the polynomial basis before multiplying them, so that their product is the actual product of two polynomials which doesn't require the computation of special coefficients.
 
@@
 Instead, one can assume that the total flux of the star should be similar for all observers, independently of the surface map of the star.
-In addition to the fact that all components of the spherical harmonics basis integrate to 0 (except the constant component $Y_{0,0}$) we deduce that the mean flux of the star must be coming from a static surface seen by any observer, no matter his viewpoint, which corresponds to a purely limb-darkened surface (with no other source of non-uniformity).\\\\
-The flux of this purely limb-darkened surface is
+In addition to the fact that all components of the spherical harmonics basis integrate to 0 (except the constant component $Y_{0,0}$) we deduce that the mean flux of the star must be coming from a static surface seen by any observer, no matter his viewpoint, which corresponds to a purely limb darkened disk (with no other source of non-uniformity).\\\\
+The flux of this purely limb darkened disk is
 \begin{equation}
+    \label{eq:fu}
     f_u = \bvec{r}^\mathsf{T}\bvec{A_1}\bvec{y}_u
@@
 However, rotations around the pole, i.e.\,with a rotation axis along $z$, have simpler expressions that can be implemented separately.
-In addition, pre-computing these matrices at lower cost is particularly important for steps 2 and 6, involving a potentially large set of angles $\{\theta\}$ and $\{\theta_z\}$, defined over times $\{t\}$, for which to compute the full light curve.
+In addition, pre-computing these matrices at lower cost is particularly important for steps 2 and 6, involving a potentially large set of angles $\{\theta\}$ and $\{\theta_z\}$, defined over times $\{t\}$, in order to compute the full light curve.
 This explains the current decomposition of the complete rotation in six separate steps.\\\\
-In \textsf{jaxoplanet}, we compute the Wigner-D-matrices by employing the Risbo recursion relations \citep{Risbo1996} implemented in \texttt{JAX} as part of the \texttt{s2fft} Python package \citep{price:s2fft}.
+In \textsf{jaxoplanet}, we compute the Wigner-D-matrices by employing the Risbo recursion relations \citep{Risbo1996} implemented in \textsf{JAX} as part of the \textsf{s2fft} Python package \citep{price:s2fft}.
 In addition, we merge rotations 3, 4 and 5 (see \autoref{fig:rotation_jaxoplanet}) into a single compound rotation of axis
 \begin{equation}
+    \label{eq:v}
     \bvec{v} = \frac{1}{\sqrt{1 - \cos^2{\left(\frac{inc}{2} \right)} \cos^2{\left(\frac{obl}{2}\right)}}} \begin{pmatrix}
@@
 \begin{equation}
+    \label{eq:QP}
     s_n = \mathcal{Q}(\bvec{G}_n) - \mathcal{P}(\bvec{G}_n)
@@
 On the other hand, the $\mathcal{P}$ integral (over the cyan contour shown in \autoref{fig:geometry}) is harder to evaluate and requires the use of recurrence relations and truncated series expansions, with a scheme that depends on the geometric configuration of the system (see \citealt{starry}, Section D.2).\\\\
-In order to translate the \textsf{starry}  implementation to \textsf{JAX}, we evaluate the $\mathcal{P}$ integrals numerically using Gauss-Legendre integrations, requiring the evaluation of the integrated functions only over a fixed number of points $q$.\\\\
+In order to translate this part of the \textsf{starry} implementation to \textsf{JAX}, we evaluate the $\mathcal{P}$ integrals numerically using Gauss-Legendre integrations, requiring the evaluation of the integrated functions only over a fixed number of points $q$.\\\\
 
@@
 \begin{equation}
+    \label{eq:k}
     k = \frac{1-r^2 - b^2 + 2br}{4br},
@@
 where $1 - z^2 = b + r - 2\cos{\varphi '}$ \footnote{note that $\varphi '$ is different from the angle $\varphi$ shown in \autoref{fig:geometry} and was introduced through the change of variable $\varphi ' = \frac{1}{2}(\varphi -\frac{3\pi}{2})$.}.
-\subsubsection{Non-uniform surface}
+\subsubsection{Non-uniform surface}\label{non-uniform}
 In the general case, for a star with a non-radially symmetric surface intensity, $n$ denotes the index of the spherical harmonics component such that
 \begin{equation}
+    \label{eq:n}
     n = l^2 + l + m,
@@
 \begin{equation}
+    \label{eq:munu}
     \begin{aligned}
@@
 \begin{equation}
+    \label{eq:delta}
     \delta=\frac{b-r}{2r}.
@@
 To minimize this number, we note that all integrands are symmetrical functions, except for $\mu=1$ and $l=1$.
-Hence, all integrals can be computed over the interval $\left[0, \frac{\kappa}{2}\right]$, instead of $\left[-\frac{\kappa}{2}, \frac{\kappa}{2}\right]$, and multiplied by 2.
+Hence, these integrals can be computed over the interval $\left[0, \frac{\kappa}{2}\right]$, instead of $\left[-\frac{\kappa}{2}, \frac{\kappa}{2}\right]$, and multiplied by 2.
 We also note that all terms for which $\frac{\mu}{2}$ is even reach machine precision for an integration order as low as $q=20$, that we hold fix.
-With these optimizations, we benchmark the precision and speed of our final implementation for the solution vector in \autoref{performance}.
+With these optimizations, we benchmark the precision and speed of our final implementation for the solution vector and the total flux in \autoref{performance}.
 
@@
 Indeed, iterative methods used in previous implementations (such as \citealt{Raposo2017} and \citealt{Brandt2021}) cannot be easily implemented in \textsf{JAX} and are not extensible to GPU or TPU acceleration.
-Instead, our implementation makes use of \cite{Markley1995} which leads to state-of-the-art precision and runtime performance.\\\\
-Although JAX features automatic differentiation, we find that explicitly implementing the derivatives of the Kepler equation solver is more efficient.
+Instead, our implementation makes use of \cite{Markley1995} which, despite its simplicity, leads to state-of-the-art precision and runtime performance.\\\\
+In addition, although \textsf{JAX} features automatic differentiation, we find that explicitly implementing the derivatives of the Kepler equation solver is more efficient.
 To do so, we use the implicit function theorem to derive the \textit{Jacobian-vector product} required by \textsf{JAX} to specify custom derivatives.
@@
     \item The \texttt{jaxoplanet.light\_curves.emission} module contains functions to compute light curves of emissive limb darkened bodies.
-This module usess the same formalism as the \texttt{limbdark} module, but allows for secondary ecclipses to be computed.
+This module use the same formalism as the \texttt{limbdark} module, but allows for secondary ecclipses to be computed.
     \item The \texttt{jaxoplanet.starry.light\_curves} module contains functions to compute occultation light curve of bodies with non-uniform surfaces \citep{starry}.
@@
 These functions can optionally be transformed using special wrappers provided in the \texttt{jaxoplanet.light\_curves.transforms}  module.
-For example, the \texttt{integrate} transform enables the computation of light curves observed at finite exposure time, using the methods described in \cite{Kipping2010}\footnote{See tutorial at}.
+For example, the \texttt{integrate} transform enables the computation of light curves observed at finite exposure time, using the methods described in \cite{Kipping2010}\footnote{See tutorial at TODO}.
 Additionlly, the \texttt{interpolate} transform can be used to precompute light curves on a grid near a transit, and then interpolate those computations to the required phases when computing the full model, leading to substantial speed ups when several transits are present in the data.\\\\
@@
 \subsection{Precision}\label{precision}
-We evaluate the precision of the \textsf{starry} and \textsf{jaxoplanet} terms against quantities computed at arbitrary precision, using a ground truth version of the code implemented with the \texttt{mpmath} Python library\footnote{\href{https://mpmath.org/}{https://mpmath.org/}}.
+We evaluate the precision of the \textsf{starry} and \textsf{jaxoplanet} terms against quantities computed at arbitrary precision, using a ground truth version of the code implemented with the \textsf{mpmath} Python library\footnote{\href{https://mpmath.org/}{https://mpmath.org/}}.
 The precision of the overall flux $f$ can only be understood by considering the precision of the terms involved in \autoref{eq:starry}: the solution vector $\bvec{s}^\mathsf{T}$ ($\bvec{r}^\mathsf{T}$ if no occultation), the change of basis matrices $\bvec{A_1}$ and $\bvec{A_2}$, and the Wigner-D rotation matrix $\bvec{R}$.
@@
 The disks at the top of the figure show the transit configurations of the occultor (light dark) and the primary surface (light gray) for different values of $b$.
-\codelink{workflows/precision/scripts/plot_error}}
+\codelink{workflows/precision/scripts/plot_error.py}}
         \label{fig:precision_s}
@@
 Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s} over the range of values $b$, but computed with $q=250$.
-\codelink{workflows/precision/scripts/plot_error_vs_degree}}
+\codelink{workflows/precision/scripts/plot_error_vs_degree.py}}
         \label{fig:precision_degree}
@@
 Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s} over the range of values $b$, but computed with $q=100$.
-For each order $n$ all limb darkening coefficients are set to zero except the $n$-th coefficient $u_n=1$.
-\codelink{workflows/precision/scripts/plot_error_limbdark}}
+For each order $n$, all limb darkening coefficients are set to zero except the $n$-th coefficient $u_n=1$.
+The purple points show the \textsf{exoplanet} errors on the flux $f$, limited to the linear and quadratic limb darkening laws.
+In order to benchmark our code (denoted \textsf{jaxoplanet (PLD)} and shown in cyan) for higher order laws, we show the error of the flux computed with the \textsf{starry} (black) and \textsf{jaxoplanet} (blue) codes, using the formalism of \cite{starry} instead \cite{Agol2020}.
+\codelink{workflows/precision/scripts/plot_error_limbdark.py}}
         \label{fig:precision_limbdark}
@@
 Note that \autoref{fig:precision_s} was produced using $q=500$ (right plot) for which $\bvec{s}$ reaches machine precision.
-\codelink{workflows/precision/scripts/plot_br_error}}
+\codelink{workflows/precision/scripts/plot_br_error.py}}
         \label{fig:relative_error_1}
@@
 In this context, numerical integration with an order as low as $q=10$ guarantees that our model reaches a precision of $10^{-16}$, an execution time on par with \textsf{exoplanet} on CPU and up to an order of magnitude faster on GPU for large datasets (see left plot of \autoref{fig:real_benchmark}).
-The second example we consider is the secondary ecclipse mapping of HD189733 b ($r\sim0.3$) using Spitzer's 8 micron channel \citep{majeau2012} obtained at a precision of about 1 ppt.
+The second example we consider is the secondary ecclipse mapping of HD 189733 b ($r\sim0.3$) using Spitzer's 8 micron channel \citep{majeau2012} obtained at a precision of about 1 ppt.
 For a spherical harmonics map of degree $l_{max}=3$, $q=10$ guarantees a model with a precision of $10^{-10}$, an execution time one order of magnitude faster than \textsf{starry} on CPU and up to two orders of magnitude faster on GPU for large datasets (see right plot of \autoref{fig:real_benchmark}).\\\\
@@
         \includegraphics[width=\textwidth]{../workflows/speed/figures/speed_vs_N_realistic.pdf}
-        \caption{Evaluation time versus number of points for different light curves models used in realistic applications.
+        \caption{Evaluation time versus number of points for different light curve models used in realistic applications.
 On the left, we evaluate the time required by \textsf{jaxoplanet} and \textsf{exoplanet} to compute the quadratically limb darkened transit light curve of TRAPPIST-1 b observed with the JWST/NIRISS instrument (for which $q=10$ yields machine precision).
-On the right we evaluate the time required by \textsf{jaxoplanet} and \textsf{starry} to compute the eclipse light curve of HD189733 b whose surface is modeled using spherical harmonics of degree $l_{max}=3$ (with $q=10$ leading to a precision of $10^{-10}$).
-On CPU, our benchmark is done on the single core of an Apple M2 max chip, while on GPU we use the single core of an NVIDIA Tesla V100 processor.}
+On the right we evaluate the time required by \textsf{jaxoplanet} and \textsf{starry} to compute the secondary eclipse light curve of HD 189733 b whose surface is modeled using spherical harmonics of degree $l_{max}=3$ (with $q=10$ leading to a precision of $10^{-10}$).
+On CPU, our benchmark is done on the single core of an Apple M2 max chip, while on GPU we use the single core of an NVIDIA Tesla V100 processor.
+\codelink{workflows/speed/scripts/plot_speed_vs_N_realistic.py}}
         \label{fig:real_benchmark}
@@
 Although this might not be required, \autoref{precise_speed} shows reference speed benchmarks for models evaluated at state-of-the-art precisions.\\\\
-In order to help users set the value of $q$ according to their required precision, the \textsf{jaxoplanet} Python package provide helper functions to estimate the precision of the light curve model given the relative radius $r$, the spherical harmonics degree $l_{max}$ (if general non-uniform maps are required), and the order of the polynomial limb darkening law\footnote{See tutorial at \href{}{}}.\\\\
+In order to help users set the value of $q$ according to their required precision, the \textsf{jaxoplanet} Python package provide helper functions to estimate the precision of the light curve model given the relative radius $r$, the spherical harmonics degree $l_{max}$ (if general non-uniform maps are required), and the order of the polynomial limb darkening law\footnote{See tutorial at TODO \href{}{}}.\\\\
 
@@
 A permanent version of the code used to generate the figures and results in this paper is archived at https://doi.org/10.5281/zenodo.1312286.
+\software{JAX \citep{jax2018github}, starry \citep{starry}, exoplanet \citep{exoplanet}, snakemake \citep{Snakemake}, numpy \citep{harris2020array}, scipy \citep{2020SciPy-NMeth}, s2fft \citep{price:s2fft}, matplotlib \citep{Hunter:2007}, mpmath \citep{mpmath}, equinox \citep{kidger2021equinox}, pint}
 
@@
         \includegraphics[width=\textwidth]{../workflows/precision/figures/error_SAR.pdf}
-        \caption{Typical errors in the matrices involved in the computation of the flux from \autoref{eq:starry}. \codelink{workflows/precision/scripts/plot_error_SAR}}
+        \caption{Typical errors in the matrices involved in the computation of the flux from \autoref{eq:starry}. \codelink{workflows/precision/scripts/plot_error_SAR.py}}
         \label{fig:precision_SAR}
@@
         \includegraphics[width=\textwidth]{../workflows/precision/figures/error_starry.pdf}
-        \caption{Same as \autoref{fig:precision_s} but with $\bvec{s}^\mathsf{T}$ and $f$ computed with the C++ implementation of \textsf{starry}. \codelink{workflows/precision/scripts/plot_error}}
+        \caption{Same as \autoref{fig:precision_s} but with $\bvec{s}^\mathsf{T}$ and $f$ computed with the C++ implementation of \textsf{starry}. \codelink{workflows/precision/scripts/plot_error.py}}
         \label{fig:precision_s_starry}
@@
         \caption{Maximum errors in the overall flux $f$ for all degrees of the spherical harmonics components for different occultor radii and orders $q$ of the Gauss-Legendre quadrature used to compute the $\mathcal{P}$ integral (see \autoref{solution_vector}).
-Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s}, maxima taken over the range of values $b$.}
+Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s}, maxima taken over the range of values $b$.
+\codelink{workflows/precision/scripts/plot_error_several_r_GL_degree_SH.py}}
         \label{fig:precision_order_gausslegendre_degree}
@@
 
-\section{Speed}\label{precise_speed}
+\section{Additional speed benchmarks}\label{precise_speed}
+Unlike \autoref{speed}, this section focuses on speed benchmarks at state-of-the-art precision.
 In \autoref{fig:speed_order_gausslegendre}, we show the evolution of the time required to compute the occultation light curve of a limb darkened star and that of a non-uniform star described by spherical harmonics of degree $l_{max}=20$, depending on the order $q$ of the numerical integration.
@@
 The processing time reported for \textsf{exoplanet} and \textsf{starry} is independent of $q$ as these two codes provide closed-form solutions for the integral $\mathcal{P}$.
-Light curves are computed for $10\,000$ points in transit and an occultor radius $r=0.1$.}
+Light curves are computed for $10\,000$ points in transit and an occultor radius $r=0.1$.
+\codelink{workflows/speed/scripts/plot_speed_vs_order.py}}
         \label{fig:speed_order_gausslegendre}
@@
         \caption{Computation time of the occultation light curve of a limb-darkened star (left) and a non-uniform star described by spherical harmonics up to order $l_{max}=20$ (right) depending on the number of points in the light curve.
-Light curves are computed at $q=200$ and an occultor radius $r=0.1$.}
+Light curves are computed at $q=200$ and an occultor radius $r=0.1$.
+\codelink{workflows/speed/scripts/plot_speed_vs_N.py}}
         \label{fig:speed_N}
@@
 %
+$\bvec{A_1}$ & change of basis matrix from $Y_{lm}$ to polynomials & \autoref{eq:p} \\
+$\bvec{A_2}$ & change of basis matrix from polynomials to Green's basis &  \autoref{occultation}\\
+$A_{kite}$ & Kite-shaped area b/w centers and contact points &  \autoref{fig:geometry}\\ 
+$b$ & impact parameter of the occultor &  \autoref{eq:k}\\
+$f_u$ & flux of a purely limb darkened star& \autoref{eq:fu} \\
+$\bvec{\tilde g}$ & Green's basis vector &  \autoref{occultation} \\
+$\tilde{g}_n$ & n-th component of the Green's basis vector &   \autoref{occultation}\\
+$\bvec{G}_n$ & Green's function &   \autoref{occultation}\\
+$I(x,y)$ & intensity of the body's surface at coordinates $(x,y)$ & \autoref{eq:I} \\
+$l$ & degree of the $Y_{lm}$ & \autoref{SH}\\
+$k$ & Elliptic integral parameter &  \autoref{eq:k}\\
+$l_{max}$ & (maximum) degree of a $Y_{lm}$ map&  \autoref{SH}\\
+$m$ & order of the $Y_{lm}$ &  \autoref{SH}\\
+$n$ & $Y_{lm}$ index in $\bvec{\tilde y}$ &  \autoref{eq:n}\\
+$\mathcal{P}$ & Primitive integral along perimiter of occultor & \autoref{eq:QP} \\
+$\bvec{p}$ & polynomial coefficients describing the the surface & \autoref{eq:p} \\
+$\bvec{\tilde p}$ & polynomial basis vector &  \autoref{eq:p}\\
+$q$ & order of the Gauss-Legendre quadrature &  \autoref{eq:p}\\
+$\mathcal{Q}$ & Primitive integral along perimiter of occulted body & \autoref{eq:QP} \\
+$r$ & occultor radius in units of occulted body's radius & \autoref{eq:k}\\
+$\bvec{R}$ & Rotation matrix of the $Y_{lm}$ basis & \autoref{eq:IR} \\
+$\bvec{R'}$ & Rotation matrix to align the occultor with the $y$-axis & \autoref{occultation} \\
+$s_n$ & n-th component of the solution vector $\bvec{s}^\mathsf{T}$  &\autoref{eq:sn_gn} \\
+$\bvec{s}^\mathsf{T}$ & solution vector & \autoref{occultation} \\
+$\bvec{u}$ & limb darkening coefficients &  \autoref{eq:limb_darkening_map}\\
+$\bvec{U}$ & change of basis matrix from $Y_{lm}$ to limb darkening & \autoref{eq:limb_darkening_map} \\
+$\bvec{v}$ & vector of compound rotation to sky-projected orientation &  \autoref{eq:v}\\
+$x$ & $x$ cartesian coordinate &   \autoref{eq:I}\\
+$y$ & $y$ cartesian coordinate &   \autoref{eq:I}\\
+$\bvec{y}$ & $Y_{lm}$ coefficients describing the surface & \autoref{eq:I} \\
+$\bvec{\tilde y}$ & $Y_{lm}$ basis vector &  \autoref{eq:I} \\
+$Y_{lm}$ & Spherical harmonics basis function &  \autoref{SH}\\
+$\bvec{y}_u$ & limb darkening coefficients of a purely limb darkened star & \autoref{eq:limb_darkening_map} \\
+$z$ & $z=\sqrt{1 -x^2 -y^2}$ cartesian coordinate & \autoref{starry} \\
+$\delta$ & numerical stability parameter &  \autoref{eq:delta}\\
+$\kappa$ & $\varphi + \frac{\pi}{2}$ & \autoref{non-uniform} \\
+$\kappa_0$ & Angular position of body's disks intersection & \autoref{fig:geometry} \\
+$\kappa_1$ & Angular position of body's disks intersection &  \autoref{fig:geometry}\\
+$\lambda$ & angular position of the occultor contact points & \autoref{fig:geometry} \\
+$\mu$ & $l-m$ & \autoref{eq:munu} \\
+$\nu$ & $l+m$ & \autoref{eq:munu} \\
+$\varphi$ & angular position of the occultor contact points & \autoref{fig:geometry} \\
+$\varphi '$ & change of variable for $\varphi$ & \autoref{non-uniform} \\
 $\omega$        & Rotation angle of the combined rotation       & \autoref{eq:combined_angle} \\
-$n$ & order of the Gauss-Legendre approximation & \\
-$r$ & occultor radius in units of occulted body's radius & \\
 %

### 2024-10-17 latex/paper.tex
feat: update readme
reorganize workflow

@@
 Through this implementation, we provide the community with a differentiable model of stellar light curves, and enable the use of powerful machine-learning tools available in the \textsf{JAX} ecosystem.\\\\
-For convenience, we provide links to Python code and Jupyter notebooks ( \codelink{}) to reproduce all of the figures throughout the paper.
+For convenience, we provide links to Python code and Jupyter notebooks (\;\codelink{}) to reproduce all of the figures throughout the paper.
 Code to generate this paper is hosted at \href{https://github.com/lgrcia/paper-jaxoplanet}{https://github.com/lgrcia/paper-jaxoplanet} and makes used of \textsf{jaxoplanet} version 3.0.0\footnote{TODO make release}.
@@
     \begin{center}
-        \includegraphics[width=0.6\textwidth]{../workflows/geometry/figures/rotation_basis.pdf}
-        \caption{Rotation of a primary body's surface to place it in the sky-projected frame of the observer. \codelink{workflows/geometry/scripts/rotations.ipynb}}
+        \includegraphics[width=0.6\textwidth]{../workflows/figures/rotation_basis.pdf}
+        \caption{Rotation of a primary body's surface to place it in the sky-projected frame of the observer. \codelink{workflows/scripts/rotation_figures.py}}
         \label{fig:rotation_basis}
@@
     \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/geometry/figures/rotation_starry.pdf}
+        \includegraphics[width=\textwidth]{../workflows/figures/rotation_starry.pdf}
         \caption{Decomposition of the full rotation of the surface into elementary rotations.
 Each schematic shows a sphere rotated from a previous orientation around an axis displayed as a black vector and an angle shown on top.
-\codelink{workflows/geometry/scripts/rotations.ipynb}}
+\codelink{workflows/scripts/rotation_figures.py}}
         \label{fig:rotation_starry}
@@
     \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/geometry/figures/rotation_jaxoplanet_1.pdf}
+        \includegraphics[width=\textwidth]{../workflows/figures/rotation_jaxoplanet_1.pdf}
         \caption{Consecutive rotations of the spherical harmonics basis in the \textsf{jaxoplanet} implementation.
 Each schematic shows a sphere rotated from a previous orientation around an axis displayed as a black vector and an angle shown on top.
-\codelink{workflows/geometry/scripts/rotations.ipynb}}
+\codelink{workflows/scripts/rotation_figures.py}}
         \label{fig:rotation_jaxoplanet}
@@
     \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/geometry/figures/occultation_geometry.pdf}
+        \includegraphics[width=\textwidth]{../workflows/figures/occultation_geometry.pdf}
         \caption{Geometric configuration and angles definitions for the two-body occultation.
@@
 By choosing an appropriate transform, the integration over $S$ can be reduced to a line integral over the contour of the occulting body, which can be decomposed into the contour of the occulting body within the area of the primary body (in cyan) and the contour of the primary body everywhere else (in blue).
-\codelink{workflows/geometry/scripts/occultation_geometry.ipynb}}
+\codelink{workflows/scripts/occultation_geometry.py}}
         \label{fig:geometry}
@@
     \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/precision/figures/error_jax.pdf}
+        \includegraphics[width=\textwidth]{../workflows/figures/error_jax.pdf}
         \caption{Error of the solution vector $\bvec{s}^\mathsf{T}$ and the overall flux $f$ for all terms of the spherical harmonics basis up to degree $l_{max}=20$.
@@
 The disks at the top of the figure show the transit configurations of the occultor (light dark) and the primary surface (light gray) for different values of $b$.
-\codelink{workflows/precision/scripts/plot_error.py}}
+\codelink{workflows/scripts/plot_error.py}}
         \label{fig:precision_s}
@@
     \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/precision/figures/error_degree.pdf}
+        \includegraphics[width=\textwidth]{../workflows/figures/error_degree.pdf}
         \caption{Maximum errors in the overall flux $f$ for all degrees of the spherical harmonics components.
 Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s} over the range of values $b$, but computed with $q=250$.
-\codelink{workflows/precision/scripts/plot_error_vs_degree.py}}
+\codelink{workflows/scripts/plot_error_vs_degree.py}}
         \label{fig:precision_degree}
@@
     \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/precision/figures/limbdark_error.PDF}
+        \includegraphics[width=\textwidth]{../workflows/figures/limbdark_error.PDF}
         \caption{Maximum errors in the overall flux $f$ of a star occulted by an opaque body for increasing orders of its polynomial limb darkening law.
@@
 In order to benchmark our code (denoted \textsf{jaxoplanet (PLD)} and shown in cyan) for higher order laws, we show the error of the flux computed with the \textsf{starry} (black) and \textsf{jaxoplanet} (blue) codes, using the formalism of \cite{starry} instead \cite{Agol2020}.
-\codelink{workflows/precision/scripts/plot_error_limbdark.py}}
+\codelink{workflows/scripts/plot_error_limbdark.py}}
         \label{fig:precision_limbdark}
@@
     \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/precision/figures/br_error.png}
+        \includegraphics[width=\textwidth]{../workflows/figures/br_error.png}
         \caption{relative numerical error in the flux $f$ of a linearly limb-darkened star (degree $l$=1) transited by an opaque companion over the $(r, b)$ parameter space.
@@
 Note that \autoref{fig:precision_s} was produced using $q=500$ (right plot) for which $\bvec{s}$ reaches machine precision.
-\codelink{workflows/precision/scripts/plot_br_error.py}}
+\codelink{workflows/scripts/plot_br_error.py}}
         \label{fig:relative_error_1}
@@
 %     \begin{center}
-%         \includegraphics[width=\textwidth]{../workflows/precision/figures/error_order.pdf}
+%         \includegraphics[width=\textwidth]{../workflows/figures/error_order.pdf}
 %         \caption{Maximum errors in the solution vector and overall flux $f$ for different occultor radii and orders $n$ of the Gauss-Legendre quadrature used to compute the P integral (see \autoref{solution_vector}).
@@
     \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/speed/figures/speed_vs_N_realistic.pdf}
+        \includegraphics[width=\textwidth]{../workflows/figures/speed_vs_N_realistic.pdf}
         \caption{Evaluation time versus number of points for different light curve models used in realistic applications.
@@
 On CPU, our benchmark is done on the single core of an Apple M2 max chip, while on GPU we use the single core of an NVIDIA Tesla V100 processor.
-\codelink{workflows/speed/scripts/plot_speed_vs_N_realistic.py}}
+\codelink{workflows/scripts/plot_speed_vs_N_realistic.py}}
         \label{fig:real_benchmark}
@@
     \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/precision/figures/error_SAR.pdf}
-        \caption{Typical errors in the matrices involved in the computation of the flux from \autoref{eq:starry}. \codelink{workflows/precision/scripts/plot_error_SAR.py}}
+        \includegraphics[width=\textwidth]{../workflows/figures/error_SAR.pdf}
+        \caption{Typical errors in the matrices involved in the computation of the flux from \autoref{eq:starry}. \codelink{workflows/scripts/plot_error_SAR.py}}
         \label{fig:precision_SAR}
@@
     \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/precision/figures/error_starry.pdf}
-        \caption{Same as \autoref{fig:precision_s} but with $\bvec{s}^\mathsf{T}$ and $f$ computed with the C++ implementation of \textsf{starry}. \codelink{workflows/precision/scripts/plot_error.py}}
+        \includegraphics[width=\textwidth]{../workflows/figures/error_starry.pdf}
+        \caption{Same as \autoref{fig:precision_s} but with $\bvec{s}^\mathsf{T}$ and $f$ computed with the C++ implementation of \textsf{starry}. \codelink{workflows/scripts/plot_error.py}}
         \label{fig:precision_s_starry}
@@
     \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/precision/figures/error_order_degree.pdf}
+        \includegraphics[width=\textwidth]{../workflows/figures/error_order_degree.pdf}
         \caption{Maximum errors in the overall flux $f$ for all degrees of the spherical harmonics components for different occultor radii and orders $q$ of the Gauss-Legendre quadrature used to compute the $\mathcal{P}$ integral (see \autoref{solution_vector}).
 Errors shown in this figure correspond to the maximum values of the scaled errors shown in \autoref{fig:precision_s}, maxima taken over the range of values $b$.
-\codelink{workflows/precision/scripts/plot_error_several_r_GL_degree_SH.py}}
+\codelink{workflows/scripts/plot_error_several_r_GL_degree_SH.py}}
         \label{fig:precision_order_gausslegendre_degree}
@@
     \begin{center} 
-        \includegraphics[width=\textwidth]{../workflows/speed/figures/speed_vs_order.pdf}
+        \includegraphics[width=\textwidth]{../workflows/figures/speed_vs_order.pdf}
         \caption{Computation time of the occultation light curve of a limb-darkened star (left) and a non-uniform star described by spherical harmonics up to order $l_{max}=20$ (right) depending on the order $q$ of the Gauss-Legendre quadrature used to compute the $\mathcal{P}$ integral numerically (see \autoref{solution_vector}).
@@
 Light curves are computed for $10\,000$ points in transit and an occultor radius $r=0.1$.
-\codelink{workflows/speed/scripts/plot_speed_vs_order.py}}
+\codelink{workflows/scripts/plot_speed_vs_order.py}}
         \label{fig:speed_order_gausslegendre}
@@
     \begin{center}
-        \includegraphics[width=\textwidth]{../workflows/speed/figures/speed_vs_N.pdf}
+        \includegraphics[width=\textwidth]{../workflows/figures/speed_vs_N.pdf}
         \caption{Computation time of the occultation light curve of a limb-darkened star (left) and a non-uniform star described by spherical harmonics up to order $l_{max}=20$ (right) depending on the number of points in the light curve.
 Light curves are computed at $q=200$ and an occultor radius $r=0.1$.
-\codelink{workflows/speed/scripts/plot_speed_vs_N.py}}
+\codelink{workflows/scripts/plot_speed_vs_N.py}}
         \label{fig:speed_N}

### 2024-10-17 workflows/precision/scripts/SAR_jax_starry.py

@@ -8,7 +8,7 @@ import numpy as np
 from jaxoplanet.starry.multiprecision import mp
 from jaxoplanet.starry.core.basis import A1, A2_inv
 import jax.numpy as jnp
-from jaxoplanet.starry.s2fft_rotation import compute_rotation_matrices
+from jaxoplanet.starry.core.s2fft_rotation import compute_rotation_matrices
 from jaxoplanet.starry.multiprecision.rotation import R
 
 from jaxoplanet.starry.core.solution import solution_vector

### 2024-10-17 workflows/precision/scripts/make_b.py

@@ -1,7 +1,9 @@
 import numpy as np
 
+RES = snakemake.params.resolution
 
-def b_range(r, logdelta=-3, logeps=-6, res=50):
+
+def b_range(r, logdelta=-3, logeps=-6, res=RES):
     delta = 10**logdelta
     eps = 10**logeps
     if r > 1:

### 2024-10-17 workflows/scripts/rotation_figures.py

@@ -1,0 +1,243 @@
+import jax
+
+jax.config.update("jax_enable_x64", True)
+import jaxoplanet.starry as jerry
+
+import matplotlib
+
+font = {"family": "serif", "size": 12, "serif": "cmr10"}
+
+matplotlib.rc("font", **font)
+matplotlib.rc("text", usetex=True)
+
+import numpy as np
+
+planet = [0.9, -0.5]
+
+inc = 0.8
+obl = 0.5
+theta = 0.2
+theta_z = -np.pi / 2 + np.arctan2(planet[1], planet[0])
+
+np.random.seed(0)
+y = jerry.ylm.Ylm.from_dense(np.hstack([1.0, np.random.rand(10) * 1e-1]))
+
+import matplotlib.pyplot as plt
+
+cmap = "Blues_r"
+
+plt.figure(figsize=(5, 2.5))
+
+u = (0.1, 0.1)
+
+
+def format(ax):
+    ax.set_aspect("equal")
+    ax.axis("off")
+    ax.set_xlim(-1.2, 1.2)
+    ax.set_ylim(-1.2, 1.2)
+
+
+ax = plt.subplot(121)
+s = jerry.Surface(y=y, u=u)
+jerry.visualization.show_surface(s, ax=ax, n=4, cmap=cmap)
+format(ax)
+plt.title("Rest frame")
+
+x = 1.05
+h = 0.08
+l = 0.3
+plt.arrow(-x, -x, 0.0, l, head_width=h, head_length=h, color="0.8")
+plt.arrow(-x, -x, l, 0.0, head_width=h, head_length=h, color="0.8")
+
+letter = dict(fontsize=10, color="0.8")
+plt.text(-x + 1.8 * l, -x, "x", ha="left", va="center", **letter)
+plt.text(-x, -x + 1.8 * l, "y", ha="center", va="bottom", **letter)
+
+ax = plt.subplot(122)
+s = jerry.surface.Surface(y=y, inc=np.pi / 2 - inc, obl=obl, u=u)
+jerry.visualization.show_surface(s, ax=ax, n=4, theta=theta, cmap=cmap)
+disk = plt.Circle(planet, 0.25, color="0.05", zorder=10)
+ax.add_artist(disk)
+plt.title("Sky-projected frame")
+format(ax)
+
+plt.tight_layout()
+plt.savefig(snakemake.output[0])
+
+from scipy.spatial.transform import Rotation
+import numpy as np
+
[…]

### 2024-10-17 workflows/speed/scripts/plot_error_vs_order.py

@@ -13,15 +13,15 @@ exoplanet_time_quadratic = np.median(
     np.load(snakemake.input.exoplanet_quadratic)["time"]
 )
 
-jax_times_l20 = np.array(
-    [np.median(np.load(f)["time"]) for f in snakemake.input.jaxoplanet_l20]
+jax_times_lmax = np.array(
+    [np.median(np.load(f)["time"]) for f in snakemake.input.jaxoplanet_lmax]
 )
 
-jax_times_l20_gpu = np.array(
-    [np.median(np.load(f)["time"]) for f in snakemake.input.jaxoplanet_l20_gpu]
+jax_times_lmax_gpu = np.array(
+    [np.median(np.load(f)["time"]) for f in snakemake.input.jaxoplanet_lmax_gpu]
 )
 
-starry_times_l20 = np.median(np.load(snakemake.input.starry_l20)["time"])
+starry_times_lmax = np.median(np.load(snakemake.input.starry_lmax)["time"])
 
 import matplotlib.pyplot as plt
 
@@ -39,9 +39,9 @@ plt.title("Quadratic limb-darkened map ($l=2$)")
 plt.yscale("log")
 
 plt.subplot(122)
-plt.axhline(starry_times_l20 * 1e3, color="C3", label="starry")
-plt.plot(orders, jax_times_l20 * 1e3, ".-", label="jaxoplanet (CPU)")
-plt.plot(orders, jax_times_l20_gpu * 1e3, ".--", label="jaxoplanet (GPU)", c="C0")
+plt.axhline(starry_times_lmax * 1e3, color="C3", label="starry")
+plt.plot(orders, jax_times_lmax * 1e3, ".-", label="jaxoplanet (CPU)")
+plt.plot(orders, jax_times_lmax_gpu * 1e3, ".--", label="jaxoplanet (GPU)", c="C0")
 plt.legend()
 plt.xlabel("order of Gauss-Legendre quadrature")
 plt.title("Non-uniform map ($l=20$)")

### 2025-01-14 latex/paper.tex
Add part of dan's comments and So comments

@@
 \author[0000-0002-4296-2246]{Lionel J. Garcia}
-\author{Soichiro Hattori}
+\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY, USA}
+\author[0000-0002-0842-863X]{Soichiro Hattori}
+\affiliation{Department of Astrophysics, American Museum of Natural History, New York, NY, USA}
+\affiliation{Department of Astronomy, Columbia University, New York, NY, USA}
 \author[0000-0002-9328-5652]{Daniel Foreman-Mackey}
@@
 From the orbital characterization of stellar systems to the study of exoplanet and stellar atmospheres.
-In this context, the analytical framework known as \textit{starry} provided the community with a powerful tool to analytically compute light curves of bodies with non-uniform surface intensities.
-However, as datasets and models complexity grow, the tractable inference of their physical parameters becomes increasingly challenging, and often motivates biased approximations.
-In this paper, we present a new implementation of the \textsf{starry} C++ implementation in \textsf{JAX}, a high-performance machine-learning library designed for large-scale computations.
-We describe the changes made to the original version of \textsf{starry} and evaluate the performance of our new implementation, that we release in the open-source Python package \textsf{jaxoplanet}.
-Through this implementation, we provide the community with a differentiable model of stellar light curves, and enable the use of powerful machine-learning tools available in the \textsf{JAX} ecosystem.
+In this context, the framework known as \textit{starry} provided the community with a powerful tool to efficiently and accurately compute light curves of bodies with non-uniform surface intensities.
+However, as datasets and models complexity grows, the tractable inference of their physical parameters becomes increasingly challenging.
+In this paper, we present a new implementation of \textsf{starry} built from scratch in \textsf{JAX}, a high-performance machine-learning library designed for large-scale numerical computation.
+We describe the changes made to the original version of \textsf{starry} and evaluate the performance of our new implementation that we release in the open-source Python package \textsf{jaxoplanet}.
+Through this implementation, we provide the community with a fast and differentiable model of stellar light curves, and enable the use of powerful tools available in the \textsf{JAX} ecosystem.
 \gitlink{}
@@
 In this context, the development of accurate and robust models of stellar light curves is crucial.\\\\
-In the age of wide-field surveys (such as the \textit{Transiting Exoplanet Survey Satellite}; \citealt{tess}) and high-precision spectroscopic instruments, these models are being compared to large datasets in probabilistic ways, requiring fast and robust evaluations of their posterior likelihood over larger and larger parameter spaces.
-These inferences can be highly facilitated by gradient-based optimization and sampling algorithms (see for example \citealt{Betancourt2017}), so that being able to differentiate these models with respect to their parameters offers clear advantages.\\\\
+In the age of wide-field surveys (such as the \textit{Transiting Exoplanet Survey Satellite}; \citealt{tess}) and high-precision spectroscopic instruments, these models are being compared to large datasets in probabilistic ways, requiring fast and robust evaluations of their likelihood over larger and larger parameter spaces.
+These inferences can be facilitated by gradient-based optimization and sampling algorithms (see for example \citealt{Betancourt2017}), so that being able to differentiate these models with respect to their parameters offers clear advantages.\\\\
 In this context, \cite{starry} and \cite{Agol2020} presented a framework to analytically compute the light curves of spherical bodies with non-uniform surface intensities, such as stars and exoplanets, and the flux observed from their mutual occultations.
 This framework, known as \textit{starry}, has been successfully applied to a wide variety of science cases, like the study of transiting exoplanets \citep{Kostov2019}, precise modeling of radial velocities \citep{Dawson2021}, the study of exoplanets atmospheres \citep{Bell2023}, the study of stellar photospheres \citep{Almenara2022}, or even the surface mapping of stellar system objects \citep{Bartolic2022}.
-However, as datasets and models complexity grow, the tractable inference of the larger parameter spaces we have to deal with becomes increasingly challenging, and often motivates biased approximations.
+However, as datasets and models complexity grows, tractable inference becomes increasingly challenging, and often necessitates biased approximations.
 This is the case, for example, in transmission spectroscopy measurements conducted with the \textit{James Webb Space Telescope} NIRSpec instrument, that should ideally consist in the simultaneous modeling of more than 3000 transit light curves, one per pixel element (e.g.
 \citealt{Alderson2023}).\\\\
-In this paper we present \textsf{jaxoplanet}, a new implementation of the \textit{starry} framework in \textsf{JAX}, a high-performance machine-learning library designed for large-scale computations, including autodifferentiation capabilities, with code compiled and deployable on various hardware accelerators.
+In this paper we present \textsf{jaxoplanet}, a new implementation of the \textit{starry} framework in \textsf{JAX}, a high-performance library designed for large-scale computations, including autodifferentiation capabilities, with code compiled and deployable on various hardware accelerators.
 In \autoref{starry}, we summarize the formalism of \textit{starry}, introduced by \cite{starry} to analytically compute stellar light curves.
@@
 
-Inspired by the idea of \cite{pal2012}, \cite{starry} and \cite{Agol2020} present a general framework to compute the light curves of spherical bodies with non-uniform surface intensities.
-In this section, we describe the mathematical formalism behind the \textit{starry} model.
+\cite{starry} and \cite{Agol2020} present a general framework to compute the light curves of spherical bodies with non-uniform surface intensities called \textit{starry}.
+In this section, we describe the mathematical formalism behind this model.
 
@@
 In order to simplify this integral, \citealt{starry} introduce a change of basis from the spherical harmonics basis $\bvec{\tilde y}$ to a simple polynomial basis $\bvec{\tilde p}$ (defined in \citealt[section 2.3]{starry}) characterized by the change of basis matrix $\bvec{A_1}$.
-In this new basis, $\bvec{y}$ can be expressed as
-\begin{equation}\label{eq:p}\bvec{p} = \bvec{A_1}\bvec{y},\end{equation}
-such that 
+In this new basis
 \begin{equation}I(x, y) = \bvec{\tilde p^\mathsf{T}}(x, y)\,\bvec{A_1}\bvec{R}\bvec{y}.\end{equation}
@@
 since $\bvec{A_1}$, $\bvec{R}$ and $\bvec{y}$ are independent of the coordinates on the sphere.\\\\
-As $\bvec{\tilde{p}}$ is simply made of polynomials of the coordinates $x$, $y$ and $z$, the elements of the \textit{rotation vector} $\bvec{r}^\mathsf{T}$ can be computed analytically and only once for a given $l_{max}$ (see \citealt[Eq.
+As $\bvec{\tilde{p}}$ is simply made of polynomials of the coordinates $x$, $y$ and $z$, the elements of the \textit{rotation vector} $\bvec{r}^\mathsf{T}$ can be computed in closed form and only once for a given $l_{max}$ (see \citealt[Eq.
 20]{starry}).\\\\
@@
     \item The Green's basis takes a simpler form (see \citealt[Equation 14]{Agol2020}).
-    \item Matrices involved in the computation of the integrated flux have $l_{max}$ columns, versus $(l_{max}+1)^2$ in the general case, reducing drastically the computational cost of the model.
+    \item Matrices involved in the computation of the integrated flux have $l_{max}$ columns, versus $(l_{max}+1)^2$ in the general case, drastically reducing the computational cost of the model.
 \end{enumerate}
-For these reasons, and because of its wide application, our implementation treats radially-symmetric maps separately, leading to highly optimize computations of limb darkened light curves.\\\\
+For these reasons, and because of its wide application, our implementation treats radially-symmetric maps separately, leading to highly optimized computations of limb darkened light curves.\\\\
 
@@
 Limb darkened maps are obtained by multiplying a base map, devoid of limb darkening, with a pure limb darkened map.
-If $\bvec{U}$ is the change of basis matrix that transforms the polynomial limb darkening coefficients $\bvec{u}$ to the spherical harmonics coefficients $\bvec{y}_u$, the pure limb darkened map is
+If $\bvec{\tilde u}$ is the change of basis matrix that transforms the polynomial limb darkening coefficients $\bvec{u}$ to the spherical harmonics coefficients $\bvec{y}_u$, the pure limb darkened map is
 \begin{equation}
     \label{eq:limb_darkening_map}
-    \bvec{y}_u = \bvec{U}\bvec{u}.
+    \bvec{y}_u = \bvec{\tilde u}^\mathsf{T} \bvec{u}.
 \end{equation}
-The multiplication of the base map represented by $\bvec{y}$ with the limb darkening map represented by $\bvec{y}_u$ can be seen as the multiplication of two polynomials, only with special coefficients (called Clebsch–Gordan coefficients) inherent to the properties of spherical harmonics.
+The multiplication of the base map represented by $\bvec{y}$ with the limb darkening map represented by $\bvec{y}_u$ can be seen as the multiplication of two polynomials, only with special coefficients called Clebsch–Gordan coefficients.
 To simplify this operation, we first transform both maps into the polynomial basis before multiplying them, so that their product is the actual product of two polynomials which doesn't require the computation of special coefficients.
@@
 As discussed in the \textsf{starry} documentation\footlink{https://starry.readthedocs.io/en/latest/notebooks/LDNormalization/} this approach is wrong and leads to an unphysical normalization.\\\\
-Instead, one can assume that the total flux of the star should be similar for all observers, independently of the surface map of the star.
-In addition to the fact that all components of the spherical harmonics basis integrate to 0 (except the constant component $Y_{0,0}$) we deduce that the mean flux of the star must be coming from a static surface seen by any observer, no matter his viewpoint, which corresponds to a purely limb darkened disk (with no other source of non-uniformity).\\\\
+As we notice that all components of the spherical harmonics basis integrate to 0 (except the constant component $Y_{0,0}$) we deduce that the mean flux of the star must be coming from a static surface seen by any observer, no matter his viewpoint, which corresponds to a purely limb darkened disk (with no other source of non-uniformity).\\\\
 The flux of this purely limb darkened disk is
@@
     \label{eq:fu}
-    f_u = \bvec{r}^\mathsf{T}\bvec{A_1}\bvec{y}_u
+    F_u = \bvec{r}^\mathsf{T}\bvec{A_1}\bvec{y}_u
 \end{equation}
-with $\bvec{y}_u$ defined in \autoref{eq:limb_darkening_map}. Hence, we normalize limb-darkened maps by $\pi/f_u$, so that the total flux for a uniform map with no limb darkening is unchanged.
+with $\bvec{y}_u$ defined in \autoref{eq:limb_darkening_map}. Hence, we normalize limb-darkened maps by $\pi/F_u$, so that the total flux for a uniform map with no limb darkening is unchanged.
 
@@
         \caption{Decomposition of the full rotation of the surface into elementary rotations.
-Each schematic shows a sphere rotated from a previous orientation around an axis displayed as a black vector and an angle shown on top.
+Each schematic illustrates a sphere undergoing rotation by the angle specified above it.
+Arrows indicate the approximate movement of the sphere between successive configurations.
+The numbers at the bottom of each sphere represent the index of each rotation.
 \codelink{workflows/scripts/rotation_figures.py}}
@@
         \caption{Consecutive rotations of the spherical harmonics basis in the \textsf{jaxoplanet} implementation.
-Each schematic shows a sphere rotated from a previous orientation around an axis displayed as a black vector and an angle shown on top.
+Each schematic illustrates a sphere undergoing rotation by the angle specified above it.
+Arrows indicate the approximate movement of the sphere between successive configurations.
+The numbers at the bottom of each sphere represent the index of each rotation.
 \codelink{workflows/scripts/rotation_figures.py}}
@@
         \\[1em]
-        2s_0 + 4\pi\eta - 2\pi & \qquad n = 2
+        2\,\mathcal{P}(\bvec{G}_0) + 4\pi\eta - 2\pi & \qquad n = 2
         \\[1em]
@@
 Hence, these integrals can be computed over the interval $\left[0, \frac{\kappa}{2}\right]$, instead of $\left[-\frac{\kappa}{2}, \frac{\kappa}{2}\right]$, and multiplied by 2.
-We also note that all terms for which $\frac{\mu}{2}$ is even reach machine precision for an integration order as low as $q=20$, that we hold fix.
+We also note that all terms for which $\frac{\mu}{2}$ is even reach machine precision for an integration order as low as $q=20$, that we hold fixed.
 With these optimizations, we benchmark the precision and speed of our final implementation for the solution vector and the total flux in \autoref{performance}.
@@
 Using all the optimizations described in the previous sections, the \textsf{jaxoplanet} Python package provides code to compute light curves of mutually occulting celestial bodies.
-Our implementation is written in \textsf{JAX} and contains fully vectorized light curve functions that can be compiled and run on several hardware accelerator.
+Our implementation is written in \textsf{JAX} and contains fully vectorized light curve functions that can be compiled and run on several devices such as GPUs and TPUs.
 \begin{itemize}
@@
 \end{figure}
-Assuming that we perform the numerical integration described in \autoref{solution_vector} at a relatively high order $q=500$, \autoref{fig:precision_s} shows that our reimplentation of \textsf{starry} reaches relative errors lower than 1 part per billion for $\bvec{s}^\mathsf{T}$ and the overall flux $f$, for almost all degrees up to $l_{max}=20$ and considering both a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$.
+Assuming that we perform the numerical integration described in \autoref{solution_vector} at a relatively high order $q=500$, \autoref{fig:precision_s} shows that \texttt{jaxoplanet} reaches relative errors lower than 1 part per billion for $\bvec{s}^\mathsf{T}$ and the overall flux $f$, for almost all degrees up to $l_{max}=20$ and considering both a small ($r=0.01$) and a large ($r=100$) occultor transiting the star accross numerically-sensitive values of impact parameters $b$.
 For reference, \autoref{fig:precision_s_starry} shows comparable errors on the same quantities computed with the C++ implentation of \textsf{starry}.
@@
 
-With the results of this section, we validate the precision of \textsf{jaxoplanet} and show it is on per with the legacy C++ \textsf{starry} implentation\footnote{version 1.2.0 \citep{starry_120}} described in \cite{starry}.\\\\
+With the results of this section, we validate the precision of \textsf{jaxoplanet} and show it is on par with the legacy C++ \textsf{starry} implentation\footnote{version 1.2.0 \citep{starry_120}} described in \cite{starry}.\\\\
 
@@
 These observations reached an estimated precision of 89 ppm for a planet with a relative radius of the order $r\sim0.1$.
-In this context, numerical integration with an order as low as $q=10$ guarantees that our model reaches a precision of $10^{-16}$, an execution time on par with \textsf{exoplanet} on CPU and up to an order of magnitude faster on GPU for large datasets (see left plot of \autoref{fig:real_benchmark}).
+In this context, numerical integration with an order as low as $q=10$ guarantees that our model reaches a precision of $10^{-16}$ and execution times on par with \textsf{exoplanet} on CPU and up to an order of magnitude faster on GPU for large datasets (see left plot of \autoref{fig:real_benchmark}).
 The second example we consider is the secondary ecclipse mapping of HD 189733 b ($r\sim0.3$) using Spitzer's 8 micron channel \citep{majeau2012} obtained at a precision of about 1 ppt.
@@
 \section{Conclusion}
-In this paper we presented \textsf{jaxoplanet}, a new implementation of the analytical light curve model \textit{starry} developed using the machine learning library \textsf{JAX}.
-We showed that our implementation is on par with the legacy C++ \textsf{starry} implementation, and that it offers a significant speed advantage on GPUs for large datasets, enabled by \textsf{JAX}'s just-in-time compilation and hardware acceleration.
+In this paper we presented \textsf{jaxoplanet}, a new implementation of the \textit{starry} light curve model developed using the machine learning library \textsf{JAX}.
+We showed that our implementation provides similar numerical and computational performance than the legacy C++ \textsf{starry} implementation, and that it offers a significant speed advantage on GPUs for large datasets, enabled by \textsf{JAX}'s just-in-time compilation and hardware acceleration.
 Our code can be used in association with the large ecosystem of probabilistic and sampling libraries developed in \textsf{JAX}, and will facilitate the development of tools and pipeline for the modeling of TESS and JWST observations, but also future missions such as PLATO \citep{Rauer2014} and ARIEL \citep{Tinetti2018}.\\\\

### 2025-01-14 workflows/scripts/rotation_figures.py

@@ -2,7 +2,8 @@ import jax
 
 jax.config.update("jax_enable_x64", True)
 import jaxoplanet.starry as jerry
-
+import matplotlib.pyplot as plt
+import matplotlib.patches as patches
 import matplotlib
 
 font = {"family": "serif", "size": 12, "serif": "cmr10"}
@@ -65,6 +66,9 @@ format(ax)
 plt.tight_layout()
 plt.savefig(snakemake.output[0])
 
+
+###############
+
 from scipy.spatial.transform import Rotation
 import numpy as np
 
@@ -86,6 +90,7 @@ def plot_lines(lines, ax, pts: int = 100, **kwargs):
     kwargs.setdefault("c", kwargs.pop("color", "0.8"))
     kwargs.setdefault("lw", kwargs.pop("linewidth", 1))
     kwargs.setdefault("alpha", 1.0)
+    kwargs.setdefault("zorder", 0)
 
     jerry.utils.plot_lines(lines[0], ax=ax, **kwargs)
     jerry.utils.plot_lines(lines[1], ax=ax, **kwargs)
@@ -94,16 +99,6 @@ def plot_lines(lines, ax, pts: int = 100, **kwargs):
 
 
 pi2 = "\pi/2"
-
-rotations = (
-    ([1.0, 0.0, 0.0], 0.5 * np.pi, f"${pi2}$"),
-    ([0.0, 0.0, 1.0], theta, r"$\{\theta\}$"),
-    ([1.0, 0.0, 0.0], -0.5 * np.pi, f"$-{pi2}$"),
-    ([0.0, 0.0, 1.0], obl, "$obl$"),
-    ([-np.cos(obl), -np.sin(obl), 0.0], -inc, "$inc - \pi/2$"),
-    ([0, 0, 1.0], -theta_z, r"$\{\theta_z\}$"),
-    ([0, 0, 1.0], -theta_z, "$\\theta_z$"),
-)
 
 
 def plot_vector(u, l=0.3, c="k"):
@@ -125,76 +120,18 @@ def plot_vector(u, l=0.3, c="k"):
         )
 
 
-import matplotlib.pyplot as plt
-from matplotlib.gridspec import GridSpec
-
-
-plt.figure(figsize=(8, 2))
-axes = GridSpec(1, len(rotations), wspace=0.0, hspace=0.0)
-ax = plt.subplot(axes[0])
-
-
-def plot_angles(axes, rotations):
-
[…]
